---
title: "Data, Methods and Descriptive Statistics {#sec-chap2}"
author: "Alexandre Fortier-Chouinard"
bibliography: Thesis.bib
format: pdf
toc: true
toc-depth: 5
abstract: "This chapter seeks to explain the dissertation's datasets, methods, and analyse descriptive statistics about the main explanatory and outcome variables used in the next chapters. It seeks to answer the following question: How can we study the development of political interest through gendered socialization processes? The types of analyses, survey weights, and decisions about which variables to include are provided."
---
@sec-chap1 has established that political interest is an important predictor of political discussion and of more concrete forms of political engagement, which makes this concept worthy of scientific inquiry. Measuring political interest, its development, and the socialization processes that lead to it requires a methodological approach that relies on data collected among both children and adults. It also requires the data collected among children to inquire about all potential sources of socialization --- parents, peers, media, and schools. The measurement of gendered personality traits among children --- self-assertion for boys and co-operation for girls --- is also an important factor that may explain how transmission processes work. This chapter answers the following question: *How can we study the development of political interest through gendered socialization processes?* It starts by detailing the datasets and methods that will be used throughout the chapters of this dissertation. It also provides descriptive statistics about some of the main variables that will be used, notably those related to political interest as well as socio-economic status. Before doing so, it provides a brief explanation of the case selection.

## Why Canada?

The data used for this dissertation comes from Canada, the country I live in and know best. As @noel2014 puts it, studying one's own country produces social scientific knowledge that is directly relevant to and usable by citizens of that country, and it therefore has a uniquely important character. Moreover, for the current study, Canada is a particularly interesting country to study, given that it is often classified by reports as one of the best for women to live in [@conant2019; @em22020; @usnews2020], even though only 30.7% of elected MPs are women, the 61st highest percentage among the world's 193 countries [@ipu2023]. Canadian Election Study (CES) data since 1997 shows that the gender gap in political interest has remained fairly stable despite a recent increase in the percentage of women politicians at the national and provincial levels [@sevi2021]. These characteristics make Canada a country worth studying to better understand the underlying reasons behind the stability of the gender gap in the aggregate measure of political interest, which mainly taps into interest in partisan politics. Studies that measure interest in different political topics have been conducted in Europe and the Middle East, but it is unclear how the gender differences they find for many topics apply in Canada. Studying the transmission of political interest to children in Canada seems like a promising place to look for explanations of these gender differences that may, or may not, apply elsewhere.

## Data

Given the abstract nature of political interest, the observation of children's behaviour would not provide an accurate answer to this dissertation's questions. Political interest is different from frequency of political discussions, and different from political participation. Surveys, compared with interviews or focus groups, make it easier to get reliable interest scores that can be compared across respondents of different backgrounds. This dissertation therefore relies on survey data.

Within Canada, this dissertation relies on five datasets as well as the 2021 Canadian Census. Two of these datasets have been collected for this dissertation while the other three are large-scale surveys mobilized to provide extra context and better estimates of political engagement in Canada --- and its evolution as Canadians age. The two datasets which were collected for this dissertation share similar questions about political interest and about interest in five specific political topics: health care, international affairs, law and crime, education, and partisan politics. These five topics are those for which past studies have reported the largest gender differences in interest: women typically report being more interested in the political aspects of health care and education, while men typically report being more interested in the political aspects of the other three topics mentioned [@campbellrosie2008; @coffe2013; @ferrin2020; @hayes1993; @kuhn2004; @sabella2004; @verba1997].

### Canadian Children Political Interest Survey (CCPIS)

To create the Canadian Children Political Interest Survey (CCPIS), data was collected among 698 Canadian children and adolescents aged 8 to 18 from seven elementary and secondary schools and one school board's Student Senate, all located in various urban areas of Quebec and Ontario,^[All are considered urban since they are part of a census metropolitan area (CMA) as defined by @statcan2023.] between August 2022 and January 2023. Students had to fill a 15-minute online survey questionnaire during classroom time. The survey was hosted on the Qualtrics Web platform and was available in both French and English. It seems worthwhile to study political socialization in the period of life where political interest is developed to better understand gender differences since there seems to be some level of path dependency in individuals' political interest afterwards [@prior2010; @prior2019].

@tbl-descriptive shows the number of classrooms and students in each of the eight schools included in the final sample of the CCPIS, as well as additional information about the schools. Participating schools include a mix of private and public schools in Quebec and Ontario, including students from all age groups.

Table: Descriptive statistics, CCPIS data {#tbl-descriptive}

| **ID** | **Type**^[The three public bodies that accepted to be part of the study are associated with three different school boards.] | **Language** | **Province** | **Ages**^[Age groups are for the school itself, not the classrooms selected.] |  **Number of students in body**  | **Number of students in sample** | **Number of classrooms in sample**  | **Number of teachers in sample** |
|---|------------|--------------------|------------------|-------------|--------------------|-----------------------|----------------------|------------------------|
| 1           | Private          | French            | Quebec            | 12--17        | 450               | 133                | 5                   | 2                  |
| 2           | Public           | French            | Quebec            | 12--17        | 690               | 196                | 10                  | 2                  |
| 3           | Private          | French            | Quebec            | 12--17        | 670               | 78                 | 3                   | 1                  |
| 4           | Private          | French            | Quebec            | 12--17        | 900               | 253                | 12                  | 3                  |
| 5^[Mixed on-site/online school.]           | Private          | English           | Ontario           | 14--18        |                   | 5                  | 3                   | 2                  |
| 6           | Public           | French            | Quebec            | 5--12         |                   | 14                 | 1                   | 1                  |
| 7           | Private          | English           | Ontario           | 5--14         |                   | 4                  | 3                   | 1                  |
| 8^[Student Senate, a body located at the school board level.]           | Public           | English           | Ontario           | 14--18        | 15                | 15                 | 1                   | 1                  |
|             |                  |                   |                   |               | **Total**    | **698**       | **38**         | **13**        |
<!--School ID 1: Collège Boisbriand 2016, nombre d'élèves https://cdn-contenu.quebec.ca/cdn-contenu/adm/min/education/publications-adm/education/organismes-lies/CCEP/CCEP-rapport-annuel-20-21.pdf
School ID 2: École Jean-de-Brébeuf, nombre d'élèves https://cssc.gouv.qc.ca/ecole/ecole-jean-de-brebeuf
School ID 3: Collège mariste de Québec, nombre d'élèves https://www.xn--collgemariste-zgb.ca/historique
School ID 4: Collège Citoyen, nombre d'élèves https://ecolespriveesquebec.ca/ecole/college-citoyen/
School ID 5: Jaya International High School
School ID 6: École de la Rose-des-Vents
School ID 7: Urban Village Academy
School ID 8: Renfrew County DSB Student Senate, number of students https://www.rcdsb.on.ca/en/parents-and-students/student-senate.aspx-->

In total, 75 school boards,^[This figure also includes institutions known under names such as school service centers, school divisions, school districts, school councils, and centers for education.] 47 public schools, and 83 private schools across Canada were originally contacted to take part in the project. Of the 75 school boards that were contacted, 10 accepted to be part of the study (13%), 15 refused (20%), 27 did not reply to emails (36%), and 23 did not follow through or required extensive information that was impossible to provide (31%). On that last point, several school boards required extensive documentation about the study or background police checks, even for virtual data collection, and most of those who were sent that information still refused to take part in the study after the documentation was provided to them, often citing a lack of time. Of the 132 schools that were contacted, 10 accepted (8%), 9 refused (7%), 2 were deemed ineligible after checking (2%), 88 did not respond (67%) --- in most cases after several attempts --- and 21 stopped replying after a few email exchanges (16%). Among the schools that accepted to take part in the study, 13 teachers made their students fill the survey (59% of all teachers contacted).^[Non-response and rejection rates after the first contact are not reported by @dostiegoulet2009en, @prior2019, @neundorf2013 or other similar studies about political interest among children and teenagers.] Within each classroom, 18 students filled the survey on average --- a number that rises to 22 per classroom when removing school 5, whose students did not have an allocated classroom period to fill the survey and therefore did so out of their free time. 22 per classroom is a number very close to the average 20 to 24 students per classroom in Ontario and Quebec schools [@bolduc2019].^[@dostiegoulet2009fr and @dostiegoulet2009en also reported high within-class response rates of 90% and 80% respectively.] In total, the 698 students who took part in the survey, while not as large as the 1,000 generally reached in regular cross-country opinion polling in Canada, make it possible to analyze trends that go beyond the local or school level with sufficient statistical power.

Low response rates for schools and school boards imply that the sample's descriptive characteristics should not be seen as representative of the Canadian student population. This sample should be considered a non-random, convenience sample. Nevertheless, as will become clearer when examining student-level descriptive statistics, there is no theoretical reason to believe that the types of correlations studied here --- transmission of political interest by role models --- should be affected by the convenience nature of the sample. Given the scarcity of available data on school-aged children's political predispositions within Canada, this dataset also provides relevant information about students' interests, role models, and relation to politics. The core of the data analysis in each chapter therefore relies on this dataset.

The CCPIS includes information about students' interest in various political topics, the political topics they say they discuss with their mothers, fathers and peers, as well as the political topics discussed by a teacher and an influencer they like. The gender of these role models is also inquired, to test the implication from social learning theory that political interest transmission should occur mainly between a child and role models of their gender, more than role models of other genders [@prior2019; @shehata2019].

For questions about students' political interests, the following questions are asked: "How interested are you in politics generally? Set the slider to a number from 0 to 10, where 0 means no interest at all, and 10 means a great deal of interest", and "If you were to open a news website and see the following articles how interested would you be in reading each article? Set the slider to a number from 0 to 10, where 0 means 'Not at all interested, I would not read it,' and 10 means 'Very interested, I would most likely read it.' (a) Health care (i.e., pandemic restrictions, working conditions of nurses); (b) International affairs (i.e., diplomatic disputes between Canada and China, Ukrainian war); (c) Law and crime (i.e., police funding, sentences for violent crimes); (d) Education (i.e., university tuition, funding of public and private schools); (e) Partisan politics (i.e., federal elections, political parties)". These questions specify concrete examples of political issues related to each topic, in order to make the political aspects of these topics more salient and avoid students answering while thinking about their own health or classes they are taking. The question's phrasing is meant to be easily understood by children and teenagers.

The main socialization-related explanatory variables for children's political interest identified in the literature concern the role of four agents of socialization: parents, peers, schools, and media. In order to determine the importance of these role models' gender, the following questions were asked to children: "Which parent do you discuss most often with? (a) Mother; (b) Father; (c) Both equally"; "What is the gender of most of your friends? (a) Girls; (b) Boys; (c) About the same for both genders"; "Think about a teacher that you like(d). Is that teacher... (a) A woman; (b) A man; (c) Other (e.g. Trans, non-binary, two-spirit, gender-queer)"; "Think about someone that you like and sometimes read or watch on social media --- including YouTube. Is that person... (a) A man; (b) A woman; (c) Other (e.g. Trans, non-binary, two-spirit, gender-queer)." These questions make it possible to assess the influence of role models, but also to test its interaction effect with role models' gender. Questions are then asked about the topics most discussed by parents, peers, teachers and influencers among the five political topics highlighted beforehand.

The CCPIS also uses the Personal Attributes Questionnaire [@spence1978; @ward2006] to assess students' degrees of agency and communality, and to assess whether interest correlations between children and their same-gender role models change when controlling for the child's degree of agency or communality. Men are typically more assertive, which brings them closer to partisan politics, while women are typical more co-operative, which leads them to be more interested in topics such as health care and education. Given the importance of the agency/communality distinction in explaining gender differences in political interest [@campbellrosie2008], this hypothesis is tested alongside hypotheses about the role of parents, peers and other role models.

Finally, the CCPIS asks socio-demographic questions about children's gender, language spoken at home, immigrant status, age, and race. These questions are mostly inspired by the Canadian Election Study [@ces2019], except for the question on race, which is based on the 2016 Census [@census2017]. The English and French CCPIS questionnaires are available in @sec-appendix1 and  @sec-appendix2.

### Datagotchi Post-Election Survey (Datagotchi PES)

The second dataset collected over the course of this dissertation is the 2022 Quebec Datagotchi Post-Election Survey, which includes data on 2,228 Quebec adult respondents [@datagotchi2023]. This dataset was collected in February and March 2023 using a panel of respondents which agreed to be contacted by email during the 2022 Quebec general election through the Datagotchi Web app [@datagotchi2022]. Opt-in surveys like this one are typically more reliable but less externally valid than traditional surveys [@herrick2019; @thielo2021]. For the purposes of this dissertation, the questions of interest which were asked to panel respondents inquired about their level of interest in the five aforementioned political topics and general socio-economic status. For these questions, the same question wording was used than among students, with the same five political topics. The Datagotchi Post-Election Survey French questionnaire is available in @sec-appendix3.

### Cross-Country Datasets

Other than these two main datasets, three often-used datasets including Canadian respondents are mobilized to give a more general empirical overlook on political interest in Canada. These datasets include only one question about political interest --- rather than interest in five topics --- but each of them asks about other forms of political engagement --- which will further be analyzed in @sec-chap3 --- and two of them provide time-series data about political interest and engagement, while the third provides a less political questionnaire and therefore provides data about a different crowd of Canadian respondents which may be more representative of Canadians' overall level of interest in politics.

First, Canadian Election Study (CES) data on various aspects of political engagement is analyzed [@ces2021]. Other than typical socio-economic status questions, the CES includes data about political interest, political knowledge, internal efficacy, external efficacy, political discussion, political debating, participation in protests, participation in boycotts, participation in petitions, party membership, donations to parties, and voting. CES data allows time series analyses of the evolution of gender gaps, with some political engagement questions having been asked since 1965.

Second, time-series World Values Survey (WVS) data is also used. These data, like CES data, make it possible to visually represent descriptive statistics by age and gender on political engagement. Moreover, the World Values Survey (WVS) Wave 7 (2017--22) includes data from Canada (2020) and 56 other countries, therefore providing the opportunity for cross-country comparisons on political interest and political engagement [@wvs2022]. Other than typical socio-economic status questions, respondents were asked a different political interest question than in other surveys: "How interested would you say you are in politics? (a) Very interested; (b) Somewhat interested; (c) Not very interested; (d) Not at all interested" [@wvs2022].^[This 4-point scale is transformed into a 0--10 scale comparable with the CES's political interest scale, using the same method as @prior2010.] Time-series data about political interest are also used in some analyses, with political interest questions being asked as early as 1982 in some countries.

Third, the 2013 and 2020 Canadian General Social Survey (GSS), cycles 27 and 35 (Social Identity), are also used to get data about Canadians' political interest and engagement coming from a less political survey. The GSS's political interest questions asks "Generally speaking, how interested are you in politics? (e.g., international, national, provincial or municipal) (a) Very interested; (b) Somewhat interested; (c) Not very interested; (d) Not at all interested" [@gss2020].

### Canadian Census

Finally, data from the 2021 Canadian Census is used to compare data from all other sources to Canadian population figures [@census2022] --- or Quebec population figures, in the case of the Datagotchi PES. More specifically, the Individual Public Use Microdata Files (PUMF) are used to cross-reference data. The PUMF includes data collected among 980,868 Canadians, or a representative segment of 2.7% of the Canadian population. These data are specifically meant to be used to create survey weights and calculate aggregate statistics about the Canadian population, which I do with the Datagotchi PES data.

## Methods

### Avoiding Social Desirability Bias

Questions about political interest can be subject to some degree of social desirability bias, which is "the difference between an individual's own intention and his/her perception of his/her peers' intention" [@chung2003, p.&nbsp;296]. Respondents may want to portray themselves as "good citizens" who are interested in public affairs, regardless of their real interest in politics. However, it is not entirely clear whether girls' scores or boys' scores would be more inflated as a result of social desirability bias. Studies have found that in Canada, Australia, and the United States, *women* generally exhibit higher social desirability bias when answering survey questions [@chung2003; @cohen1998; @cohen2001]. However, studies in these same countries and more instead show that *men* are more likely to over-report having voted in the previous elections [@belli2001; @herrick2020; @stockemer2023]. For political interest, it is unclear which gender --- if any --- tends to over-report it, but disaggregating the concept in five topics might be the best approach to avoid systematic social desirability bias associated with one particular question. This is therefore the approach taken throughout this dissertation.

### Controlling for Classroom-Level Effects

In the CCPIS dataset, students are clustered within classrooms, which are themselves nested within schools and provinces. There may be some schools or classrooms in which relationships between socialization and political interest develop differently. Multilevel regressions with two levels are therefore used to help disentangle classroom-level effects from individual-level effects. Since the numbers of schools (8) and provinces (2) in the final sample are too low for multilevel regressions to be conducted, only classroom-level fixed effects are used in the analyses.

Empty models are calculated to assess what percentage in the total variance in political interest --- and in each of the five topics --- is located at the classroom level. Results show that 6.2% of variance in political interest is located at the classroom level, as well as 4.6% of variance in interest in health care, 3.6% of variance in interest in international affairs, 1.4% of variance in interest in law and crime, 8.1% of variance in interest in education, and 2.6% of variance in interest in partisan politics. In all cases, the bulk of variance therefore seems to be located at another level than the classroom --- presumably the individual level.

The four datasets collected among adults do not involve individuals clustered within classrooms or other broader groupings of interest for this study. Instead of multilevel regressions, simple and multiple ordinary least squares (OLS) regressions are instead performed.

### What are Potential Confounding Variables to Links Between Socialization and Political Interest?

Other than @sec-chap3, which takes a time-series approach to studying political interest, most analyses of the coming chapters have political interest as the outcome variable and rely on CCPIS data. Typically, a simple regression model with one explanatory variable --- often a given role model's interest in a specific topic --- is followed by a multiple regression models with control variables for three blocs of variables: (1) personality traits, (2) role models' gender, and (3) socio-economic status variables. Personality traits include the agency and communality scales described in the following section, while socio-economic status variables include confounding variables which have been linked with political interest and could therefore mediate any given relationship found between two persons' political interest. These include gender, age --- @sec-chap3 explores the relationships between gender, age and political interest in more detail --- as well as language, ethnicity, and immigrant status. For analyses involving adult data, income and level of education, both of which can also be linked to political interest, are also added as control variables.

## Descriptive Statistics

### General

Before analyzing average levels of interest in politics in each of the datasets, it is first important to assess the extent to which each of them is (un)representative of the Canadian population on some of the main socio-economic status variables that may relate to political interest, including age, gender, education, income, and so on. In cases where the data is deemed unrepresentative, a weighting method is applied and described.

@fig-ccpis1 shows the distribution of CCPIS students by age, gender, language, race, immigrant status, degree of agency and communality, and family situation. Most respondents are aged 12 to 17, with a large proportion (41%) being exactly 16 years old. This roughly corresponds to the moment when adolescents' political interest starts increasing alongside a gender gap in which boys report being more interested in politics [@fraile2020; @janmaat2022; @prior2019]. This age distribution should make it possible to assess the influence of all types of role models at the moment when they are the most likely to be influential for students' political interest. The sample is therefore compared with 2021 Census microdata on 12--17-year-old Canadian teenagers. The gender composition of the sample is roughly balanced, with 50.4% of girls and 47.1% of boys while, among all Canadians, 51.5% of teenagers are boys and 48.5% are girls.^[The number of transgender and non-binary people for this age group is not disclosed in the microdata.] Respondents are mainly white --- 57.9%, which is close to the Canadian average of 59.8%. They are also overwhelmingly Francophone due to the over-representation of Quebec within sampled schools --- 66.5% mostly speak French at home, which is much higher than the Canadian average of 18.3% --- but several students who mainly speak another language than English or French at home have also filled the survey --- 25%, which is higher than the Canadian average of 13.2%. 84.4% of students in the sample were born in Canada --- close to the 86.8% countrywide average. The distributions of agency and communality among students both follow a bell-shaped distribution centered around the averages of 0.62 and 0.68 respectively. Most students live with one mother and one father (71.8%), with a substantial minority (18.3%) also having stepparents. Other family arrangements are less common. Other than language and province, these students do not appear to be markedly different from the Canadian teenager population on the main socio-demographic variables. Descriptive statistics about role models' gender and interest in political topics are analyzed in @sec-chap4, @sec-chap5, and @sec-chap6, alongside further analysis related to these role models' influence on children's political interest.

```{r}
#| label: fig-ccpis1
#| echo: false
#| fig-cap: "CCPIS Descriptive Statistics - General"
knitr::include_graphics("_graphs/CCPISDescriptive1.pdf")
```

For agency and communality, scales are replicated from @ward2006. Factor analysis in @fig-factor1 and @fig-factor2 show that, with one exception on the agency list --- an item that was reverse-coded^[This could be a data quality issue, since this is the only question for which the agentic pole is reverse-coded among the first eight questions asked to children, and none of the communality scale questions are reverse-coded. This means some children might have paid less attention to the actual question and responded quickly. However, given the fact that the non-reverse-coded measure is also unrelated to other elements of the scale, and since children responded differently to the agency and communality scales, this seems to be a minor concern. The scale is an exact replication of the one used by @ward2006.] --- all items scale well together, with factor loadings at least medium --- above 0.3 [@shevlin2000] --- and first eigenvalues larger than the conventionally accepted value of 1 [@williams2010]. The Cronbach's alpha for the agency scale (0.63) is slightly under the 0.7 to 0.9 range suggested by scholars [@tavakol2011], while it is within that range for the communality scale (0.78), therefore meeting the standard benchmarks for valid and reliable scales established in peer-reviewed studies about measurement scales.

```{r}
#| label: fig-factor1
#| echo: false
#| fig-cap: "CCPIS Factor Analysis: Agency Scale"
knitr::include_graphics("_graphs/AgencyScale.pdf")
```

```{r}
#| label: fig-factor2
#| echo: false
#| fig-cap: "CCPIS Factor Analysis: Communality Scale"
knitr::include_graphics("_graphs/CommunalityScale.pdf")
```

@fig-dg shows descriptive statistics for the Datagotchi PES, which is not a representative sample before a form of weighting is applied. The unweighted sample includes an over-representation of men (56.8%), white people (97.2%), Francophones (94% responded to the French version of the survey), university-educated people (71%), people born in Canada (86%), and those with a yearly household income over \$110,000 (45%). By comparison, Census numbers among adult Quebeckers are 49.3% of men, 82.1% of white people, 82.1% of people who speak French more than English, 28.2% university-educated people, 80.7% people born in Canada, and 34.3% with a yearly household income above \$110,000 [@census2022]. The median age is 46, with a bimodal distribution centered around 35--40 and 65, compared with a median age within the 50--54 years old range among adult Quebeckers.

```{r}
#| label: fig-dg
#| echo: false
#| fig-cap: "Datagotchi PES Descriptive Statistics - General"
knitr::include_graphics("_graphs/DGDescriptive1.pdf")
```

Given the mismatch between many of these characteristics --- notably gender, level of education and race --- and the Quebec population, survey weights are created and used to make the the Datagotchi PES sample more representative of the Quebec population on these two aspects, as well as income and age. Using the `anesrake` package which is used to weight results from the American National Election Studies [@pasek2018], a raking procedure is used to produce these weights, which range from 0 to 5. Default settings of the `anesrake` function are used. After using this procedure, the population percentages match with sample percentages within 10 percentage points --- and typically less than 5 --- for each category of the five variables.

@fig-ces shows descriptive statistics for the 2021 CES. Compared to 2021 Canadian Census data collected among adults [@census2022], the unweighted sample includes a slight over-representation of women --- 54.5% vs. 51.1% --- and a larger over-representation of Allophones --- 33.1% vs. 16.5% --- of white people --- 89.1% vs. 70.2% --- of university-educated people --- 54.4% vs. 30.7% --- and of people born in Canada --- 84.4% vs. 70.5%. The median yearly household income is within the \$60,001--\$90,000 range, which includes the narrower \$85,000 to \$89,999 median range among adult Canadians. The median age is 53, slightly above the median 45--49 age range for the adult Canadian population, and there is again a bimodal distribution centered, this time centered around 30--35 and 60--70. Respondents from Ontario are slightly under-represented --- 35.0% vs. 38.6% --- while those from Quebec are over-represented --- 29.6% vs. 22.9%. Given the mismatch between many of these characteristics --- notably province, education and age --- and the Canadian population, the CES's survey weights based on the 2016 Census, which account for province, gender, age, and education, are used. These weights' values vary between 0.2 and 5. After using this procedure, the population percentages match with sample percentages on these aspects.

```{r}
#| label: fig-ces
#| echo: false
#| fig-cap: "2021 CES Descriptive Statistics - General"
knitr::include_graphics("_graphs/CESDescriptive1.pdf")
```

@fig-wvs shows descriptive statistics for the 2020 WVS in Canada. Compared to 2021 Canadian Census data collected among adults, the unweighted sample includes a similar proportion of women --- 48.8% vs. 51.1% --- an under-representation of Allophones --- 6.0% vs. 16.5% --- and an over-representation of university-educated people --- 55.5% vs. 30.7% --- of white people --- 80.5% vs. 70.2% --- and of people born in Canada --- 82.1% vs. 70.5%. The median age is 45, which broadly matches with the median 45--49 age range for the adult Canadian population. Respondents from Ontario are under-represented --- 25.0% vs. 38.6%. Given the mismatch between many of these characteristics --- notably province and education --- and the Canadian population, the WVS's survey weights for Canada based on the 2016 Census, which account for age, gender, education, and region, are used. These weights' values vary between 0.12 and 3.69. After using this procedure, the population percentages match with sample percentages on these aspects.

```{r}
#| label: fig-wvs
#| echo: false
#| fig-cap: "2020 WVS Descriptive Statistics - General"
knitr::include_graphics("_graphs/WVSDescriptive1.pdf")
```

@fig-gss shows descriptive statistics for the 2020 GSS. Compared to 2021 Canadian Census data collected among adults [@census2022], the unweighted sample includes a similar proportion of women --- 51% vs. 51.1% --- an over-representation of Allophones --- 23.1% vs. 16.5% --- of university-educated people --- 40.8% vs. 30.7% --- of visible minorities --- 42% vs. 29.8% --- and of immigrants --- 41.1% vs. 29.5%. The median yearly household income is within the \$50,000--\$74,999 range, compared with the higher \$85,000 to \$89,999 median range among adult Canadians. The median age is within the 45--54 age range, which broadly matches with the median 45--49 age range for the adult Canadian population. Respondents from Ontario are slightly under-represented --- 34.4% vs. 38.6%. Given the mismatch between many of these characteristics and the Canadian population, the GSS's survey weights based on the 2016 Census, which account for gender, age, province, CMA, and visible minority status, are used. These weights' values vary between 1 and 32,631. After using this procedure, the population percentages match with sample percentages on these aspects.

```{r}
#| label: fig-gss
#| echo: false
#| fig-cap: "2020 GSS Descriptive Statistics (Cycle 35 - Social Identity) - General"
knitr::include_graphics("_graphs/GSSDescriptive1.pdf")
```

Overall, the four studies conducted among adults have samples that vary from each other. Some of them over-sample immigrants while others under-sample them; some over-sample older citizens while others under-sample them. A common theme in the three studies of Canadian adults --- CES, WVS and GSS --- is the under-sampling of Ontarians. However, all these variations are corrected using raking or post-stratification weights.

### Political Interest

After survey weights are applied, what is the average level of political interest in each of the datasets? What is the distribution of political interest among respondents, and which political topics are Canadians most interested in? By looking at the distribution of political interest among respondents topic by topic, this section lays the ground for future chapters who will try to explain these political interest scores.

@fig-ccpis2 shows the distribution of CCPIS students by interest in politics by topic. 0 is associated with the lowest level of interest and 10 with the highest. For all topics, 9 and 10 are among the least common answers. International relations, and law and crime are generally well-balanced, with students' mean interest above 5/10 (5.2--5.3). On the other hand, fewer students report high interest in health care (4.1), education (4.1), or partisan politics (3.5). Partisan politics in particular has a relatively steady positive skew, where the mode is a 0/10 level of interest. For other topics, the distributions are relatively wide, often taking a flattened bell shape.

```{r}
#| label: fig-ccpis2
#| echo: false
#| fig-cap: "CCPIS Descriptive Statistics - Political Interest"
knitr::include_graphics("_graphs/CCPISDescriptive2.pdf")
```

@fig-dginterest and @fig-ceswvsgssinterest show the level of interest in politics among the four adult Canadian surveys, including interest in the five topics in the case of the Datagotchi PES, after survey weights are applied. While adult Quebeckers (Datagotchi PES) report being more interested in each of the five topics compared with Canadian students (CCPIS), interest in law and crime (5.5/10) is lower than for other topics, followed by partisan politics (5.9/10). General political interest (7.1/10) is higher on average than interest in any of the five topics. Interest in the five political topics also follows a more bell-shaped, normal distribution than what CCPIS found among Canadian youth, although a negative skew can be found for all topics. Canadian WVS respondents are slightly less interested in politics (5.6/10) than those to the 2020 GSS and 2021 CES (6.0/10 and 6.1/10 respectively). Datagotchi PES respondents score an even higher average general political interest score of 7.1/10, which might not be surprising given the fact that these respondents are former users of the Datagotchi Web app. Analyses using the Datagotchi PES throughout this dissertation can therefore be considered to be a high bar for level of interest in each of the political topics.

```{r}
#| label: fig-dginterest
#| echo: false
#| fig-cap: "Datagotchi PES Descriptive Statistics - Political Interest"
knitr::include_graphics("_graphs/DGInterest.pdf")
```

```{r}
#| label: fig-ceswvsgssinterest
#| echo: false
#| fig-cap: "CES, WVS and GSS Descriptive Statistics - Political Interest"
knitr::include_graphics("_graphs/CESWVSGSSInterest.pdf")
```

How can one make sense of these numbers? Are they relatively low or high compared with what other studies have found in various countries among respondents of different age groups? The next chapter will provide more thorough answers to these questions while studying what period of life is associated with the development of political interest.

## References
