---
title: "Data, Methods and Descriptive Statistics"
author: "Alexandre Fortier-Chouinard"
bibliography: Thesis.bib
format: pdf
toc: true
toc-depth: 5
---
Political interest is an important predictor of political discussion and of more concrete forms of political engagement. Measuring political interest and the socialization processes that may lead to it require a well-integrated methodological approach. This chapter details the datasets and methods that will be used throughout the chapters of this dissertation. It also provides descriptive statistics about some of the main variables that will be used. Before doing so, it provides a brief explanation of the case selection.

## Why Canada?

The data used for this dissertation comes from Canada, the country I live in and know best. As @noel2014 puts it, studying one's own country produces social scientific knowledge that is directly relevant to and usable by citizens of that country, and it therefore has a uniquely important character. Moreover, for the current study, Canada is a particularly interesting country to study, given that it is often classified by reports as one of the best for women to live in [@conant2019; @em22020; @usnews2020], even though only 30.7% of elected MPs are women, the 61st highest percentage among the world's 193 countries [@ipu2023]. Canadian Election Study (CES) data since 1997 shows that the gender gap in political interest has remained fairly stable despite a recent increase in the percentage of women politicians at the national and provincial levels [@sevi2021]. These characteristics make Canada a country worth studying to better understand the underlying reasons behind the stability of the gender gap in the aggregate measure of political interest, which mainly taps into interest in partisan politics. Studies that measure interest in different political topics have been conducted in Europe and the Middle East, but it is unclear how the gender differences they do find for many topics apply in Canada. Studying the transmission of political interest to children in Canada seems like a promising place to look for explanations of these gender differences that may, or may not, apply elsewhere.

## Data Collection

Within Canada, this dissertation relies on two main datasets which have been collected for this dissertation. However, three other large-scale surveys are mobilized to provide extra context and better estimates of political engagement in Canada --- and its evolution as Canadians age.

The two datasets which were collected for this dissertation share similar questions about political interest and about interest in five specific political topics: health care, international affairs, law and crime, education, and partisan politics. These five topics are those for which past studies have reported the largest gender differences in interest: women typically report being more interested in the political aspects of health care and education, while men typically report being more interested in the political aspects of the other three topics mentioned [@campbellrosie2008; @coffe2013; @ferrin2020; @hayes1993; @kuhn2004; @sabella2004; @verba1997].

Given the abstract nature of political interest, the observation of children's behaviour would not provide answer to this dissertation's aims. Political interest is different from frequency of political discussions, and different from political participation. Surveys, compared with interviews or focus groups, make it easier to get reliable interest scores that can be compared across children of different schools. This dissertation therefore relies on survey data, of which student data is the main component.

### Canadian Children Political Interest Survey (CCPIS)

To create the Canadian Children Political Interest Survey (CCPIS), data was collected among 698 Canadian children and adolescents aged 8 to 18 from seven elementary and secondary schools and one school board's Student Senate, all located in various urban areas of Quebec and Ontario,^[All are considered urban since they are part of a census metropolitan area (CMA) as defined by @statcan2023.] between August 2022 and January 2023. It seems worthwhile to study political socialization in the period of life where political interest is developed to better understand gender differences since there seems to be some level of path dependency in individuals' political interest afterwards [@prior2010; @prior2019].

Students had to fill a 15-minute online survey questionnaire during classroom time. The survey was hosted on the Qualtrics Web platform and was available in both French and English. @tbl-descriptive shows the number of classrooms and students in each of the eight schools included in the final sample of the CCPIS, as well as additional information about the schools. A total of 75 school boards,^[This figure also includes institutions known under names such as school service centers, school divisions, school districts, school councils, and centers for education.] 47 public schools, and 83 private schools across Canada were originally contacted to take part in the project. Of the 75 school boards that were contacted, 10 accepted to be part of the study (13%), 15 refused (20%), 27 did not reply to emails (36%), and 23 did not follow through or asked for information that was not possible to provide (31%). Several school boards required extensive documentation about the study or background police checks, even for virtual data collection, and most of them refused to take part of the study even after that documentation was provided to them, often citing a lack of time. Of the 132 schools that were contacted, 10 accepted (8%), 9 refused (7%), 2 were deemed ineligible after checking (2%), 88 did not respond (67%) --- in most cases after several attempts --- and 21 stopped replying after a few email exchanges (16%). Among the schools that accepted to take part in the study, 13 teachers made their students fill the survey (59% of all teachers contacted).^[Nonresponse and rejection rates after the first contact are not reported in any of the cited studies about political interest among children and teenagers.] Within each classroom, on average 18 students filled the survey --- a number that rises to 22 per classroom when removing school 5, whose students did not have an allocated classroom period to fill the survey and therefore did so out of their free time, a number very close to the 20 to 24 students per classroom on average in Ontario and Quebec schools [@bolduc2019].^[@dostiegoulet2009fr and @dostiegoulet2009en also reported high within-class response rates of 90% and 80% respectively.]

Table: Descriptive statistics, CCPIS data {#tbl-descriptive}

| **ID** | **Public/ Private**^[The three public bodies that accepted to be part of the study are associated with three different school boards.] | **Language** | **Province** | **Ages**^[Age groups are for the school itself, not the classrooms selected.] |  **Number of students in body**  | **Number of students in sample** | **Number of classrooms in sample**  | **Number of teachers in sample** |
|----|-----------------|-------------------|-----------------|-----------|------------------------|------------------------|------------------------|------------------------|
| 1           | Private          | French            | Quebec            | 12--17        | 450               | 133                | 5                   | 2                  |
| 2           | Public           | French            | Quebec            | 12--17        | 690               | 196                | 10                  | 2                  |
| 3           | Private          | French            | Quebec            | 12--17        | 670               | 78                 | 3                   | 1                  |
| 4           | Private          | French            | Quebec            | 12--17        | 900               | 253                | 12                  | 3                  |
| 5^[Mixed on-site/online school.]           | Private          | English           | Ontario           | 14--18        |                   | 5                  | 3                   | 2                  |
| 6           | Public           | French            | Quebec            | 5--12         |                   | 14                 | 1                   | 1                  |
| 7           | Private          | English           | Ontario           | 5--14         |                   | 4                  | 3                   | 1                  |
| 8^[Student Senate, a body located at the school board level.]           | Public           | English           | Ontario           | 14--18        | 15                | 15                 | 1                   | 1                  |
|             |                  |                   |                   |               | **Total**    | **698**       | **38**         | **13**        |
<!--School ID 1: Collège Boisbriand 2016, nombre d'élèves https://cdn-contenu.quebec.ca/cdn-contenu/adm/min/education/publications-adm/education/organismes-lies/CCEP/CCEP-rapport-annuel-20-21.pdf
School ID 2: École Jean-de-Brébeuf, nombre d'élèves https://cssc.gouv.qc.ca/ecole/ecole-jean-de-brebeuf
School ID 3: Collège mariste de Québec, nombre d'élèves https://www.xn--collgemariste-zgb.ca/historique
School ID 4: Collège Citoyen, nombre d'élèves https://ecolespriveesquebec.ca/ecole/college-citoyen/
School ID 5: Jaya International High School
School ID 6: École de la Rose-des-Vents
School ID 7: Urban Village Academy
School ID 8: Renfrew County DSB Student Senate, number of students https://www.rcdsb.on.ca/en/parents-and-students/student-senate.aspx-->

Those low response rates imply that the sample's descriptive characteristics should not be seen as representative of the Canadian student population. This sample should be considered a non-random, convenience sample. There are no reasons to believe that the types of correlations studied here should be affected by that however. Nevertheless, a census of students does not exist, and due to the scarcity of available data on school-aged children's political predispositions within Canada, this dataset provides relevant information about students' interests, role models, and relation to politics. The core of the data analysis in each chapter therefore relies on this dataset.

The CCPIS includes information about students' interest in various political topics, the political topics they say they discuss with their mothers, fathers and peers, as well as the political topics discussed by a teacher and an influencer they like. The gender of these role models is also inquired, to test the implication from social learning theory that political interest transmission should occur mainly between a child and role models of their gender, more than role models of other genders [@prior2019; @shehata2019].

The CCPIS also uses the Personal Attributes Questionnaire [@spence1978; @ward2006] to assess students' degrees of agency and communality, and to assess whether interest correlations between children and their same-gender role models change when controlling for the child's degree of agency or communality. Men are typically more assertive, which brings them closer to partisan politics, while women are typical more co-operative, which leads them to be more interested in topics such as health care and education. Given the importance of the agency/communality distinction in explaining gender differences in political interest [@campbellrosie2008], this hypothesis is tested alongside hypotheses about the role of parents, peers and other role models.

Finally, the CCPIS asks socio-demographic questions about children's gender, language spoken at home, immigrant status, age, and race. These questions are mostly inspired by the Canadian Election Study [@ces2019], except for the question on race, which is based on the 2016 Census [@census2017]. The English and French CCPIS questionnaires are available in Appendices I and II.

The following questions are asked about students' interests: "How interested are you in politics generally? Set the slider to a number from 0 to 10, where 0 means no interest at all, and 10 means a great deal of interest", and "If you were to open a news website and see the following articles how interested would you be in reading each article? Set the slider to a number from 0 to 10, where 0 means 'Not at all interested, I would not read it,' and 10 means 'Very interested, I would most likely read it.' (a) Health care (i.e., pandemic restrictions, working conditions of nurses); (b) International affairs (i.e., diplomatic disputes between Canada and China, Ukrainian war); (c) Law and crime (i.e., police funding, sentences for violent crimes); (d) Education (i.e., university tuition, funding of public and private schools); (e) Partisan politics (i.e., federal elections, political parties)". These questions specify concrete examples of political issues related to each topic, in order to make the political aspects of these topics more salient and avoid students answering while thinking about their own health or classes they are taking. The question's phrasing is meant to be easily understood by children and teenagers.

The main socialization-related explanatory variables for children's political interest identified in the literature concern the role of four agents of socialization: parents, peers, schools, and media. In order to determine the importance of these role models' gender, the following questions were asked to children: "Which parent do you discuss most often with? (a) Mother; (b) Father; (c) Both equally"; "What is the gender of most of your friends? (a) Girls; (b) Boys; (c) About the same for both genders"; "Think about a teacher that you like(d). Is that teacher... (a) A woman; (b) A man; (c) Other (e.g. Trans, non-binary, two-spirit, gender-queer)"; "Think about someone that you like and sometimes read or watch on social media --- including YouTube. Is that person... (a) A man; (b) A woman; (c) Other (e.g. Trans, non-binary, two-spirit, gender-queer)." Questions are then asked about the topics most discussed by parents, peers, teachers and influencers among the five political topics highlighted beforehand.

### Datagotchi Post-Election Survey (Datagotchi PES)

The second dataset collected over the course of this dissertation is the Datagotchi post-election survey, which includes 2,228 Quebec adult respondents [@datagotchi2023]. This dataset was collected in February and March 2023 using a panel of respondents which agreed to be contacted by email during the 2022 Quebec general election through the Datagotchi Web app [@datagotchi2022]. For the purposes of this dissertation, the questions of interest which were asked to panel respondents inquired about their level of interest in various political topics and general socio-economic status. For these questions, the same question wording was used than among students, with the same five political topics. The Datagotchi Post-Election Survey French questionnaire is available in Appendix III.

### Canadian Election Survey (CES)

Other than these two main datasets, three often-used datasets including Canadian respondents are mobilized to give a more general empirical overlook on political interest in Canada. First, Canadian Election Study (CES) data on various aspects of political engagement is analyzed [@ces2021]. This includes data about political interest, political knowledge, internal efficacy, external efficacy, political discussion, political debating, participation in protests, participation in boycotts, participation in petitions, party membership, donations to parties, and voting. CES data allows time series analyses of the evolution of gender gaps, with some political engagement questions having been asked since 1965.

### World Values Survey (WVS)

Second, time-series World Values Survey (WVS) data is also used. These data, like CES data, make it possible to visually represent descriptive statistics by age and gender on political engagement. Moreover, the World Values Survey (WVS) Wave 7 (2017--22) includes data from Canada and 56 other countries, therefore providing the opportunity for cross-country comparisons on political interest and political engagement. Respondents were asked a different political interest question than in other surveys: "How interested would you say you are in politics? (a) Very interested; (b) Somewhat interested; (c) Not very interested; (d) Not at all interested" [@wvs2022].^[This 4-point scale is transformed into a 0--10 scale comparable with the CES's political interest scale, using the same method as @prior2010.]

### General Social Survey (GSS)

Finally, microdata from the 2013 Canadian General Social Survey (GSS), cycle 27.2 (Social Identity) are used to complement the CES and WVS data with a less non-political questionnaire which might be more representative of Canadians' overall level of interest in politics [@gss2013].

## Variable Description & Descriptive Statistics

### CCPIS

@fig-ccpis1 shows the distribution of CCPIS students by age, gender, language, race, immigrant status, degree of agency and communality, family situation, and the gender of their role models. The gender composition of the sample is roughly balanced, with 50% of girls and 47% of boys. Most respondents are aged 13 to 17, with a large proportion (41%) being exactly 16 years old. This roughly corresponds to the moment when adolescents' political interest starts increasing alongside a gender gap in which boys report being more interested in politics [@fraile2020; @janmaat2022; @prior2019]. This age distribution should make it possible to assess the influence of all types of role models at the moment when they are the most likely to be influential for students' political interest. Respondents are mainly white (57%) and Francophone (66%), but several students who speak another language than English or French at home have also filled the survey (25%). 85% of them were born in Canada. Most students live with one mother and one father (72%), with a substantial minority (18%) also having stepparents. Other family arrangements are less common.

```{r}
#| label: fig-ccpis1
#| echo: false
#| fig-cap: "CCPIS Descriptive Statistics - General"
knitr::include_graphics("_graphs/CCPISDescriptive1.pdf")
```

For agency and communality, scales are replicated from @ward2006. Factor analysis in @fig-factor1 and @fig-factor2 show that, with one exception on the agency list --- an item that was reverse-coded^[This could be a data quality issue, since this is the only question for which the agentic pole is reverse-coded among the first eight questions asked to children, and none of the communality scale questions are reverse-coded. This means some children might have paid less attention to the actual question and responded quickly. However, given the fact that the non-reverse-coded measure is also unrelated to other elements of the scale, and since children responded differently to the agency and communality scales, this seems to be a minor concern. The scale is an exact replication of the one used by @ward2006.] --- all items scale well together, with factor loadings at least medium --- above 0.3 [@shevlin2000] --- and first eigenvalues larger than the conventionally accepted value of 1 [@williams2010]. The Cronbach's alpha for the agency scale (0.63) is slightly under the 0.7 to 0.9 range suggested by scholars [@tavakol2011], while it is within that range for the communality scale (0.78), therefore meeting the standard benchmarks for valid and reliable scales established in peer-reviewed studies about measurement scales.

```{r}
#| label: fig-factor1
#| echo: false
#| fig-cap: "CCPIS Factor Analysis: Agency Scale"
knitr::include_graphics("_graphs/AgencyScale.pdf")
```

```{r}
#| label: fig-factor2
#| echo: false
#| fig-cap: "CCPIS Factor Analysis: Communality Scale"
knitr::include_graphics("_graphs/CommunalityScale.pdf")
```

@fig-ccpis2 shows the distribution of students by interest in each of the political topics. 0 is associated with the lowest level of interest and 10 with the highest. International relations, and law and crime are generally well-balanced, with students' mean interest above 5/10 (5.2--5.3). On the other hand, fewer students report high interest in health care (4.1), education (4.1), or partisan politics (3.5).

```{r}
#| label: fig-ccpis2
#| echo: false
#| fig-cap: "CCPIS Descriptive Statistics - Political Interest"
knitr::include_graphics("_graphs/CCPISDescriptive2.pdf")
```

### Datagotchi PES, CES, WVS and GSS

@fig-dg shows descriptive statistics for the Datagotchi PES, which is not a representative sample before a form of weighting is applied. The unweighted sample includes an over-representation of men (57%), white people (97%), Francophones (94%), university-educated people (71%), and people born in Canada (86%) compared to 2021 Canadian Census data for Quebec [@census2022]. 45% of respondents have a yearly household income over \$110,000. The average age is 48, with a bimodal distribution centered around 35--40 and 65. Given the mismatch between many of these characteristics --- notably gender, level of education and race --- and the Quebec population, survey weights are created and used to make the sample more representative of the Quebec population on these two aspects, as well as income, language and age. A raking procedure is used to produce these weights. After using this procedure, the population percentages match with sample percentages.

```{r}
#| label: fig-dg
#| echo: false
#| fig-cap: "Datagotchi PES Descriptive Statistics - General"
knitr::include_graphics("_graphs/DGDescriptive1.pdf")
```

```{r}
#| label: fig-gss
#| echo: false
#| fig-cap: "2013 GSS Descriptive Statistics (Wave 27 - Social Identity) - General"
knitr::include_graphics("_graphs/GSSDescriptive1.pdf")
```

@fig-dginterest and @fig-ceswvsgssinterest show the level of interest in politics among the four adult Canadian surveys, including interest in the five topics in the case of the Datagotchi PES. While adult Quebeckers (Datagotchi PES) report being more interested in each of the five topics compared with Canadian students (CCPIS), interest in law and crime (5.5/10) is lower than for other topics, followed by law and crime (6.1/10). General political interest (7.4/10) is higher on average than interest in any of the five topics. Interest in the five political topics also follows a more bell-shaped, normal distribution than what CCPIS found among Canadian youth.

GSS and WVS respondents are less slightly interested in politics (5.7/10 for both) than those to the 2021 CES (6.4/10), potentially related to the CES's more explicitly political nature. Datagotchi PES respondents score an even higher average general political interest score of 7.4/10, which might not be surprising given the fact that these respondents are former users of the Datagotchi Web app. Analyses using the Datagotchi PES throughout this dissertation can therefore be considered to be a high bar for level of interest in each of the political topics.

```{r}
#| label: fig-dginterest
#| echo: false
#| fig-cap: "Datagotchi PES Descriptive Statistics - Political Interest"
knitr::include_graphics("_graphs/DGInterest.pdf")
```

```{r}
#| label: fig-ceswvsgssinterest
#| echo: false
#| fig-cap: "CES, WVS and GSS Descriptive Statistics - Political Interest"
knitr::include_graphics("_graphs/CESWVSGSSInterest.pdf")
```

## Methods

### Multilevel Regressions and OLS

For CCPIS data, multilevel regressions with two levels are used to help disentangle classroom-level effects from individual-level effects. While students are clustered within classrooms, which are themselves nested within schools and provinces, the number of schools (8) and provinces (2) in the final sample is too low for multilevel regressions to be conducted, and only the classroom level is kept in the analyses. Simple regression models and multiple regression models are computed, with control variables for agency, communality, role models' gender, and socio-demographic variables.

Empty models are calculated to assess what percentage in the total variance in political interest --- and in each of the five topics --- is located at the classroom level. Results show that 6.2% of variance in political interest is located at the classroom level, as well as 4.6% of variance in interest in health care, 3.6% of variance in interest in international affairs, 1.4% of variance in interest in law and crime, 8.1% of variance in interest in education, and 2.6% of variance in interest in partisan politics. In all cases, the bulk of variance therefore seems to be located at another level --- presumably the individual level.

For the other four datasets, simple and multiple ordinary least squares (OLS) regression are instead performed, with the same control variables.

### Note on Social Desirability Bias

Questions about political interest can be subject to some degree of social desirability bias, which is "the difference between an individual's own intention and his/her perception of his/her peers' intention" [@chung2003, p.&nbsp;296]. However, it is not entirely clear whether girls' scores or boys' scores would be more inflated as a result of social desirability bias. Studies have found that in Canada, Australia, and the United States, *women* generally exhibit higher social desirability bias when answering survey questions [@chung2003; @cohen1998; @cohen2001]. However, studies in these same countries and more instead show that *men* are more likely to over-report having voted in the previous elections [@belli2001; @herrick2020; @stockemer2023]. For political interest, it is unclear which gender --- if any --- tends to over-report it, but disaggregating the concept in five topics might be the best approach to avoid systematic social desirability bias associated with one particular question.

## References
