---
title: "Gender and Political Interest Development: Canadian Trends from Childhood to Adulthood {#sec-chap3}"
author: "Alexandre Fortier-Chouinard"
bibliography: Thesis.bib
format: pdf
toc: true
toc-depth: 5
---
As @sec-chap1 showed, studies have long shown a gender gap in political interest among adults in Western countries, in which men report being more interested in politics than women, and one of the main explanations for that gap is socialization. This process can be influenced by four main agents: parents, peers, media and schools. Moreover, when dissecting interests by topic, women typically develop stronger interests in cooperation-focused political topics, while men typically develop stronger interests in self-assertion-focused political topics. Socialization can happen at various stages throughout the life course. It can start from an early age but does not stop completely after someone reaches adulthood. The timing of gendered political socialization processes can be puzzling. This chapter therefore seeks to provide an overview of how political interest evolves while Canadians age. Identifying the critical periods of life in which political interest can fluctuate will lay the groundwork for @sec-chap4 and @sec-chap5, which will focus their attention on explaining how various socialization agents can influence the development of political interest, especially among children and teenagers.

This chapter addresses two related questions: *Throughout the average person's life, when does political interest increase, decrease, or remain stable? How does the evolution of political interest over the life course differ between girls and boys, and later between women and men?* After exploring what studies have found about these two questions in various countries, the chapter uses a variety of Canadian and international datasets to provide context-specific answers to these two questions and break results down by age, gender, other socio-demographic characteristics, other forms of political engagement and, most importantly, interest in various political topics. This chapter hypothesizes that for both boys and girls, adolescence is the moment in time when gender differences in interest in various political topics emerge. Notably, it is argued that boys develop more interest in partisan politics, which is a contributing factor to their higher interest in running for elected office when they reach adulthood.

## Political Interest Evolution Over the Life Course

Political interest is remarkably stable over the life course, as longitudinal studies conducted in European countries have shown [@fraile2020; @neundorf2013; @prior2010; @prior2019]. This finding is robust to changes in survey question wording. However, there is an important exception to this rule: children start with a lower and less stable level of political interest than adults.^[Children also seem to discuss politics more with parents and peers as they age [@rebenstorf2004chap9].]

Before the age of 15, it is unclear if children and teenagers experience rising, falling or stable levels of interest in politics. In the United States context, @hess1967 find that children aged 7 to 14 report less and less political interest as they grow older, and @bos2022 show the same for children aged 6 to 12. The explanation for this decline is not provided, but it seems to affect mostly girls, a result attributed to internalizing gender roles and the idea that politics is a men's domain, as exemplified by girls becoming more likely as they age to draw a man when asked to draw a politician. On the other hand, more recent peer-reviewed studies by @russo2017 and @shehata2019 both find a slight increase in political interest between 13 years old and 15 years old. In the Canadian context, @dostiegoulet2009fr also finds children's political interest falls between ages 14 and 15, before increasing between 15 and 16 years old.

Around the age of 15, numerous European studies find children start experiencing an important uptick in political interest, which keeps increasing until they reach 25 years old approximately, after which it stabilizes [@neundorf2013; @prior2019; @quintelier2014; @russo2017; @shehata2019]. @russo2017 and @prior2019 suggest the rise in political interest among adolescents could be due to an increasingly clear sense of what politics is during those formative years (15--25), although it has also been found that by age 10, children have already gained an understanding of what politics means [@hess1967]. @russo2017 put it in these terms:

> \[W\]e observed a general increase in interest in politics, which is much steeper between 16 and 18 years of age than in the 13–15 age range. One interpretation of this finding lies in the ideas that --- at these ages --- adolescents obtain cognitive abilities that allow abstract thinking and reasoning \[...\] and they learn more about society and the wider world. Another possibility is that youths become more interested in political issues because they are approaching voting age, and their 'social environment (as parents and teachers) anticipate a 'life event' in becoming an enfranchised voter' \[...\] Even if our results are consistent with both these ideas, it is worth noting that there was no national election in Sweden in 2012 (when we collected data from the 18 year-olds). Hence, cognitive maturation is a more plausible explanation for the increase in political interest that we observed between 16 and 18 years of age (654).

Adolescence and early adulthood are important, as @russo2017 suggest that it "is during this period that parents, teachers, and role models in general can potentially raise youths' interest in political and societal issues" (655). Learning when exactly political interest increases during one's lifetime can help in understanding the political socialization processes at play.

How large is this increase in political interest between 15 and 25 years old? In panel data collected among British, Swiss and German respondents, @prior2019 estimates there is a 10--15 percentage point increase. Political interest keeps increasing after 25 years old, but at a slower pace, and almost entirely due to cohort effects: older cohorts of voters, especially those born in the 1940s, are particularly interested in politics. However, within each cohort of people, after reaching 25 years old, political interest remains very stable until death. Prior's [-@prior2019] findings are similar in all three countries studied.

Political interest also becomes more *stable* at the individual level over the teenage years. @prior2019 finds an increase in the stability of political interest between the ages of 11 and 20. @russo2017 also show that the stability of political interest increases drastically from 13 years old to 20 years old, after which it remains high. Using a 5-category response scale to measure political interest, they find that 21.8% of adolescents aged 13--15 changed their answer by two or more response categories over two years, compared with only 4.5% of those aged 26 to 28.

## Gender Differences in Political Interest Evolution

### Size of the Gender Gap

Among scientific studies on the gender gap in political interest among children and teenagers, a first group of studies finds that boys already report higher levels of political interest. Among children aged 7 to 14, @hess1967 find a gender gap varying between two and five percentage points, with boys being more interested in United States government and current events. @owen1988 also find that boys aged 10--13 and 14--17 in the United States are more interested in politics than girls of that age. More recently, @arens2017 find a gender gap of 7.4 points for 12-year-old Germans, which increases to 10.8 points for 15-year-olds, where boys report being more interested in politics.

However, the findings are not all consistent across studies and contexts. @dowse1971 find no statistically significant gender gap (-0.3 to +1.4 percentage point, where positive numbers are associated with boys) in political interest for children aged 11--17. @mayer2004 also find that no strong gender differences in political interest exist between boys and girls aged 12 to 15 in China, Mexico, the United States, and Japan. Similar results are found among Quebec students aged 14 and 15 [@beauregard2008; @dostiegoulet2009fr]. However, @bos2022 find that *girls* are slightly but significantly (3.3 points) more interested in politics at 6--7 years old, but then the gap quickly reverses and grows larger until early adolescence.^[Political interest is measured using age-appropriate questions adapted from the Noyce Enthusiasm for Science scale. "Interest in political activities is an index of agree/disagree responses to the following sentiments: (1) politics, government, and history are exciting topics; (2) curiosity to learn about politics, government, and history; (3) desire to have a political job; and (4) learning about government is boring [reverse coded]" [@bos2022, 488].] By age 12, *boys* are 6.7 points more interested in politics. Overall, US boys are found to be 2.4 points more interested in politics than girls [@bos2020; @bos2022].

For older teenagers and young adults, starting at age 15 --- the time when political interest starts increasing markedly --- the literature is clearer: studies conducted in various contexts generally show an important gender gap in self-reported political interest, where men are the most interested, and the gap is growing through time for those who report longitudinal data. Using a -100--+100 scale where positive numbers are associated with men, the gender gap in political interest has been measured at +2 (@koskimaa2015, Finland, 16--18 years old and @dostiegoulet2009fr, Canada, 14--16 years old); +5 (@janmaat2022, UK, 16 years old); +11 (@cicognani2012, Belgium, 15--19 years old and @lawless2013, USA, 18--25 years old); +15 (@burns2001, United States, 18 years old and @muxel2002, France, 18--25 years old); +20 (@fraile2020, UK, 15 years old); +22 (@janmaat2022, UK, 30 years old); +27 (@hyman1959, Germany, 15--24 years old); and +30 (@fraile2020, UK, 25 years old). Among these results, only Koskimaa and Rapeli's [-@koskimaa2015] are not statistically significant.

The increases in the gender gap between late adolescence and early adulthood reported by the recent studies of @fraile2020 and @janmaat2022 are substantially large, although both rely on the same panel dataset. @fraile2020 also test for sub-periods within the 1991--2009 time frame and find similar results.

Among the general adult population, studies also point to a greater interest in politics by men compared with women. @vandeth2000 shows a gender gap in political interest in the Netherlands, which has remained steady or increased through time. Among adults, @prior2019 finds that men are 10--15 percentage points more interested in politics than women, a wider gap than the one found among younger respondents. @sanchezvitores2019 finds a statistically significant gender gap in 13 countries. @fraile2020 suggest that after the formative years of 15 to 25 years old, "attitudes crystallize and so does the gender gap, remaining at the same size (around 30 percentage points of difference between women and men) over the life course" (89). Using 2002 European Social Survey data, they find a gender gap across 15 European countries, varying between 4 and 13 percentage points. Similarly, @kittilson2012 use 2000 World Values Survey data among 29 countries and measure gender gaps of +1 (Argentina) to +28 (India), with an average gender gap of approximately +15.

### Gaps in Interest for Certain Topics

Some studies conducted among teenagers report the types of interests girls and boys have, but do not always find significant differences there. @beauregard2008 finds no gender gap in reported interest in domestic and international politics. @burns2001 also find null results with regards to interest in community and social issues, which contrasts with the gender gap in political interest they report. Finally, @oswald1998 find that "girls are more interested than boys in topics like peace, ecology and problems of the Third World, whereas boys are more interested in governmental and international affairs than girls" (153). The authors reason that girls might not be "interested as much in the institutions of politics and in the everyday business of negotiation in government and parliament and that the single question measures mainly this sphere of front-page politics" (153). All three studies focus on a relatively limited number of topics.

Among adults, on average, studies have found that women report more interest in topics such as health care, education and gender issues, while men report more interest in foreign policy, partisan politics, and law and order [@campbellrosie2008; @coffe2013; @ferrin2020; @hayes1993; @kuhn2004; @sabella2004; @verba1997]. Topics such as taxes and local politics seem to be equally interesting to men and women.

### Political Interest Evolution in Canada

Political interest has generally been found to be higher in Canada than in other Western countries. @howe2010 finds that 59% of Canadian citizens report being very or somewhat interested in politics, a higher percentage than in most European countries. Similarly, using World Values Survey data, @gidengil2004 show that Canada ranks fourth among seventeen democratic countries when it comes to the average level of political interest.^[The timing of the World Values Survey might be a confounding factor since Canadian data was collected shortly before the Meech Lake Accord failed --- a time of intense political discussion [@howe2010].] On political interest evolution, mirroring trends in other Western countries, Canadians aged 21--29 have long been less interested in politics than those aged 50--65, and this age gap seems to be growing [@gidengil2004; @howe2010]. Political interest also varies by province: "Residents of Quebec and Saskatchewan are typically less interested (5.0) in politics in general than Canadians at large. Meanwhile, interest is highest in British Columbia (5.8), followed by Manitoba, Ontario, Newfoundland and Labrador (5.7), and Alberta (5.6)" [@gidengil2004, 24].

Again using 2000 WVS data, @kittilson2012 find that Canadians stand close to the middle of the pack in terms of countries' gender gap in political interest, with a 13-point gap in which men report being more interested. For their part, @gidengil2004 evaluate the gender gap in political interest between Canadian women and men to be worth 5 percentage points, relying on older 1990 WVS data.

## Theorizing Change Over Time

It seems worthwhile to study political socialization in the period of life where political interest is developed --- childhood to early adulthood --- since there seems to be some level of path dependency in individuals' political interest afterwards. Gender differences also seem to become starker at that moment: the early increase in self-reported political interest seems to be stronger for men [@jennings1981, 276], and political socialization seems to be faster-paced during the teenage years. @fraile2020 suggest that "the development of gender roles during early childhood is a crucial phase in the source of the gender gap, deserving further attention from scholars" (89). Moreover, no study has measured the evolution of interest in various political topics with age among children or adults.

***Hypothesis 0a**: Interest in specific political topics starts rising around age 15 and increases until age 25, after which it stabilizes.*

***Hypothesis 0b**: Gender differences in interest in specific political topics start rising around age 15 and increase until age 25, after which they stabilize.*

Given findings among adults that women and men are interested in different topics depending on these topics being assertion-focused or cooperation-focused (see @sec-chap1), Hypotheses 0c and 0d provide two alternative views about when these differences should start to occur. Hypothesis 0c suggests such differences in interest exist prior to the increase in self-reported political interest around age 15, perhaps as a result of tendencies towards agency and communality which start developing early.^[Gender differences in assertive speech use may develop between 1 and 2 years old, as @fagot1985 suggest, or may not have already developed at that age, as @brownell2006 suggest. A meta-analysis of studies done by @leaper2004 including children from various age groups more generally finds that boys use more assertive speech than girls, although this is not the case in mixed-gender interactions. @noakes2006 also find that, among boys and girls aged 9--14, boys reported having more interpersonal disagreements related to status and hierarchy, and girls were more likely to use cooperative conflict resolution strategies. Finally, @caravita2012 find that, among 9--15-year-old girls and boys, boys appear to have more agentic goals and girls appear to have more communal goals, although aging makes the gender gap in agency appear and the gap in communality disappear, which could be due to a relatively small sample size.] Hypothesis 0d instead suggests that --- likely as a result of cognitive maturation --- personality traits only start fostering interest in assertion-focused or cooperation-focused political topics between ages 15 and 25.

***Hypothesis 0c**: Gender differences in interest in specific political topics already exist prior to age 15, with boys already more interested in assertion-focused political topics such as law and crime, international affairs and partisan politics, and girls more interested in cooperation-focused political topics such as health care and education politics at 10--15 years old.*

***Hypothesis 0d**: Between ages 15 and 25, boys develop more interest in assertion-focused political topics, while girls develop more interest in cooperation-focused political topics. These differences then carry on at the adult age.*

Hypothesis 0a, 0b and 0c should apply to interest in the five topics used throughout this book: health care, education, partisan politics, law and crime, and international affairs. Since past studies have found the self-reported measure of political interest is mostly associated with interest in partisan politics [@campbellrosie2008], it should be treated as an assertion-focused political topic.

Hypothesis 0e suggests, again as a result of the cognitive maturation theory laid out by @russo2017, that issues related to partisan politics --- such as elections and parties --- are more political than other not explicitly partisan issues such as the working conditions of nurses or tuition fees. This would be congruent with Hypothesis 0d.

***Hypothesis 0e**: Boys and girls both see issues related to partisan politics as more political than other political issues starting at age 15.*

Studies have found strong causal or correlational associations between political interest, political knowledge, political efficacy, and other indicators of political engagement [@bennett1989; @coffe2010; @coffe2013; @ondercin2011; @prior2019] --- who are sometimes bundled together to create general political engagement scales. @prior2019 notably finds modest positive correlations between political interest and a measure of political efficacy in three countries, and positive bidirectional causal links between efficacy and interest in one of these countries --- Switzerland. Therefore, Hypotheses 0f and 0g suggest that the value of these indicators tends to covary with age and that gender differences in these indicators should also emerge within similar time frames.

***Hypothesis 0f**: Various indicators of political engagement, such as political interest, political knowledge and political efficacy, mostly increase at the same time --- under age 25.*

***Hypothesis 0g**: Gender differences in various indicators of political engagement, such as political interest, political knowledge and political efficacy, when they exist, mostly arise at the same time --- under age 25.*

## Data and Methods

This chapter wants to portray the evolution of political interest by age and through time, and therefore relies on five datasets across several survey years: the Canadian Children Political Interest Survey (CCPIS), the Datagotchi Post-Election Survey (Datagotchi PES), the Canadian Election Study (CES), the World Values Survey (WVS) and the Canadian General Social Survey (GSS). The CCPIS includes survey data collected among 9-to-18--year-olds in school settings in 2022, while the Datagotchi PES includes data collected among Quebec adults in 2023. More about these datasets is explained in @sec-chap2. Both the CCPIS and Datagotchi PES include questions about interests in specific political topics, which can better measure the concept. These datasets are therefore used to assess the relationship between gender and interest in five topics: health care, international affairs, law and crime, education, and partisan politics. OLS and WLS regressions are used with Datagotchi PES data, while multilevel regressions with classroom fixed effects are used for CCPIS data. While the CES, WVS and GSS do not ask these specific political topics, they each have unique advantages. The CES includes the most survey years --- with a question on political interest being asked in surveys from 1997 to 2021 --- making it possible to measure other indicators of political engagement than political interest, including political knowledge, efficacy, and participation. The WVS makes it possible to compare Canadians' level of interest with other countries. Finally, two GSS surveys --- 2013 and 2020 --- provide the largest datasets and, since it is not designed as a political survey per survey, may provide more accurate measures of Canadians' level of interest in politics --- despite the limitations of a single-item measure.

Apart from questions about political interest, this chapter also relies on four other measures of political engagement taken from the Canadian Election Studies: internal political efficacy, external political efficacy, knowledge of political figures' names, and political participation.

*Internal political efficacy* is defined as "individuals' self-perceptions that they are capable of understanding politics and competent enough to participate in political acts such as voting", while *external political efficacy* is individuals' belief that the public can influence political outcomes because government leaders and institutions are responsive to their needs (@craig1982, 86). Internal efficacy is measured by the degree of agreement with the following statement: "Sometimes, politics and government seem so complicated that a person like me can't really understand what's going on." [@ces2021]. External efficacy is measured by the degree of agreement with the following statement: "People like me don't have any say about what the government does." [@ces2021]. Both questions are reverse-coded, with more agreement associated with higher political efficacy.

Knowledge of political figures' names is measured by correct answers to questions about who the minister of Finance, governor general of Canada, and their provincial premier are. Political participation is measured by a scale including 13 items: volunteering for groups and organizations, attending political meetings and speeches, attending protests, boycotting products, signing petitions, following politicians on social media, volunteering for politicians, contacting elected officials, donating to candidates, donating to causes, being a group's active member, commenting political content, and discussing politics on social media. CES respondents were asked how often they had done these activities in the past 12 months.

Factor analysis for knowledge of political figures' names (@fig-factor3) and for the political participation scale (@fig-factor4) show that items generally scale well together, with factor loadings at least medium --- above 0.3 [@shevlin2000] --- for all elements of the political participation scale and two out of three of the knowledge scale, and first eigenvalues larger than the conventionally accepted value of 1 [@williams2010]. The Cronbach's alpha for the knowledge scale (0.39) is under the 0.7 to 0.9 range suggested by scholars [@tavakol2011], while it is within that range for the political participation scale (0.84), therefore meeting the standard benchmarks for valid and reliable scales established in peer-reviewed studies about measurement scales.

```{r}
#| label: fig-factor3
#| echo: false
#| fig-cap: "CCPIS Factor Analysis: Knowledge Scale"
knitr::include_graphics("_graphs/KnowScale.pdf")
```

```{r}
#| label: fig-factor4
#| echo: false
#| fig-cap: "CCPIS Factor Analysis: Political Participation Scale"
knitr::include_graphics("_graphs/ParticScale.pdf")
```

## Results

### Evolution of Political Interest Among Adults

In order to explore what level of general (self-reported) political interest Canadian men and women report, three datasets are mobilized: the 2021 CES, the WVS's Wave 7 (2017--22) and the 2020 GSS. @fig-yeargender shows the evolution in Canadians' average level of political interest by gender from 1990 to 2021. In general, CES respondents are the most interested in politics according to these data --- 6.3/10 --- which is unsurprising given these surveys are conducted during election campaigns, and ask questions of a more clearly political nature. WVS respondents are the least interested --- 4.5/10 --- with GSS respondents sitting somewhere in between --- 5.8/10. In general, political interest seems to have increased between 2000 and 2021, with an average interest of 4.9/10 in the 2000 CES and WVS and an average interest of 6.1/10 in the 2019--20 CES, WVS and GSS.

```{r}
#| label: fig-yeargender
#| echo: false
#| fig-cap: "General Political Interest by Year and Gender Among Canadian Adults, CES, WVS (Canada) and GSS \\newline \\textit{Notes}: On the \\textit{y} axis, 0 = no interest at all, and 10 = a great deal of interest. 95% confidence intervals represented by shaded areas. CES, WVS and GSS weights are applied."
knitr::include_graphics("_graphs/InterestYearGender.pdf")
```

Despite fluctuations in the level of political interest between surveys *and* within surveys taken at different times, the gender gap in self-reported political interest remains relatively constant through time, with a notable exception in the 2021 CES, where a sharp drop in women's political interest occurs while men's political interest remains largely stable, bringing the largest gender gap of these studies --- 1.3/10 points. The size of the gender gap slightly varies between studies, standing on average at 0.6 points in the GSS, 0.7 points in the CES and 0.9 points in the WVS. While differences between the three survey organizations are twice as large as gender differences in political interest, men are significantly more likely to report being interested in politics in every survey and every year.

Again relying on data from the CES, WVS and GSS, @fig-timeceswvsgss shows average self-reported political interest by age and gender among Canadians, using a local regression model (LOESS). The trends among all three surveys are very similar and match to some degree with cross-country WVS results as well.

```{r}
#| label: fig-timeceswvsgss
#| echo: false
#| fig-cap: "Self-Reported Level of General Political Interest by Age Among Canadian Adults, 2021 CES, WVS Wave 7 and 2020 Canadian GSS \\newline \\textit{Notes}: On the \\textit{y} axis, 0 = no interest at all, and 10 = a great deal of interest. Dots represent average interest by age and gender. 95% confidence intervals represented by shaded areas. CES, WVS and GSS weights are applied."
knitr::include_graphics("_graphs/TimeCESWVSGSS.pdf")
```

In all three datasets, interest in politics increases as people age, but it does so relatively slowly from 18 to 50 years old --- or not at all, according to CES data. Findings from the GSS seem partially consistent with Hypothesis 0a, with a slightly larger increase in interest between ages 15 and 24 than afterwards, but CES data and WVS data from Canada are both inconsistent with Hypothesis 0a, with the WVS even showing a decrease in political interest between ages 18 and 25.

A large gender gap also appears for those within that age group in all three datasets. For instance, CES data shows that on average, women aged 18 to 50 report being neither interested nor disinterested in politics --- 5/10; men, on the other hand, report being somewhat interested in politics --- 6.5/10. This 1.5-point gap is statistically significant (*p*<0.001), larger than the 1-point increase in political interest between ages 50 and 75 and almost as large as the 1.8-point gap in average political interest between CES respondents and Canadian WVS respondents over the years. The gender gap seems to increase mostly between ages 15 and 25 in the GSS and WVS data, corroborating Hypothesis 0b, although CES data rather shows stability within that age range.

After age 50, both men and women start reporting higher levels of interest in politics, and this interest keeps increasing through their 60s, 70s, 80s and 90s. Moreover, again starting at age 50, the gender gap progressively reduces, as women's political interest increases more quickly than men's. Around age 75--80, the gap becomes statistically non-significant, although confidence intervals also become wider due to smaller sample sizes. For instance, in their early 90s, CES data shows that the average interest in politics stands at 8/10 for both men and women. Overall, throughout people's life course, CES data shows that women's political interest averages 5.4, while men's averages 6.8 (*p*<0.001).

How does Canada compare with other countries? Across all 57 countries surveyed during wave 7 of the WVS, the average political interest is 4.8/10 for men and 4/10 for women (*p*<0.001), and while political interest increases with people's age, the size of the gender gap remains relatively stable. Among Canadian WVS respondents, the average political interest is 6.2/10 for men and 5/10 for women: this gender gap is 0.35 points *larger* than the WVS average and is statistically significant (*p*<0.001). Moreover, the age pattern is somewhat different in the Canadian WVS compared with the WVS's cross-country results, becoming statistically insignificant around age 75 while a large and significant gap remains in other countries. Despite smaller sample sizes for people aged 75 and over in all Canadian surveys, all of them show a shrinking of the gap at that age in Canada, which reinforces confidence in that finding.

Within Canada, across all three studies, average political interest is higher in larger provinces, with an average of 6.1/10 in Alberta, 6/10 in Quebec and Ontario, and 5.9/10 in British Columbia. New Brunswick stands last at 5.5/10. Notably, the gap between the province with the highest and lowest scores is only 0.6 points, smaller than the gender gap in any of the three survey organizations over time. However, between-study differences in provinces' average political interest are large, suggesting these results need to be interpreted with caution.

@fig-cesgapyearage uses a LOESS model to look at the size of the gender gap by age for each CES, making it possible to assess the presence of generational effects. Positive values indicate women report being more interested, while negative values indicate men report being more interested. The 2021 CES stands out as having the largest gender gap in political interest on average for respondents aged 20 to 80. All other Canadian Election Studies are associated with similar trends in the evolution of the gender gap with age: the gap slowly reduces with time. Values tend to be less consistent among both older and younger voters, as a result of smaller sample sizes at the extremes of the age distribution. When respondents aged 18--20 and those aged 85 and over are removed from the sample, differences between surveys become smaller.

```{r}
#| label: fig-cesgapyearage
#| echo: false
#| fig-cap: "General Political Interest by Year and Gender Among Canadian Adults, CES \\newline \\textit{Notes}: On the \\textit{y} axis, 0 = no gender difference in interest at all, positive values (up to +10) = women more interested, negative values (down to -10) = men more interested. CES weights are applied."
knitr::include_graphics("_graphs/CESGapYearAge.pdf")
```

@fig-timepoliticalengagement shows the level of political engagement among 2021 CES respondents. Levels of external political efficacy and political participation remain low and stable for respondents aged 18 to 90, and no statistically significant gender gap is found. Conversely, large 2-point gender gaps in internal political efficacy and knowledge of political figures' names are found among young respondents. These gaps become smaller, as respondents' age increases.

```{r}
#| label: fig-timepoliticalengagement
#| echo: false
#| fig-cap: "Level of Political Engagement Across Several Measures Among Canadian Adults, 2021 CES \\newline \\textit{Notes}: On the \\textit{y} axis, 0 = no engagement at all, and 10 = a great deal of engagement. Dots represent average interest by age and gender. 95% confidence intervals represented by shaded areas. CES weights are applied."
knitr::include_graphics("_graphs/TimePoliticalEngagement.pdf")
```

Some of these trends mirror quite closely what is found for political interest among these same CES respondents. Notably, large gender gaps in knowledge of political figures' names and internal political efficacy appear among younger respondents. These gaps are larger than the 13-point gap in political interest. Similar to political interest, these gaps also seem to progressively fade among older respondents, and after age 80--85, the level of political engagement seems to be the same for women and men across all indicators. With regards to knowledge, the gap was to be expected and mirrors findings from past research --- notably @stolle2010 --- as recalling the names of political figures is closely tied to partisan politics and assertion as opposed to cooperation. Unfortunately, the data does not make it possible to assess respondents' knowledge about health care or education policy, but @stolle2010 suggest that the gender gap should be expected to be in the opposite direction, with women more knowledgeable about these topics on average.

Hypothesis 0f suggested that the various indicators of political engagement were likely to vary simultaneously, with a notable increase between ages 18 and 25. @fig-timepoliticalengagement does not corroborate this claim: internal political efficacy increases at a relatively even pace until age 40, the knowledge scale also increases constantly until age 75, while external efficacy and political participation both slightly *decline* between ages 18 and 25. These findings, while inconsistent with Hypothesis 0f, are consistent with the finding that political interest is also stable, not increasing, until age 25 --- in the CES at least. However, these results do not indicate that various indicators of political engagement are unrelated to each other. Indeed, correlations of political interest with internal political efficacy (0.42), external political efficacy (0.15), knowledge of political figures' names (0.34) and political participation (0.39) are all positive and significant at the 99.9% confidence level. Correlations between political efficacy and interest are somewhat higher than the correlations of 0.09 to 0.27 found by @prior2019 in three countries while aggregating both internal and external dimensions. Correlations between political participation and interest are also higher than the correlations of 0.07 to 0.31 found by @grechyna2023 in the EU and the UK.

Hypothesis 0g is also not corroborated by the data: the gender gap in internal political efficacy is stable, not increasing, among 18-to-25--year-olds, while for knowledge of political figures' names, the gap starts shrinking within that age group. External political efficacy and political participation do not display any notable gender gap for that age group or other age groups.

Is the gender gap in political interest concentrated among certain demographic groups only? @fig-interestwavegroup provides a robustness check to the existence of a gender gap in self-reported political interest by looking at gender differences through an intersectional lens, with gender--ethnicity pairs and gender--immigrant status pairs included across several surveys and years for which the information was available. Within each survey--year, self-reported political interest is higher among white men than white women, and non-white men than non-white women. The same applies to the gender gap among immigrants and non-immigrants. By comparison, differences in interest between Caucasians and non-Caucasians are found to be smaller, and there are no differences in interest between people born in Canada and those born abroad. The gender gap in interest therefore seems to trump other socio-demographic characteristics such as ethnicity and immigrant status, which reinforces the need to understand how gendered socialization shapes this gap.^[With regard to immigration, these results are consistent with the findings of previous studies, which found no significant difference between levels of political interest of immigrants and non-immigrants [@hochman2019]. @prior2018 found higher levels of interest in politics among white than black Americans. In the Dominican Republic, @spierings2012 found that Protestant women report levels of political interest similar to men and higher than Catholic women, while Mestizo women report lower levels of political interest than Indigenous women.]

```{r}
#| label: fig-interestwavegroup
#| echo: false
#| fig-cap: "General Political Interest by Year, Gender, Ethnicity and Immigrant Status Among Canadian Adults, CES, WVS (Canada) and GSS \\newline \\textit{Notes}: On the \\textit{y} axis, 0 = no interest at all, and 10 = a great deal of interest. 95% confidence intervals shown. CES, WVS and GSS weights are applied. Not all surveys included here asked questions about both ethnicity and immigrant status."
knitr::include_graphics("_graphs/InterestWaveGroup.pdf")
```

Now, how do adults' interest in various political topics change depending on age and gender? This question is best answered using Datagotchi PES data. As Table \ref{tab:olsInterestDG} shows, in models without controls, Quebec men generally report being more interested in politics than women, with a 0.9-point gap, similar to the other three studies. Men's interest in international affairs and partisan politics is also higher than women's (0.7-point and 0.6-point difference respectively; both *p*<0.001). However, women are significantly more interested in health care and education (0.7-point and 0.5-point difference respectively, both *p*<0.001) than men are. Interest in law and crime is about the same for both genders. Interestingly, the largest coefficient found is for interest in politics in general --- not any of specific topic.^[@fig-interestdg in @sec-appendix4 shows the gender coefficient for an intermediate model with SES (socio-economic status) controls but no interactions. Before adding interactions between (i) gender and age, (ii) gender and ethnicity, and (iii) age squared, coefficients from the simple regression model remain significant.]

\begin{table}
\centering\centering
\caption{Interest in Topic by Gender, Datagotchi PES \label{tab:olsInterestDG}}
\centering
\fontsize{6}{8}\selectfont
\begin{tabular}[t]{lcccccc}
\toprule
  & Politics (general) & Health care & International affairs & Law and crime & Education & Partisan politics\\
\midrule
\addlinespace[0.5em]
\multicolumn{7}{l}{\textit{Without Controls}}\\
\midrule \hspace{1em}(Intercept) & 7.724*** & 6.449*** & 7.453*** & 5.421*** & 6.586*** & 6.334***\\
\hspace{1em} & (0.064) & (0.075) & (0.071) & (0.079) & (0.075) & (0.084)\\
\hspace{1em}Gender (1 = women) & -0.877*** & 0.690*** & -0.592*** & 0.104 & 0.517*** & -0.645***\\
\hspace{1em} & (0.097) & (0.116) & (0.109) & (0.121) & (0.116) & (0.128)\\
\hspace{1em}Num.Obs. & 1575 & 1575 & 1575 & 1575 & 1575 & \vphantom{1} 1575\\
\hspace{1em}R2 & 0.049 & 0.022 & 0.018 & 0.000 & 0.013 & 0.016\\
\hspace{1em}R2 Adj. & 0.048 & 0.022 & 0.018 & 0.000 & 0.012 & 0.015\\
\hspace{1em}Log.Lik. & -3254.524 & -3523.301 & -3437.633 & -3591.294 & -3524.218 & -3687.310\\
\addlinespace[0.5em]
\multicolumn{7}{l}{\textit{With Controls}}\\
\midrule \hspace{1em}(Intercept) & 7.751*** & 4.166*** & 7.652*** & 6.469*** & 5.947*** & 8.463***\\
\hspace{1em} & (0.819) & (0.815) & (0.834) & (1.020) & (0.841) & (0.941)\\
\hspace{1em}Gender (1 = women) & -0.580 & 1.733* & -2.020* & -0.251 & -1.744* & -3.951***\\
\hspace{1em} & (0.694) & (0.836) & (0.856) & (0.865) & (0.864) & (0.966)\\
\hspace{1em}Age & -0.030 & -0.005 & -0.042+ & 0.013 & -0.055* & -0.144***\\
\hspace{1em} & (0.020) & (0.023) & (0.024) & (0.025) & (0.024) & (0.027)\\
\hspace{1em}Age squared & 0.000+ & 0.000 & 0.001* & 0.000 & 0.001* & 0.001***\\
\hspace{1em} & (0.000) & (0.000) & (0.000) & (0.000) & (0.000) & (0.000)\\
\hspace{1em}Ethnicity (1 = white) & 0.503 & -0.957* & -0.833+ & -1.305* & -0.722 & 0.056\\
\hspace{1em} & (0.504) & (0.477) & (0.488) & (0.628) & (0.492) & (0.551)\\
\hspace{1em}Immigrant & -0.443+ & 1.361*** & 0.057 & -0.428 & 1.035*** & 0.045\\
\hspace{1em} & (0.229) & (0.298) & (0.305) & (0.285) & (0.308) & (0.344)\\
\hspace{1em}French spoken at home & 0.106 & 0.659** & 0.324 & -0.243 & 0.307 & 0.237\\
\hspace{1em} & (0.340) & (0.224) & (0.229) & (0.423) & (0.231) & (0.259)\\
\hspace{1em}Income between \$60,000 and \$150,000 & 0.040 & 0.371+ & 0.182 & 0.257 & 0.360+ & 0.987***\\
\hspace{1em} & (0.125) & (0.190) & (0.194) & (0.156) & (0.196) & (0.219)\\
\hspace{1em}Income above \$150,000 & 0.122 & 0.172 & 0.433* & 0.439* & 0.395* & 0.622**\\
\hspace{1em} & (0.163) & (0.182) & (0.186) & (0.203) & (0.188) & (0.210)\\
\hspace{1em}Education: college & 0.118 & 0.335+ & 0.273 & 0.180 & 0.472** & 0.254\\
\hspace{1em} & (0.206) & (0.175) & (0.179) & (0.256) & (0.181) & (0.203)\\
\hspace{1em}Education: university & 0.466* & 0.735*** & 0.529* & -0.124 & 1.396*** & 0.802***\\
\hspace{1em} & (0.192) & (0.204) & (0.209) & (0.239) & (0.211) & (0.236)\\
\hspace{1em}Gender (1 = women):Age & 0.018** & 0.021** & 0.020** & -0.009 & 0.017* & 0.039***\\
\hspace{1em} & (0.006) & (0.007) & (0.007) & (0.007) & (0.007) & (0.008)\\
\hspace{1em}Gender (1 = women):Ethnicity (1 = white) & -1.181+ & -1.872* & 0.775 & 0.863 & 1.764* & 1.551+\\
\hspace{1em} & (0.657) & (0.802) & (0.821) & (0.819) & (0.828) & (0.926)\\
\hspace{1em}Num.Obs. & 1575 & 1575 & 1575 & 1575 & 1575 & 1575\\
\hspace{1em}R2 & 0.076 & 0.129 & 0.053 & 0.017 & 0.128 & 0.065\\
\hspace{1em}R2 Adj. & 0.069 & 0.122 & 0.046 & 0.010 & 0.121 & 0.057\\
\hspace{1em}Log.Lik. & -3232.057 & -39061.721 & -39098.369 & -3578.081 & -39112.301 & -39288.383\\
\bottomrule
\multicolumn{7}{l}{\rule{0pt}{1em}+ p $<$ 0.1, * p $<$ 0.05, ** p $<$ 0.01, *** p $<$ 0.001}\\
\multicolumn{7}{l}{\rule{0pt}{1em}Without controls: Ordinary least squares (OLS) regressions}\\
\multicolumn{7}{l}{\rule{0pt}{1em}With controls: OLS for Politics (general) and Law and Crime; Weighted least squares (WLS) for other regressions}\\
\end{tabular}
\end{table}

In models with controls for socio-demographic variables and interaction terms for gender and ethnicity as well as gender and age, Table \ref{tab:olsInterestDG} shows some of the relationships found in models without controls remain significant while others lose significance. Self-reported political interest is not associated with being a man, but it is positively associated with being an older woman and marginally negatively associated with being a white woman. Women remain more likely to be interested in health care politics under these specifications, and again, there is a positive association with being an older woman and a negative association with being a white woman. International affairs are positively associated with being a man in general, but there is again a positive association with being an older woman. Interest in law and crime remains unlinked with gender, while interest in partisan politics is again strongly associated with being a man --- also with a significant impact of being an older woman. The coefficients for interest in education politics are notable, as being a white woman and being an older woman are both positively related to the outcome, but the coefficient for gender is now negative, suggesting that men are expected to be more interested in education politics after controlling for interactions between gender, ethnicity and age.

@fig-datagotchi inquires further into age trends in the evolution of interest in each of the topics, including gendered aspects of these age trends. Similar to @fig-ceswvsgssinterest, the gender gap in self-reported political interest diminishes among older respondents, as older women tend to become more interested. This trend of increased interest in politics among older women compared with younger ones can also be observed in interest in health care, international affairs, law and crime, and partisan politics. After age 50, men are no more likely than women to report being interested in partisan politics or international affairs, despite a significant gender gap among younger respondents.

```{r}
#| label: fig-datagotchi
#| echo: false
#| fig-cap: "Self-Reported Level of Interest in Various Topics by Age Among Canadian Adults, 2022 Datagotchi PES \\newline \\textit{Notes}: On the \\textit{y} axis, 0 = no interest at all, and 10 = a great deal of interest. Dots represent average interest by age and gender."
knitr::include_graphics("_graphs/InterestAgeGenderDG.pdf")
```

Among Quebec men, there seems to be a general increase in the level of political interest among older respondents for general political interest, interest in health care and international affairs. Interest in law and crime seems to be mostly stable, while a downward trend is found for partisan politics. Large variations among older respondents can be due to smaller sample sizes. This decline in interest in partisan politics is difficult to explain. One possibility is that this is the result of lower levels of college attendance among men than women.

Interest in the politics of education shows a unique trajectory, with a shape that resembles a cubic equation function for both women and men. One potential explanation for this trend is that education increases quickly between ages 25--40 as women and men become parents and care about the potential effects of education policy on their children. As children grow older, parents lose interest in education policy, but after retiring around age 65, women and men often become grandparents and start caring about the education of their grandchildren. While @campbellrosie2008 find that parenthood is associated with a lower general political interest in the UK context, @grechyna2023 find no statistically significant relationship. Yet, none of these other studies test specifically for interest in education politics.

Overall, @fig-datagotchi provides little evidence to support either Hypothesis 0a, that interest in politics increases until age 25, Hypothesis 0b, that gender differences increase mostly until age 25. Interest in all topics generally increases with age among women and men, while adult men's interest in partisan politics declines with age. But most of the movement in average political interest and in gender gaps happens after age 25. Hypothesis 0d, that men develop more interest in assertion-focused topics in their young adult years while women develop in cooperation-focused topics, is also not corroborated by the data. For partisan politics and international affairs, the gender gap is concentrated among younger respondents and disappears among older ones; for health care and education, the gender gap becomes significant only among respondents aged 35 and over; finally, no gender gap is found for interest in law and crime.

### Evolution of Political Interest Among Children

What kinds of political topics are children and adolescents interested in? Are there gender differences in how interested they are in these topics? Does gender predict political interest in itself, or is it only a proxy for other variables? Data from the CCPIS is used to answer these questions. Table \ref{tab:lmeInterestCCPIS} shows the link between gender and interest in each of the topics among elementary and high school students. The upper part includes gender as the only predictor, while the lower part includes controls for socio-demographic variables and personality traits.

\begin{table}
\centering\centering
\caption{Interest in Topic by Gender, CCPIS \label{tab:lmeInterestCCPIS}}
\centering
\fontsize{6}{8}\selectfont
\begin{tabular}[t]{lcccccc}
\toprule
  & Politics (general) & Health care & International affairs & Law and crime & Education & Partisan politics\\
\midrule
\addlinespace[0.5em]
\multicolumn{7}{l}{\textit{Without Controls}}\\
\midrule \hspace{1em}(Intercept) & 4.579*** & 4.041*** & 5.724*** & 4.956*** & 4.219*** & 4.007***\\
\hspace{1em} & (0.184) & (0.167) & (0.180) & (0.173) & (0.206) & (0.171)\\
\hspace{1em}Gender (1 = girl) & -0.434* & 0.128 & -0.980*** & 0.488* & -0.103 & -0.854***\\
\hspace{1em} & (0.207) & (0.197) & (0.229) & (0.231) & (0.223) & (0.232)\\
\hspace{1em}SD (Intercept Class) & 0.651 & 0.542 & 0.473 & 0.377 & 0.771 & 0.344\\
\hspace{1em}SD (Observations) & 2.499 & 2.397 & 2.802 & 2.837 & 2.701 & 2.855\\
\hspace{1em}Num.Obs. & 617 & 623 & 620 & 619 & 623 & 620\\
\hspace{1em}R2 Marg. & 0.007 & 0.001 & 0.029 & 0.007 & 0.000 & 0.022\\
\addlinespace[0.5em]
\multicolumn{7}{l}{\textit{With Controls}}\\
\midrule \hspace{1em}(Intercept) & -2.429 & 10.710 & 13.284 & 7.365 & 19.424* & 21.801**\\
\hspace{1em} & (8.113) & (7.495) & (8.444) & (8.799) & (8.916) & (8.210)\\
\hspace{1em}Gender (1 = girl) & 1.497 & 0.074 & -2.296 & 0.903 & -0.121 & -0.640\\
\hspace{1em} & (2.046) & (1.984) & (2.271) & (2.326) & (2.233) & (2.275)\\
\hspace{1em}Age & 0.293 & -1.304 & -1.420 & -0.601 & -2.627* & -2.814*\\
\hspace{1em} & (1.111) & (1.031) & (1.159) & (1.211) & (1.221) & (1.127)\\
\hspace{1em}Age squared & -0.004 & 0.049 & 0.047 & 0.026 & 0.098* & 0.092*\\
\hspace{1em} & (0.038) & (0.035) & (0.040) & (0.042) & (0.042) & (0.039)\\
\hspace{1em}Ethnicity (1 = white) & 0.630* & 0.214 & 0.886* & -0.053 & 0.050 & 0.662+\\
\hspace{1em} & (0.320) & (0.315) & (0.354) & (0.365) & (0.352) & (0.357)\\
\hspace{1em}Immigrant & -0.163 & 0.167 & -0.713* & -0.690+ & 0.212 & -0.413\\
\hspace{1em} & (0.321) & (0.317) & (0.357) & (0.371) & (0.357) & (0.359)\\
\hspace{1em}English spoken at home & -0.445 & -0.218 & -1.644*** & -0.514 & 0.253 & -0.106\\
\hspace{1em} & (0.451) & (0.436) & (0.493) & (0.507) & (0.499) & (0.491)\\
\hspace{1em}French spoken at home & -0.035 & 0.093 & -0.358 & -0.245 & 0.422 & 0.550+\\
\hspace{1em} & (0.279) & (0.275) & (0.308) & (0.319) & (0.307) & (0.308)\\
\hspace{1em}Agency & 3.494*** & 1.319* & 2.544*** & 2.077** & 1.321+ & 3.335***\\
\hspace{1em} & (0.629) & (0.614) & (0.694) & (0.719) & (0.690) & (0.702)\\
\hspace{1em}Communality & 1.662* & 1.066 & 2.253** & -0.174 & 1.283+ & 1.243\\
\hspace{1em} & (0.658) & (0.656) & (0.739) & (0.760) & (0.733) & (0.756)\\
\hspace{1em}Gender (1 = girl):Age & -0.092 & 0.025 & 0.115 & -0.010 & 0.018 & 0.030\\
\hspace{1em} & (0.130) & (0.126) & (0.145) & (0.148) & (0.142) & (0.145)\\
\hspace{1em}Gender (1 = girl):Ethnicity (1 = white) & -0.764+ & -0.500 & -0.635 & -0.218 & -0.505 & -0.929+\\
\hspace{1em} & (0.428) & (0.420) & (0.473) & (0.490) & (0.470) & (0.477)\\
\hspace{1em}SD (Intercept Class) & 0.515 & 0.401 & 0.335 & 0.366 & 0.642 & 0.203\\
\hspace{1em}SD (Observations) & 2.385 & 2.371 & 2.686 & 2.774 & 2.636 & 2.717\\
\hspace{1em}Num.Obs. & 558 & 563 & 559 & 559 & 561 & 561\\
\hspace{1em}R2 Marg. & 0.109 & 0.041 & 0.119 & 0.039 & 0.057 & 0.109\\
\bottomrule
\multicolumn{7}{l}{\rule{0pt}{1em}+ p $<$ 0.1, * p $<$ 0.05, ** p $<$ 0.01, *** p $<$ 0.001}\\
\multicolumn{7}{l}{\rule{0pt}{1em}Method: Multilevel linear regression}\\
\multicolumn{7}{l}{\rule{0pt}{1em}Fixed Effects: Classroom}\\
\multicolumn{7}{l}{\rule{0pt}{1em}Reference Category for Language: Other languages spoken at home}\\
\end{tabular}
\end{table}

In the upper half of the table, taking into account classroom fixed effects, boys generally report being more interested in politics than girls, but the gender gap is relatively minimal, standing at 0.4 for the 11-point political interest scale (*p*<0.05). Boys' interest in international affairs and partisan politics is higher than girls' (1-point and 0.9-point difference respectively; both *p*<0.001). This seems to be in line with the results among adult respondents. However, girls' interest in law and crime is also slightly higher (0.5 point, *p*<0.05). This result is more surprising given previous literature showing the contrary. Yet, the direction of the gender gap for these three topics matches findings among Datagotchi PES adult respondents. Interest in the other two topics, health care and education, are almost even between the genders. Coupled with Datagotchi PES data showing significant gaps for both of these topics only emerge after age 35, when women become more interested than men in both cases, this suggests socialization for these two topics may happen well into the adult age rather than during the formative years, while gender patterns for interest in partisan politics and international affairs emerge early and may fade after age 50. The experiences of several women as mothers and caregivers later in life might be the elements that shape the importance they start giving to both issues.

However, when controlling for socio-demographic factors and personality traits and adding interactions for gender and age as well as gender and ethnicity, all relationships between gender and interest disappear, as the lower part of Table \ref{tab:lmeInterestCCPIS} shows. Moreover, none of the interaction terms are significant at the *p*<0.05 level. Yet, when interaction terms are removed but other control variables are kept, coefficients from the simple regression model remain significant except for general political interest, which marginally loses statistical significance.^[@fig-interestccpis in @sec-appendix4 shows the gender coefficient for two intermediate models, one with SES controls and one with SES and personality traits but no interactions.] These results may therefore be an artifact of high multicollinearity.

@fig-interestyoungold re-analyzes those results for children aged 9--15 (*n*=277) and teenagers aged 16--18 (*n*=365). Among younger students, the gender gap in self-reported political interest is 0.1/10 (not statistically significant). Among those aged 16--18, this gap grows to 0.5/10 (*p*<0.05). Notably, among both groups, both interest in international affairs and interest in partisan politics are higher among boys than girls, suggesting some gender differences in interests may already exist before adolescence.

```{r}
#| label: fig-interestyoungold
#| echo: false
#| fig-cap: "Gender Differences in Interest for Specific Political Topics by Age Group Among Canadian Children, 2022 CCPIS \\newline \\textit{Notes}: No controls are added."
knitr::include_graphics("_graphs/GenderCCPISYO.pdf")
```

Moreover, interest in each of the five topics is higher among children aged 16--18 than those aged 10--15, confirming insights from past literature. Gender gaps amount to 0.8 points for general political interest (*p*<0.001), 0.8 points for health care (*p*<0.001), 0.6 points for international affairs (*p*<0.05), 0.7 points for law and crime (*p*<0.01), 1.1 points for education (*p*<0.001), and 0.2 points for partisan politics (N.S.).^[The full regression table for @fig-interestyoungold, Table \ref{tab:lmeInterestYoungOldCCPIS}, can be found in @sec-appendix4.]

These results, coupled with those among young adults, support the idea that, while a gender gap in self-reported political interest emerges between ages 15--25, gender gaps in the other five topics develop in different ways, contrary to the expectations of Hypotheses 0a, 0b and 0d. For education politics, significant fluctuations in the overall level of interest seem to be related to life-cycle effects including the birth of children and grandchildren. For health care politics, the gender gap becomes significant only around age 35, perhaps as care experiences among women have become more common. For partisan politics and international affairs, young boys already express significantly more interest in these topics than young girls, which lends partial support for Hypothesis 0c, that gender differences in assertion-focused topics would emerge at 10--15 years old or earlier. The gender gaps for these two topics then disappear around age 40--50 --- the same moment when the gender gap in general political interest starts narrowing in Canadian election studies.

@fig-political shows how students view each of the 10 concrete issues associated with the 5 topics as political or non-political depending on their age. As expected by Hypothesis 0e, the two issues related to partisan politics are almost universally seen as political, followed by issues related to international affairs. These are the two topics for which boys report being more interested. Issues related to the other three topics are perceived by 25% to 60% of students as being non-political --- substantially large proportions. @campbellrosie2008 and @ferrin2020 had similarly found that issues related to partisan politics were seen as more political than those related to other topics. Visibly, this trend also applies to Canadian children, which suggests that topics that are seen as political are also those for which men typically report a higher degree of interest.

```{r}
#| label: fig-political
#| echo: false
#| fig-cap: "Views of Topics as Political or Non-Political By Canadian Students By Age Group, 2022 CCPIS"
knitr::include_graphics("_graphs/CCPISPolitical.pdf")
```

Contrary to Hypothesis 0e, age does not seem to matter in defining which types of issues students see as political. Older students are more likely to define issues as political, but the ordering of issues as political does not change much between both groups --- partisan politics and international affairs always are at the top of the list. It seems that ideas about what politics is and is not are already well-defined in children's minds at a young age. This is particularly striking as interest in partisan politics is not significantly higher among older children than younger ones, suggesting ideas about partisan politics --- both what it is and how interesting it is --- are already ingrained in the early teens.

## Discussion

This chapter highlights gender differences in interest in various political topics, but also shows differences between children and adults. For international relations and partisan politics, male students report more interest in these topics, just like adults in previous studies. However, girls' higher interest in law and crime and similar levels of interest in health care and education compared to boys contrast with data previously found among adults [@campbellrosie2008; @coffe2013; @ferrin2020; @hayes1993; @kuhn2004; @sabella2004; @verba1997].

@russo2017 suggested that the 13--15 age range is a period when parents, teachers, and other role models have the highest potential for transmission of interest in politics to children. Applying these results to the Canadian context, it appears that the 15--25-year-old period, similar to what previous studies have found, is also the point in time where a gender gap in interest for politics more generally emerges, with boys/men reporting higher interest in politics in general than girls/women.

However, these findings do not seem to apply to any political topic individually. Given the results of the Datagotchi PES, which shows higher levels of interest in the political aspects of health care and education among adult women than men, it seems that the gender gap in interest in these topics emerges around 30--40 years old, perhaps as a result of life events that happen at that stage, such as women becoming more likely to care for children and relatives. Moreover, with the help of CCPIS data, it is possible to note, for the first time, that some gender differences in interests are not necessarily pre-existent among adolescents. On the other hand, reported interest in politics in general, partisan politics and international affairs is already higher among boys aged 10 to 18, showing for the first time that gender differences in interest for certain aspects of politics seem to emerge before the emergence of a gap in interest in politics in general. Moreover, young students already think of issues related to these two topics as being more political than other issues related to health care or education politics, highlighting the fact that political thinking may be starting early.

Overall, measuring political interest using a single-item may give the impression that political interest emerges between ages 15 and 25, alongside a gender gap where men start reporting higher interest. Yet, looking at different data sources, it seems that the evolution of interest in politics varies not only by gender, but also by topic.