---
title: "Gender and Political Interest Development: Canadian Trends from Childhood to Adulthood {#sec-chap3}"
author: "Alexandre Fortier-Chouinard"
bibliography: Thesis.bib
format: pdf
toc: true
toc-depth: 5
---
As @sec-chap1 showed, studies have long shown a gender gap in political interest among adults in Western countries, in which men report being more interested in politics than women, and one of the main explanations for that gap is socialization. This process can be influenced by four main agents: parents, peers, media and schools. Moreover, when dissecting interests by topic, women typically develop stronger interests in cooperation-focused political topics, while men typically develop stronger interests in self-assertion-focused political topics. Socialization can happen at various stages throughout the life course. It can start from an early age but does not stop completely after someone reaches adulthood. The timing of gendered political socialization processes can be puzzling. This chapter therefore seeks to provide an overview of how political interest evolves while Canadians age. By identifying the critical periods of life in which political interest can fluctuate, it will lay the groundwork for @sec-chap4, @sec-chap5 and @sec-chap6, which will focus their attention on explaining how various socialization agents can influence the development of political interest, especially among children and teenagers.

This chapter addresses two related questions: *Throughout the average person's life, when does political interest increase, decrease, or remain stable? How does the evolution of political interest over the life course differ between girls and boys, and later between women and men?* After exploring what studies have found about these two questions in various countries, the chapter uses a variety of Canadian and international datasets to provide context-specific answers to these two questions and break results down by age, gender, other socio-demographic characteristics, other forms of political engagement and, most importantly, interest in various political topics. This chapter hypothesizes that for both boys and girls, adolescence is the moment in time when gender differences in interest for various political topics emerge. Notably, it is argued that boys develop more interest in partisan politics, which is a contributing factor for their higher interest in running for elected office when they reach adulthood.

## Political Interest Evolution Over the Life Course

Political interest is remarkably stable over the life course, as longitudinal studies conducted in European countries have shown [@fraile2020; @neundorf2013; @prior2010; @prior2019]. This finding is robust to changes in survey question wording. However, there is an important exception to this rule: children start with a lower and less stable level of political interest than adults.

Before the age of 15, it is unclear if children and teenagers experience rising, falling or stable levels of interest in politics. In the United States context, @hess1967 find that children aged 7 to 14 report less and less political interest as they grow older, and @bos2022 show the same for children aged 6 to 12. The explanation for this decline is not provided, but it seems to affect mostly girls, a result attributed to internalizing gender roles and the idea that politics is a men's domain, as exemplified by girls becoming more likely as they age to draw a man when asked to draw a politician. On the other hand, more recent peer-reviewed studies by @russo2017 and @shehata2019 both find a slight increase in political interest between 13 years old and 15 years old. In the Canadian context, @dostiegoulet2009fr also finds children's political interest falls between ages 14 and 15, before increasing between 15 and 16 years old.

Around the age of 15, numerous European studies find children start experiencing an important uptick in political interest, which keeps increasing until they reach 25 years old approximately, after which it stabilizes [@neundorf2013; @prior2019; @quintelier2014; @russo2017; @shehata2019]. @russo2017 and @prior2019 suggest the rise in political interest among adolescents could be due to an increasingly clear sense of what politics is during those formative years (15--25), although it has also been found that by age 10, children have already gained an understanding of what politics means [@hess1967]. @russo2017 put it in these terms:

> \[W\]e observed a general increase in interest in politics, which is much steeper between 16 and 18 years of age than in the 13–15 age range. One interpretation of this finding lies in the ideas that --- at these ages --- adolescents obtain cognitive abilities that allow abstract thinking and reasoning \[...\] and they learn more about society and the wider world. Another possibility is that youths become more interested in political issues because they are approaching voting age, and their 'social environment (as parents and teachers) anticipate a 'life event' in becoming an enfranchised voter' \[...\] Even if our results are consistent with both these ideas, it is worth noting that there was no national election in Sweden in 2012 (when we collected data from the 18 year-olds). Hence, cognitive maturation is a more plausible explanation for the increase in political interest that we observed between 16 and 18 years of age (654).

Adolescence and early adulthood are important, as @russo2017 suggest that it "is during this period that parents, teachers, and role models in general can potentially raise youths' interest in political and societal issues" (655). Learning when exactly political interest increases during one's lifetime can help understanding the political socialization processes at play.

How large is this increase in political interest between 15 and 25 years old? In panel data collected among British, Swiss and German respondents, @prior2019 estimates there is a 10--15 percentage point increase. Political interest keeps increasing after 25 years old, but at a slower pace, and almost entirely due to cohort effects: older cohorts of voters, especially those born in the 1940s, are particularly interested in politics. However, within each cohort of people, after reaching 25 years old, political interest remains very stable until death. Prior's [-@prior2019] findings are similar in all three countries studied.

Political interest also becomes more *stable* at the individual level over the teenage years. @prior2019 finds an increase in the stability of political interest between the ages of 11 and 20. @russo2017 also show that the stability of political interest increases drastically from 13 years old to 20 years old, after which it remains high. Using a 5-category response scale to measure political interest, they find that 21.8% of adolescents aged 13--15 changed their answer by two or more response categories over two years, compared with only 4.5% of those aged 26 to 28.

## Gender Differences in Political Interest Evolution

### Size of the Gender Gap

Among scientific studies on the gender gap in political interest among children and teenagers, a first group of studies finds that boys already report higher levels of political interest. Among children aged 7 to 14, @hess1967 find a gender gap varying between two and five percentage points, with boys being more interested in United States government and current events. @owen1988 also find that boys aged 10--13 and 14--17 in the United States are more interested in politics than girls of that age. More recently, @arens2017 find a gender gap of 7.4 points for 12-year-old Germans, which increases to 10.8 points for 15-year-olds, where boys report being more interested in politics.

However, the findings are not all consistent across studies and contexts. @dowse1971 find no statistically significant gender gap (-0.3 to +1.4 percentage point, where positive numbers are associated with boys) in political interest for children aged 11--17. @mayer2004 also find that no strong gender differences in political interest exist between boys and girls aged 12 to 15 in China, Mexico, the United States, and Japan. Similar results are found among Quebec students aged 14 and 15 [@beauregard2008; @dostiegoulet2009fr]. However, @bos2022 find that *girls* are slightly but significantly (3.3 points) more interested in politics at 6--7 years old, but then the gap quickly reverses and grows larger until early adolescence.^[Political interest is measured using age-appropriate questions adapted from the Noyce Enthusiasm for Science scale. "Interest in political activities is an index of agree/disagree responses to the following sentiments: (1) politics, government, and history are exciting topics; (2) curiosity to learn about politics, government, and history; (3) desire to have a political job; and (4) learning about government is boring [reverse coded]" [@bos2022, 488].] By age 12, *boys* are 6.7 points more interested in politics. Overall, US boys are found to be 2.4 points more interested in politics than girls [@bos2020; @bos2022].

For older teenagers and young adults, starting at age 15 --- the time when political interest starts increasing markedly --- the literature is clearer: studies conducted in various contexts generally show an important gender gap in self-reported political interest, where men are the most interested, and the gap is growing through time for those who report longitudinal data. Using a -100--+100 scale where positive numbers are associated with men, the gender gap in political interest has been measured at +2 (@koskimaa2015, Finland, 16--18 years old and @dostiegoulet2009fr, Canada, 14--16 years old); +5 (@janmaat2022, UK, 16 years old); +11 (@cicognani2012, Belgium, 15--19 years old and @lawless2013, USA, 18--25 years old); +15 (@burns2001, United States, 18 years old and @muxel2002, France, 18--25 years old); +20 (@fraile2020, UK, 15 years old); +22 (@janmaat2022, UK, 30 years old); +27 (@hyman1959, Germany, 15--24 years old); and +30 (@fraile2020, UK, 25 years old). Among these results, only Koskimaa and Rapeli's [-@koskimaa2015] are not statistically significant.

The increases in the gender gap between late adolescence and early adulthood reported by the recent studies of @fraile2020 and @janmaat2022 are substantially large, although both rely on the same panel dataset. @fraile2020 also test for sub-periods within the 1991--2009 time frame and find similar results.

Among the general adult population, studies also point to a greater interest in politics by men compared with women. @vandeth2000 shows a gender gap in political interest in the Netherlands, which has remained steady or increased through time. Among adults, @prior2019 finds that men are 10--15 percentage points more interested in politics than women, a wider gap than the one found among younger respondents. @sanchezvitores2019 finds a statistically significant gender gap in 13 countries. @fraile2020 suggest that after the formative years of 15 to 25 years old, "attitudes crystallize and so does the gender gap, remaining at the same size (around 30 percentage points of difference between women and men) over the life course" (89). Using 2002 European Social Survey data, they find a gender gap across 15 European countries, varying between 4 and 13 percentage points.

### Gaps in Interest for Certain Topics

Some studies conducted among teenagers report the types of interests girls and boys have, but do not always find significant differences there. @beauregard2008 finds no gender gap in reported interest in domestic and international politics. @burns2001 also find null results with regards to interest in community and social issues, which contrasts with the gender gap in political interest they report. Finally, @oswald1998 find that "girls are more interested than boys in topics like peace, ecology and problems of the Third World, whereas boys are more interested in governmental and international affairs than girls" (153). The authors reason that girls might not be "interested as much in the institutions of politics and in the everyday business of negotiation in government and parliament and that the single question measures mainly this sphere of front-page politics" (153). All three studies focus on a relatively limited number of topics.

Among adults, on average, studies have found that women report more interest in topics such as health care, education and gender issues, while men report more interest in foreign policy, partisan politics, and law and order [@campbellrosie2008; @coffe2013; @ferrin2020; @hayes1993; @kuhn2004; @sabella2004; @verba1997]. Topics such as taxes and local politics seem to be equally interesting to men and women.

### Political Interest Evolution in Canada

Political interest has generally been found to be higher in Canada than in other Western countries. @howe2010 finds that 59% of Canadian citizens report being very or somewhat interested in politics, a higher percentage than in most European countries. Similarly, using World Values Survey data, @gidengil2004 show that Canada ranks fourth among seventeen democratic countries when it comes to the average level of political interest.^[The timing of the World Values Survey might be a confounding factor since Canadian data was collected shortly before the Meech Lake Accord failed --- a time of intense political discussion [@howe2010].] On political interest evolution, mirroring trends in other Western countries, Canadians aged 21--29 have long been less interested in politics than those aged 50--65, and this age gap seems to be growing [@gidengil2004; @howe2010]. Political interest also varies by province: "Residents of Quebec and Saskatchewan are typically less interested (5.0) in politics in general than Canadians at large. Meanwhile, interest is highest in British Columbia (5.8), followed by Manitoba, Ontario, Newfoundland and Labrador (5.7), and Alberta (5.6)" [@gidengil2004, 24].

@gidengil2004 evaluate the gender gap in political interest between Canadian women and men to be worth 5 percentage points, which is a modest gap according to them, although they emphasize this gap is persistent.

## Theorizing Change Over Time

It seems worthwhile to study political socialization in the period of life where political interest is developed --- chuldhood to early adulthood --- since there seems to be some level of path dependency in individuals' political interest afterwards. Gender differences also seem to become starker at that moment: the early increase in self-reported political interest seems to be stronger for men [@jennings1981, 276], and political socialization seems to be faster-paced during the teenage years. @fraile2020 suggest that "the development of gender roles during early childhood is a crucial phase in the source of the gender gap, deserving further attention from scholars" (89). Moreover, no study has measured the evolution of interest in various political topics with age among children or adults.

***Hypothesis 0a**: Interest in specific political topics starts rising around age 15 and increases until age 25, after which it stabilizes.*

***Hypothesis 0b**: Gender differences in interest in specific political topics start rising around age 15 and increase until age 25, after which they stabilize.*

Given findings among adults that women and men are interested in different topics depending on these topics being assertion-focused or cooperation-focused, Hypotheses 0c and 0d provide two alternative views about when these differences should start to occur. Hypothesis 0c suggests such differences in interest exist prior to the increase in self-reported political interest around age 15, perhaps as a result of tendencies towards agency and communality which start developing early.^[Gender differences in assertive speech use may develop between 1 and 2 years old, as @fagot1985 suggest, or may not have already developed at that age, as @brownell2006 suggest. A meta-analysis of studies done by @leaper2004 including children from various age groups more generally finds that boys use more assertive speech than girls, although this is not the case in mixed-gender interactions. @noakes2006 also find that, among boys and girls aged 9--14, boys reported having more interpersonal disagreements related to status and hierarchy, and girls were more likely to use cooperative conflict resolution strategies. Finally, @caravita2012 find that, among 9--15-year-old girls and boys, boys appear to have more agentic goals and girls appear to have more communal goals, although aging makes the gender gap in agency appear and the gap in communality disappear, which could be due to a relatively small sample size.] Hypothesis 0d instead suggests that --- likely as a result of cognitive maturation --- personality traits only start fostering interest in assertion-focused or cooperation-focused political topics between ages 15 and 25.

***Hypothesis 0c**: Gender differences in interest in specific political topics already exist prior to age 15, with boys already more interested in assertion-focused political topics and girls more interested in cooperation-focused political topics at 10--15 years old.*

***Hypothesis 0d**: Between ages 15 and 25, boys develop more interest in assertion-focused political topics such as law and crime, international affairs, and partisan politics, while girls develop more interest in cooperation-focused political topics such as health care and education politics. These differences then carry on at the adult age.*

Hypothesis 0a, 0b and 0c should apply to interest in the five topics used throughout this book: health care, education, partisan politics, law and crime, and international affairs. They should also apply, as shown in previous studies, to the self-reported measure of political interest, since this measure is mostly associated to interest in partisan politics --- an assertion-focused topic.

Hypothesis 0e suggests, again as a result of the cognitive maturation theory laid out by @russo2017, that issues related to partisan politics --- such as elections and parties --- are more political than other not explicitly partisan issues such as the working conditions of nurses or tuition fees. This would be congruent with Hypothesis 0d.

***Hypothesis 0e**: Boys and girls both see issues related to partisan politics as more political than other political issues starting at age 15.*

Studies have found strong causal or correlational associations between political interest, political knowledge, political efficacy, and other indicators of political engagement [@bennett1989; @coffe2010; @coffe2013; @ondercin2011; @prior2019] --- who are sometimes bundled together to create general political engagement scales. Therefore, Hypotheses 0f and 0g suggest that the value of these indicators tend to covary by age, and that gender differences in these indicators should also emerge within similar time frames.

***Hypothesis 0f**: Various indicators of political engagement, such as political interest, political knowledge and political efficacy, mostly increase at the same time --- ages 15--25.*

***Hypothesis 0g**: Gender differences in various indicators of political engagement, such as political interest, political knowledge and political efficacy, when they exist, mostly arise at the same time --- ages 15--25.*

## Data Analysis

### Pan-Canadian Surveys among Adults

In order to explore what level of general political interest Canadian men and women report, three datasets are mobilized: the 2021 Canadian Election Study (CES), the World Values Survey (WVS)'s Wave 7 and the 2020 General Social Survey (GSS). @fig-timeceswvsgss shows the rate of self-reported political interest by age and gender among Canadians, using a local regression model (LOESS). The trends among all three surveys are very similar and match to some degree with cross-country WVS results as well.

```{r}
#| label: fig-timeceswvsgss
#| echo: false
#| fig-cap: "Self-Reported Level of General Political Interest by Age Among Canadian Adults, 2021 CES, WVS Wave 7 and 2020 Canadian GSS \\newline \\textit{Notes}: On the \\textit{y} axis, 0 = no interest at all, and 10 = a great deal of interest. Dots represent average interest by age and gender. CES, WVS and GSS weights are applied."
knitr::include_graphics("_graphs/TimeCESWVSGSS.pdf")
```

In all three datasets, interest in politics increases as people age, but it does so relatively slowly between 18 to 50 years old --- or not at all, according to CES data. A large gender gap also appears for those within that age group in all three datasets. For instance, CES data shows that on average, women aged 18 to 50 report being neither interested nor disinterested in politics --- 5/10; men, on the other hand, report being somewhat interested in politics --- 6.5/10. This 1.5-point gap is statistically significant (*p*<0.001).

After age 50, both men and women start reporting higher levels of interest in politics, and this interest keeps increasing through their 60s, 70s, 80s and 90s. Moreover, again starting at age 50, the gender gap progressively reduces, as women's political interest increases more quickly than men's. Around age 75--80, the gap becomes statistically non-significant, although confidence intervals also become wider due to smaller sample sizes. For instance, in their early 90s, CES data shows that the average interest in politics stands at 8/10 for both men and women. Overall, throughout people's life course, CES data shows that women's political interest averages 5.4, while men's averages 6.8 (*p*<0.001).

How does Canada compare with other countries? Across all 57 countries surveyed during wave 7 of the WVS (2017--22), the average political interest was 4.8/10 for men and 4/10 for women, and while political interest increased with people's age, the size of the gender gap remained relatively similar with age. Among Canadian WVS respondents, the average political interest was 6.2/10 for men and 5/10 for women: this gender gap is 0.35 points larger than the WVS average and is statistically significant (*p*<0.001). Moreover, as mentioned previously, the age pattern was somewhat different in the Canadian WVS compared with the WVS's cross-country results, becoming statistically insignificant around age 75. However, sample sizes are smaller for people aged 75 and over, and age averages vary substantially from year to year as the CES and WVS scatterplots show. The fact that all three surveys show a shrinking of the gap at that age, however, reinforces confidence in that finding.

@fig-yearceswvs and @fig-yearethnicityceswvs

```{r}
#| label: fig-yearceswvs
#| echo: false
#| fig-cap: "General Political Interest by Year and Gender Among Canadian Adults, CES, WVS (Canada) and GSS \\newline \\textit{Notes}: On the \\textit{y} axis, 0 = no interest at all, and 10 = a great deal of interest. CES and WVS weights are applied."
knitr::include_graphics("_graphs/InterestYearGender.pdf")
```

```{r}
#| label: fig-yearethnicityceswvs
#| echo: false
#| fig-cap: "General Political Interest by Year, Gender and Ethnicity Among Canadian Adults, 2021 CES and 2020 WVS (Canada) \\newline \\textit{Notes}: On the \\textit{y} axis, 0 = no interest at all, and 10 = a great deal of interest. CES and WVS weights are applied."
knitr::include_graphics("_graphs/InterestWaveGroup.pdf")
```

### Canadian Children Political Interest Survey

What kinds of political topics are children and adolescents interested in? Are there gender differences in how interested they are in these topics? Does gender predict political interest in itself, or is it only a proxy for other variables? Data from the Canadian Children Political Interest Survey (CCPIS), described in further detail in @sec-chap2, is used to answer these questions. @tbl-lmeInterestCCPIS shows the link between gender and interest in each of the topics among elementary and high school students. The upper part includes gender as the only predictor, while the lower part includes controls for socio-demographic variables.

In the upper half of the table, taking into account classroom fixed effects, boys generally report being more interested in politics than girls, but the gender gap is relatively minimal, standing at 0.43 for the 11-point political interest scale (*p*<0.05). Boys' interest in international affairs and partisan politics is higher than girls' (1-point and 0.85-point difference respectively; both *p*<0.001). This seems to be in line with the results among adult respondents. However, girls' interest in law and crime is also slightly higher (0.5 point, *p*<0.05). This result is more surprising given previous literature showing the contrary. Results for the other two topics, health care and education, are almost even between the genders, and not statistically significant, despite adult women usually reporting higher levels of interest in these two topics across studies.

However, in the lower half of the table, controlling for socio-demographic factors, all relationships between gender and interest disappear. It might be the case that girls are less interested in some of these topics by virtue of other socio-demographic characteristics, but there does not seem to be one other socio-demographic variable that reliably predicts interest in any of the topics. These multiple regression results might be tainted by some multicollinearity between ethnicity, immigrant status and language.

Table: Students' political interest by gender {#tbl-lmeInterestCCPIS}

|                                     | Politics (general) | Health care | International affairs | Law and crime | Education | Partisan politics | 
|:------------------------------------|-------------------:|------------:|----------------------:|--------------:|----------:|------------------:|
| (Intercept)                         |      4.579***      |   4.041***  |        5.724***       |    4.956***   |  4.219*** |      4.007***     | 
|                                     |       (0.184)      |   (0.167)   |        (0.180)        |    (0.173)    |  (0.206)  |      (0.171)      | 
| Gender (1 = girls)                  |       -0.434*      |    0.128    |       -0.980***       |     0.488*    |   -0.103  |     -0.854***     | 
|                                     |       (0.207)      |   (0.197)   |        (0.229)        |    (0.231)    |  (0.223)  |      (0.232)      | 
| SD (Intercept Class)                |        0.651       |    0.542    |         0.473         |     0.377     |   0.771   |       0.344       | 
| SD (Observations)                   |        2.499       |    2.397    |         2.802         |     2.837     |   2.701   |       2.855       | 
| Num.Obs.                            |         617        |     623     |          620          |      619      |    623    |        620        | 
| R2 Marg.                            |        0.007       |    0.001    |         0.029         |     0.007     |   0.000   |       0.022       | 
|:------------------------------------|-------------------:|------------:|----------------------:|--------------:|----------:|------------------:|
| (Intercept)                         |       -4.021       |    -1.235   |         0.253         |     4.881     |  27.025+  |       10.330      | 
|                                     |      (13.493)      |   (13.062)  |        (14.952)       |    (15.600)   |  (15.173) |      (14.425)     | 
| Gender (1 = girls)                  |        3.071       |    -0.092   |         -3.122        |     0.357     |   0.845   |       1.237       | 
|                                     |       (2.651)      |   (2.478)   |        (2.825)        |    (2.829)    |  (2.693)  |      (2.914)      | 
| Age                                 |        0.346       |    0.359    |         0.218         |     -0.393    |  -3.560+  |       -1.227      | 
|                                     |       (1.796)      |   (1.736)   |        (1.994)        |    (2.080)    |  (2.008)  |      (1.932)      | 
| Age squared                         |        0.000       |    -0.003   |         0.000         |     0.022     |   0.129+  |       0.045       | 
|                                     |       (0.060)      |   (0.058)   |        (0.066)        |    (0.069)    |  (0.066)  |      (0.064)      | 
| Ethnicity (1 = white)                    |       0.659+       |    0.154    |         0.860*        |     -0.203    |   0.049   |       0.356       | 
|                                     |       (0.378)      |   (0.356)   |        (0.409)        |    (0.411)    |  (0.395)  |      (0.422)      | 
| Immigrant                           |       -0.072       |    -0.165   |        -0.748+        |    -0.956*    |   -0.179  |       -0.428      | 
|                                     |       (0.421)      |   (0.392)   |        (0.452)        |    (0.458)    |  (0.433)  |      (0.467)      | 
| English spoken at home              |       -0.918       |   -1.020+   |        -1.255+        |     -0.515    |   -0.453  |       -0.264      | 
|                                     |       (0.609)      |   (0.565)   |        (0.651)        |    (0.653)    |  (0.625)  |      (0.670)      | 
| French spoken at home               |       -0.088       |    -0.099   |         -0.591        |     -0.112    |   0.426   |       0.351       | 
|                                     |       (0.342)      |   (0.319)   |        (0.370)        |    (0.371)    |  (0.351)  |      (0.382)      | 
| Agency scale                        |      3.268***      |    1.127    |         2.185*        |     1.748*    |   0.674   |      2.343**      | 
|                                     |       (0.785)      |   (0.730)   |        (0.844)        |    (0.846)    |  (0.806)  |      (0.875)      | 
| Communality scale                   |        1.251       |    0.719    |         1.986*        |     0.954     |   1.042   |       0.603       | 
|                                     |       (0.839)      |   (0.791)   |        (0.910)        |    (0.913)    |  (0.875)  |      (0.937)      | 
| School #4                           |        0.196       |   -0.958*   |         -0.577        |     -0.475    |   0.002   |       -0.444      | 
|                                     |       (0.351)      |   (0.407)   |        (0.453)        |    (0.493)    |  (0.531)  |      (0.390)      | 
| School #3                           |        0.990       |    -0.113   |         0.797         |     -0.518    |   -0.467  |       0.686       | 
|                                     |       (0.708)      |   (0.754)   |        (0.836)        |    (0.890)    |  (0.933)  |      (0.761)      | 
| School #6                           |        0.875       |    0.465    |         3.969         |     1.924     |   -3.158  |       1.760       | 
|                                     |       (2.157)      |   (2.122)   |        (2.396)        |    (2.479)    |  (2.468)  |      (2.338)      | 
| School #2                           |        0.195       |   -0.817+   |         -0.551        |     -0.547    |   0.111   |       -0.447      | 
|                                     |       (0.373)      |   (0.433)   |        (0.483)        |    (0.527)    |  (0.564)  |      (0.416)      | 
| School #5                           |       -0.529       |    -1.081   |         -0.263        |     -1.482    |   1.013   |       -0.320      | 
|                                     |       (1.530)      |   (1.461)   |        (1.678)        |    (1.698)    |  (1.651)  |      (1.710)      | 
| School #8                           |       2.824**      |    0.324    |         -1.542        |     -0.392    |   3.343*  |       0.768       | 
|                                     |       (0.903)      |   (0.971)   |        (1.093)        |    (1.161)    |  (1.216)  |      (1.007)      | 
| School #7                           |       -1.601       |    0.070    |         -3.829        |     -3.370    |   -2.073  |       -2.819      | 
|                                     |       (2.045)      |   (1.957)   |        (2.891)        |    (2.923)    |  (2.843)  |      (2.950)      | 
| Gender (1 = girls):Age              |       -0.174       |    0.050    |         0.173         |     0.030     |   -0.030  |       -0.095      | 
|                                     |       (0.167)      |   (0.156)   |        (0.178)        |    (0.179)    |  (0.170)  |      (0.183)      | 
| Gender (1 = girls):Ethnicity (1 = white) |       -1.191*      |    -0.542   |         -0.718        |     -0.355    |  -1.035+  |      -0.962+      | 
|                                     |       (0.520)      |   (0.488)   |        (0.560)        |    (0.563)    |  (0.538)  |      (0.579)      | 
| SD (Intercept Class)                |        0.001       |    0.432    |         0.450         |     0.571     |   0.706   |       0.001       | 
| SD (Observations)                   |        2.460       |    2.297    |         2.647         |     2.646     |   2.524   |       2.758       | 
| Num.Obs.                            |         423        |     427     |          425          |      423      |    427    |        427        | 
| R2 Marg.                            |        0.121       |    0.071    |         0.130         |     0.073     |   0.102   |       0.085       | 
|:------------------------------------|-------------------:|------------:|----------------------:|--------------:|----------:|------------------:|

__Note:__
^^ + p < 0.1, * p < 0.05, ** p < 0.01, *** p < 0.001

__Note:__
^^ Multilevel regression with random effects at the classroom level

__Note:__
^^ Reference category for language: Other languages spoken at home

@tbl-lmeInterestYoungCCPIS and @tbl-lmeInterestOldCCPIS re-analyze those results for children aged 10--14 (*n*=200) and those aged 15--18 (*n*=446). Among younger students, the gender gap in self-reported political interest is 0.06/10 (not statistically significant). Among those aged 15--18, this gap grows to 0.53/10 (*p*<0.05). Notably, among both groups, both interest in international affairs and interest in partisan politics are higher among boys than girls, suggesting some gender differences in interests might take place before adolescence.

Table: Political interest by gender (ages 10-15) {#tbl-lmeInterestYoungCCPIS}

|                                     | Politics (general) | Health care | International affairs | Law and crime | Education | Partisan politics | 
|:------------------------------------|-------------------:|------------:|----------------------:|--------------:|----------:|------------------:|
| (Intercept)                         |      3.990***      |   3.638***  |        5.356***       |    4.519***   |  3.569*** |      3.854***     | 
|                                     |       (0.294)      |   (0.232)   |        (0.310)        |    (0.293)    |  (0.257)  |      (0.321)      | 
| Gender (1 = girls)                  |       -0.113       |    0.077    |        -0.891*        |     0.630+    |   -0.073  |      -0.755*      | 
|                                     |       (0.319)      |   (0.309)   |        (0.375)        |    (0.379)    |  (0.357)  |      (0.364)      | 
| SD (Intercept Class)                |        0.703       |    0.205    |         0.561         |     0.392     |   0.000   |       0.707       | 
| SD (Observations)                   |        2.527       |    2.467    |         2.965         |     3.022     |   2.847   |       2.860       | 
| Num.Obs.                            |         256        |     256     |          253          |      256      |    254    |        251        | 
| R2 Marg.                            |        0.000       |    0.000    |         0.021         |     0.011     |   0.000   |       0.016       | 
|:------------------------------------|-------------------:|------------:|----------------------:|--------------:|----------:|------------------:|
| (Intercept)                         |       19.493       |    -3.115   |        -52.289        |    -56.980    |   22.959  |       15.364      | 
|                                     |      (46.116)      |   (45.145)  |        (53.853)       |    (54.491)   |  (52.090) |      (53.562)     | 
| Gender (1 = girls)                  |        2.134       |    -1.485   |         -4.173        |     -4.736    |   -0.123  |       0.617       | 
|                                     |       (4.901)      |   (4.762)   |        (5.362)        |    (5.406)    |  (5.185)  |      (5.331)      | 
| Age                                 |       -3.163       |    0.778    |         8.228         |     9.313     |   -2.621  |       -1.619      | 
|                                     |       (6.914)      |   (6.771)   |        (8.072)        |    (8.166)    |  (7.808)  |      (8.029)      | 
| Age squared                         |        0.123       |    -0.023   |         -0.303        |     -0.357    |   0.082   |       0.050       | 
|                                     |       (0.257)      |   (0.252)   |        (0.299)        |    (0.303)    |  (0.290)  |      (0.298)      | 
| Ethnicity (1 = white)                    |        0.798       |    0.386    |         1.214         |     -0.307    |   1.312   |       0.185       | 
|                                     |       (0.770)      |   (0.741)   |        (0.830)        |    (0.844)    |  (0.815)  |      (0.838)      | 
| Immigrant                           |        0.683       |    -0.430   |         -0.171        |     2.142+    |   0.835   |       0.416       | 
|                                     |       (1.110)      |   (1.087)   |        (1.237)        |    (1.254)    |  (1.204)  |      (1.238)      | 
| English spoken at home              |       -1.859+      |   -2.249*   |        -2.509*        |     -0.496    |  -2.152*  |      -2.699*      | 
|                                     |       (1.048)      |   (0.982)   |        (1.116)        |    (1.139)    |  (1.070)  |      (1.100)      | 
| French spoken at home               |       -0.209       |    -0.889   |         -0.577        |     0.150     |   -0.750  |       0.094       | 
|                                     |       (0.617)      |   (0.608)   |        (0.694)        |    (0.703)    |  (0.667)  |      (0.686)      | 
| Agency scale                        |       2.997*       |    1.857    |         1.902         |     1.545     |   0.477   |       1.246       | 
|                                     |       (1.416)      |   (1.372)   |        (1.585)        |    (1.605)    |  (1.532)  |      (1.575)      | 
| Communality scale                   |       3.035*       |    0.343    |         2.812+        |     1.665     |   0.843   |       1.238       | 
|                                     |       (1.449)      |   (1.422)   |        (1.613)        |    (1.640)    |  (1.561)  |      (1.605)      | 
| School #4                           |        0.540       |    -1.080   |         -1.473        |     -1.230    |   -0.298  |       -0.599      | 
|                                     |       (0.887)      |   (0.823)   |        (0.865)        |    (0.930)    |  (0.792)  |      (0.814)      | 
| School #3                           |        1.080       |    0.028    |         -0.071        |     -0.844    |   -0.588  |       0.275       | 
|                                     |       (1.031)      |   (0.945)   |        (0.965)        |    (1.038)    |  (0.879)  |      (0.904)      | 
| School #6                           |       -0.713       |    0.834    |         6.154         |     4.838     |   -3.186  |       0.685       | 
|                                     |       (3.242)      |   (3.130)   |        (3.620)        |    (3.700)    |  (3.462)  |      (3.560)      | 
| School #2                           |        1.302       |    -0.371   |         -1.267        |     0.618     |   1.419   |       -0.284      | 
|                                     |       (0.993)      |   (0.922)   |        (0.953)        |    (1.036)    |  (0.861)  |      (0.885)      | 
| School #5                           |       -0.136       |    -2.098   |         -0.287        |     -4.974    |   0.679   |       -1.899      | 
|                                     |       (3.017)      |   (2.937)   |        (3.309)        |    (3.380)    |  (3.191)  |      (3.281)      | 
| School #8                           |       4.630+       |    1.587    |         0.726         |     -0.525    |   5.375*  |       5.597*      | 
|                                     |       (2.250)      |   (2.156)   |        (2.397)        |    (2.475)    |  (2.284)  |      (2.349)      | 
| School #7                           |       -1.114       |    0.649    |         -3.243        |     -3.716    |   -2.055  |       -1.443      | 
|                                     |       (2.416)      |   (2.311)   |        (3.249)        |    (3.325)    |  (3.122)  |      (3.211)      | 
| Gender (1 = girls):Age              |       -0.102       |    0.184    |         0.238         |     0.444     |   0.098   |       -0.074      | 
|                                     |       (0.341)      |   (0.332)   |        (0.376)        |    (0.379)    |  (0.363)  |      (0.373)      | 
| Gender (1 = girls):Ethnicity (1 = white) |       -1.261       |    -1.049   |         -0.264        |     -0.967    |  -1.985+  |       -0.387      | 
|                                     |       (0.986)      |   (0.956)   |        (1.073)        |    (1.087)    |  (1.051)  |      (1.080)      | 
| SD (Intercept Class)                |        0.587       |    0.482    |         0.336         |     0.499     |   0.000   |       0.000       | 
| SD (Observations)                   |        2.516       |    2.468    |         2.826         |     2.860     |   2.746   |       2.823       | 
| Num.Obs.                            |         159        |     159     |          158          |      159      |    159    |        159        | 
| R2 Marg.                            |        0.165       |    0.094    |         0.207         |     0.129     |   0.129   |       0.169       | 
|:------------------------------------|-------------------:|------------:|----------------------:|--------------:|----------:|------------------:|

__Note:__
^^ + p < 0.1, * p < 0.05, ** p < 0.01, *** p < 0.001

__Note:__
^^ Multilevel regression with random effects at the classroom level

__Note:__
^^ Reference category for language: Other languages spoken at home

Table: Political interest by gender (ages 16-18) {#tbl-lmeInterestOldCCPIS}

|                                     | Politics (general) | Health care | International affairs | Law and crime | Education | Partisan politics | 
|:------------------------------------|-------------------:|------------:|----------------------:|--------------:|----------:|------------------:|
| (Intercept)                         |      4.997***      |   4.365***  |        5.903***       |    5.291***   |  4.621*** |      4.097***     | 
|                                     |       (0.184)      |   (0.202)   |        (0.209)        |    (0.208)    |  (0.247)  |      (0.202)      | 
| Gender (1 = girls)                  |       -0.546*      |    0.213    |        -0.932**       |     0.473     |   -0.013  |      -0.886**     | 
|                                     |       (0.271)      |   (0.261)   |        (0.293)        |    (0.293)    |  (0.293)  |      (0.302)      | 
| SD (Intercept Class)                |        0.174       |    0.458    |         0.318         |     0.321     |   0.674   |       0.000       | 
| SD (Observations)                   |        2.484       |    2.354    |         2.682         |     2.659     |   2.621   |       2.815       | 
| Num.Obs.                            |         345        |     349     |          349          |      345      |    351    |        351        | 
| R2 Marg.                            |        0.012       |    0.002    |         0.029         |     0.008     |   0.000   |       0.024       | 
|:------------------------------------|-------------------:|------------:|----------------------:|--------------:|----------:|------------------:|
| (Intercept)                         |      -136.895      |   -30.503   |        149.356        |     98.815    |  220.357  |      -246.198     | 
|                                     |      (156.856)     |  (137.403)  |       (161.764)       |   (167.899)   | (149.124) |     (164.357)     | 
| Gender (1 = girls)                  |       -0.483       |    5.854    |         -3.756        |     15.528    |   5.428   |       2.834       | 
|                                     |       (9.780)      |   (8.948)   |        (10.326)       |    (10.522)   |  (9.723)  |      (10.684)     | 
| Age                                 |       16.603       |    3.632    |        -17.463        |    -12.298    |  -26.984  |       29.632      | 
|                                     |      (18.852)      |   (16.517)  |        (19.446)       |    (20.234)   |  (17.927) |      (19.758)     | 
| Age squared                         |       -0.494       |    -0.094   |         0.525         |     0.399     |   0.836   |       -0.884      | 
|                                     |       (0.566)      |   (0.496)   |        (0.584)        |    (0.609)    |  (0.538)  |      (0.593)      | 
| Ethnicity (1 = white)                    |        0.552       |    0.026    |         0.722         |     -0.086    |   -0.382  |       0.448       | 
|                                     |       (0.440)      |   (0.412)   |        (0.472)        |    (0.463)    |  (0.448)  |      (0.491)      | 
| Immigrant                           |       -0.172       |    0.094    |        -1.108*        |    -1.520**   |   -0.366  |       -0.490      | 
|                                     |       (0.467)      |   (0.430)   |        (0.493)        |    (0.492)    |  (0.466)  |      (0.516)      | 
| English spoken at home              |       -0.249       |    -0.201   |         -0.466        |     -0.496    |   0.658   |       1.448+      | 
|                                     |       (0.767)      |   (0.708)   |        (0.813)        |    (0.797)    |  (0.765)  |      (0.852)      | 
| French spoken at home               |        0.030       |    0.317    |         -0.626        |     -0.168    |   0.980*  |       0.568       | 
|                                     |       (0.422)      |   (0.384)   |        (0.443)        |    (0.436)    |  (0.415)  |      (0.463)      | 
| Agency scale                        |       3.209**      |    0.760    |         1.971+        |     2.017*    |   0.703   |       2.733*      | 
|                                     |       (0.964)      |   (0.885)   |        (1.015)        |    (0.998)    |  (0.957)  |      (1.065)      | 
| Communality scale                   |        0.232       |    0.935    |         1.294         |     0.460     |   1.115   |       0.267       | 
|                                     |       (1.064)      |   (0.985)   |        (1.131)        |    (1.111)    |  (1.068)  |      (1.179)      | 
| School #4                           |        0.224       |   -0.896+   |         -0.170        |     -0.203    |   0.441   |       -0.220      | 
|                                     |       (0.433)      |   (0.503)   |        (0.540)        |    (0.518)    |  (0.660)  |      (0.480)      | 
| School #2                           |       -0.041       |   -0.995+   |         -0.196        |     -0.789    |   -0.011  |       -0.371      | 
|                                     |       (0.437)      |   (0.503)   |        (0.541)        |    (0.522)    |  (0.650)  |      (0.485)      | 
| School #5                           |       -0.653       |    -0.478   |         -0.833        |     -1.418    |   0.555   |       0.690       | 
|                                     |       (1.873)      |   (1.752)   |        (2.004)        |    (1.962)    |  (1.954)  |      (2.061)      | 
| School #8                           |       2.129+       |    -0.313   |         -1.866        |     -0.127    |   2.901*  |       -1.063      | 
|                                     |       (1.056)      |   (1.107)   |        (1.221)        |    (1.184)    |  (1.358)  |      (1.177)      | 
| Gender (1 = girls):Age              |        0.054       |    -0.326   |         0.226         |     -0.899    |   -0.322  |       -0.163      | 
|                                     |       (0.598)      |   (0.547)   |        (0.631)        |    (0.643)    |  (0.594)  |      (0.653)      | 
| Gender (1 = girls):Ethnicity (1 = white) |       -1.343*      |    -0.186   |        -1.164+        |     -0.536    |   -0.736  |      -1.486*      | 
|                                     |       (0.638)      |   (0.591)   |        (0.677)        |    (0.667)    |  (0.640)  |      (0.707)      | 
| SD (Intercept Class)                |        0.000       |    0.460    |         0.426         |     0.386     |   0.758   |       0.000       | 
| SD (Observations)                   |        2.423       |    2.229    |         2.562         |     2.513     |   2.400   |       2.700       | 
| Num.Obs.                            |         264        |     268     |          267          |      264      |    268    |        268        | 
| R2 Marg.                            |        0.088       |    0.049    |         0.099         |     0.083     |   0.099   |       0.092       | 
|:------------------------------------|-------------------:|------------:|----------------------:|--------------:|----------:|------------------:|

__Note:__
^^ + p < 0.1, * p < 0.05, ** p < 0.01, *** p < 0.001

__Note:__
^^ Multilevel regression with random effects at the classroom level

__Note:__
^^ Reference category for language: Other languages spoken at home

For health care and education, results among Canadian children --- which show no gender gap --- might come as a surprise given the number of studies that show adult women report being more interested in the political aspects of these topics in various contexts. Still, the experiences of several women as mothers and caregivers later in life might be the elements that shape the importance they give to both issues.

@fig-political shows how students view each of the 10 concrete issues associated with the 5 topics as political or non-political. As expected, the two issues related to partisan politics are almost universally seen as political, followed by issues related to international affairs. These are the two topics for which boys report being more interested. Findings for issues related to the other three topics are perceived by 25% to 60% of students as being non-political --- substantially large proportions. @campbellrosie2008 and @ferrin2020 had similarly found that issues related to partisan politics were seen as more political than those related to other topics. Visibly, this trend also applies to Canadian children, which suggests that topics which are seen as political are also those for which men typically report a higher degree of interest.

```{r}
#| label: fig-political
#| echo: false
#| fig-cap: "Views of Topics as Political or Non-Political By Canadian Students, 2022 CCPIS"
knitr::include_graphics("_graphs/CCPISPolitical.pdf")
```

### Datagotchi PES

How do these interests change when people reach adulthood? This question is best answered using Datagotchi PES data. As @tbl-olsInterestDg shows, Quebec men generally report being more interested in politics than women. Men's interest in international affairs and partisan politics is also higher than women's (0.7-point and 0.6-point difference respectively; both *p*<0.001). Men's interest in law and order is also slightly higher (0.2 point, *p*<0.1). However, women are significantly more interested in health care and education (0.7-point and 0.5-point difference respectively, both *p*<0.001) than men are.

Table: Political interest by gender {#tbl-olsInterestDg}

|                                     | Politics (general) | Health care | International affairs | Law and crime | Education | Partisan politics | 
|:------------------------------------|-------------------:|------------:|----------------------:|--------------:|----------:|------------------:|
| (Intercept)                         |      7.736***      |   6.473***  |        7.517***       |    5.408***   |  6.607*** |      6.342***     | 
|                                     |       (0.059)      |   (0.070)   |        (0.066)        |    (0.073)    |  (0.070)  |      (0.077)      | 
| Gender (1 = women)                  |      -0.828***     |   0.706***  |       -0.653***       |     0.184+    |  0.530*** |     -0.565***     | 
|                                     |       (0.090)      |   (0.108)   |        (0.102)        |    (0.112)    |  (0.107)  |      (0.119)      | 
| Num.Obs.                            |        1800        |     1800    |          1800         |      1800     |    1800   |        1800       | 
| R2                                  |        0.045       |    0.023    |         0.022         |     0.002     |   0.013   |       0.012       | 
| R2 Adj.                             |        0.044       |    0.023    |         0.022         |     0.001     |   0.013   |       0.012       | 
| Log.Lik.                            |      -3702.748     |  -4020.146  |       -3922.656       |   -4088.971   | -4010.567 |     -4202.267     | 
| F                                   |       83.948       |    42.885   |         40.871        |     2.710     |   24.381  |       22.384      | 
|:------------------------------------|-------------------:|------------:|----------------------:|--------------:|----------:|------------------:|
| (Intercept)                         |      7.266***      |   4.359***  |        6.160***       |    5.391***   |  6.052*** |      6.922***     | 
|                                     |       (0.676)      |   (0.791)   |        (0.754)        |    (0.843)    |  (0.798)  |      (0.897)      | 
| Gender (1 = women)                  |       -0.557       |    1.538*   |         -0.562        |     0.426     |   -0.288  |       -1.005      | 
|                                     |       (0.627)      |   (0.733)   |        (0.698)        |    (0.781)    |  (0.740)  |      (0.831)      | 
| Age                                 |       -0.034+      |    0.038+   |         -0.006        |     0.014     |   -0.025  |      -0.073**     | 
|                                     |       (0.019)      |   (0.023)   |        (0.022)        |    (0.024)    |  (0.023)  |      (0.026)      | 
| Age squared                         |       0.000*       |    0.000    |         0.000         |     0.000     |   0.000   |       0.001*      | 
|                                     |       (0.000)      |   (0.000)   |        (0.000)        |    (0.000)    |  (0.000)  |      (0.000)      | 
| Ethnicity (1 = white)                    |        0.496       |    -0.262   |         -0.244        |    -0.936+    |   -0.449  |       0.558       | 
|                                     |       (0.443)      |   (0.518)   |        (0.493)        |    (0.552)    |  (0.523)  |      (0.587)      | 
| Immigrant                           |       -0.354+      |    0.345    |         0.262         |    -0.463+    |   0.139   |       -0.332      | 
|                                     |       (0.207)      |   (0.242)   |        (0.230)        |    (0.258)    |  (0.244)  |      (0.274)      | 
| French spoken at home               |        0.163       |    0.132    |         0.481         |     -0.058    |   0.381   |       0.209       | 
|                                     |       (0.292)      |   (0.341)   |        (0.325)        |    (0.363)    |  (0.344)  |      (0.387)      | 
| Income between \$60,000 and \$150,000 |        0.054       |    0.138    |         0.214         |     0.263+    |   0.080   |       0.164       | 
|                                     |       (0.123)      |   (0.144)   |        (0.138)        |    (0.154)    |  (0.146)  |      (0.164)      | 
| Income above \$150,000               |        0.232       |    0.306+   |        0.493**        |    0.482**    |   0.207   |       0.434*      | 
|                                     |       (0.143)      |   (0.167)   |        (0.159)        |    (0.178)    |  (0.169)  |      (0.190)      | 
| Education: college                  |        0.164       |    0.254    |         0.366+        |     0.201     |   0.604*  |       0.116       | 
|                                     |       (0.199)      |   (0.232)   |        (0.221)        |    (0.248)    |  (0.234)  |      (0.263)      | 
| Education: university               |       0.485**      |    0.407+   |        0.615**        |     -0.121    |  1.324*** |       0.475+      | 
|                                     |       (0.185)      |   (0.216)   |        (0.206)        |    (0.230)    |  (0.218)  |      (0.245)      | 
| Gender (1 = women):Age              |       0.018**      |    0.005    |        0.019**        |     -0.011    |   -0.001  |      0.026***     | 
|                                     |       (0.006)      |   (0.007)   |        (0.006)        |    (0.007)    |  (0.007)  |      (0.008)      | 
| Gender (1 = women):Ethnicity (1 = white) |       -1.119+      |    -1.014   |         -0.953        |     0.369     |   0.840   |       -0.806      | 
|                                     |       (0.592)      |   (0.692)   |        (0.659)        |    (0.737)    |  (0.698)  |      (0.785)      | 
| Num.Obs.                            |        1800        |     1800    |          1800         |      1800     |    1800   |        1800       | 
| R2                                  |        0.071       |    0.087    |         0.075         |     0.019     |   0.050   |       0.030       | 
| R2 Adj.                             |        0.065       |    0.081    |         0.069         |     0.012     |   0.044   |       0.024       | 
| Log.Lik.                            |      -3677.474     |  -3959.286  |       -3872.398       |   -4073.506   | -3976.231 |     -4185.857     | 
| F                                   |       11.392       |    14.219   |         12.132        |     2.809     |   7.889   |       4.628       | 
|:------------------------------------|-------------------:|------------:|----------------------:|--------------:|----------:|------------------:|

__Note:__
^^ + p < 0.1, * p < 0.05, ** p < 0.01, *** p < 0.001

__Note:__
^^ Ordinary least squares (OLS) regression

@fig-datagotchi inquires about any time trends in the evolution of interest in each of the topics, including gendered aspects. No clear gender pattern emerges, beyond the finding that the gap in general political interest becomes statistically insignificant among Quebeckers aged 75 and over.

```{r}
#| label: fig-datagotchi
#| echo: false
#| fig-cap: "Self-Reported Level of Interest in Various Topics by Age Among Canadian Adults, 2022 Datagotchi PES \\newline \\textit{Notes}: On the \\textit{y} axis, 0 = no interest at all, and 10 = a great deal of interest. Dots represent average interest by age and gender. No weights are currently applied."
knitr::include_graphics("_graphs/InterestAgeGenderDGPanel.pdf")
```

## Discussion

This chapter highlights gender differences in interest in various political topics, but also shows differences between children and adults. For international relations and partisan politics, male students report more interest in these topics, just like adults in previous studies. However, girls' higher interest in law and crime and similar levels of interest in health care and education compared to boys contrast with data previously found among adults [@campbellrosie2008; @coffe2013; @ferrin2020; @hayes1993; @kuhn2004; @sabella2004; @verba1997]. Given the results of the Datagotchi PES, which shows higher levels of interest in the political aspects of health care and education among adult women than men, it seems that the gender gap in interest in these topics emerges in early adulthood, perhaps as a result of life events that happen at that stage, such as women becoming more likely to care for children and relatives. Given the absence of longitudinal data collected among Canadians, it is not possible to tell at what point in time girls or women become more interested in health care and education issues relative to men. However, given the current results of two datasets, it is possible to note, for the first time, that some differences in interests are not necessarily pre-existent among adolescents. On the other hand, reported interest in politics in general, partisan politics and international affairs is already higher among boys aged 10 to 18.

@russo2017 suggested that the 13--15 age range is a period when parents, teachers, and other role models have the highest potential for transmission of interest in politics to children. Applying these results to the Canadian context, it appears that the 15--25-year-old period, similar to what previous studies have found, is also the point in time where a gender gap in interest for politics more generally emerges, with boys/men reporting higher interest in politics in general than girls/women.

## References
