---
title: "Parent--Child Political Interest Transmission: Do Moms Influence their Daughters and Dads Influence their Sons? {#sec-chap4}"
author: "Alexandre Fortier-Chouinard"
bibliography: Thesis.bib
format: pdf
toc: true
toc-depth: 5
---
Parents play a prominent role in raising their children and therefore are among the main actors who can transmit political interest to them. Political scientists have long studied the extent of parents' role in that regard and the mechanisms through which political interest can be transmitted. @campbell1960 already suggested that "interest in politics, like partisanship, is readily transmitted within the family from generation to generation" (p.&nbsp;413). More recent research has highlighted inequalities in the extent to which boys and girls match their parents' level of political interest. This chapter aims to uncover the complex role gender can play in mediating parent--child political interest transmission. When and how does political interest transmission occur? To what extent do mothers influence political interest development in their children? Do they have a greater influence on their sons or daughters? What about fathers? Do these results hold for different political topics, when political interest is measured by sector? The chapter relies on social learning theory and seeks to test this dissertation's first hypothesis highlighted in @sec-chap1: *Children's interest in specific political topics is more related to political discussions with their same-gender parent(s) than other-gender parent(s).*

## Political Interest Transmission

The transmission of interest in politics between parents and children has typically been measured by using methods to statistically compare their answers to the same political interest questions. Studies have generally shown a significant relationship between parents' political interest and their children's political interest [@beauregard2008; @janmaat2022; @neundorf2013; @prior2019; @shehata2019]. While @jennings2009 find no statistically significant relationship, their results show that successful parent--child political interest transmission occurs when the family environment is more politicized. @prior2019 estimates moderately strong Pearson correlation coefficients of 0.3 to 0.4 for parent--child political interest scores across three countries. By comparison, correlation coefficients of different indicators of political interest for the same individual vary between 0.6 and 0.7.

Parent--child political interest correlations vary by age. @prior2019 finds very weak (0.05) correlations at age 11, followed by a steady growth until age 15. Parent--child correlations then remain stronger when both parents share a similar level of political interest, when this parental political interest is stable through time, and when children move out late from their parents' place. Children who move out early of their parents' place tend to see a quick drop in the extent to which their political interest matches their parents'.

While parents' political interest often seems to match their children's political interest, most studies lack the kind of data needed to establish a causal link between both. However, @prior2019 uses panel data collected among parents and children and finds a weak but noticeable causal link in parental transmission of political interest. In time series, an increase in mothers' political interest is often accompanied by an increase in their children's political interest. The same goes for fathers, and for decreases rather than increases. These trends are clearer in the United Kingdom and Germany, despite weaker evidence in Switzerland. In the first two countries, it seems reasonable to assume that a change in parents' political interest could cause a similar change in their children's political interest.

The main causal mechanism for this transmission process seems to be parent--child political discussions. Scholars have long suggested that the development of political interest can happens through increasingly complex discussions about political topics between parents and children at home [@easton1969; @greenstein1965]. These discussions can be initiated either by the parent or by the child, and the more they occur, the likelier the child is to be interested in politics. This relationship between the frequency of political discussions with parents and the child's interest in politics is statistically significant in the United States [@shehata2019], China, Mexico, Japan [@mayer2004], and Canada [@dostiegoulet2009en]. Furthermore, @shehata2019 find that changes in the frequency of political discussions with parents positively predict changes in adolescents' level of political interest.

It is also important to recognize that children can play a role in shaping political discussions with their parents --- and potentially their parents' interest in politics. @mcdevitt2002 find evidence of trickle-up political socialization, in which adolescent children initiate political discussions with parents, who react by increasing their news consumption or finding other ways to gain knowledge about politics in order to maintain a leadership role in the family. In this study, children's interest is first triggered through a civics curriculum.^[The influence of schools, teachers, and citizenship education are studied in more detail in @sec-chap6.] @york2019 even finds that adolescents' news use and political discussions with peers have a positive influence on their political discussions with parents, while the opposite effects --- political discussions with parents influencing adolescents' news use or political discussions with peers --- are not found to hold.

Yet, other studies about the relative influence of various socialization agents on the development of political interest in children and adolescents [@dostiegoulet2009en; @koskimaa2015; @shehata2019] mostly find that parents play a larger role in transmitting political interest than any other socialization agent --- including friends, media and schools. Trickle-down political socialization --- from parents to children --- therefore remains an important explanation of children's development of political interest.

## Gender Differences in Transmission

Research has found that the trickle-down effect of political interest from parents to children works in gendered ways, and much of that evidence suggests transmission works better for parent--child pairs of the same gender. From the perspective of parents, mothers' political interest has a stronger effect on their daughters than sons' political interest, while fathers' political interest has a stronger effect on their sons' political interest [@beauregard2008; @owen1988; @prior2019]. Disregarding parents' gender, @sabella2004political finds that parents' potential to transmit political interest to their sons is stronger than to their daughters. From the perspective of children, daughters' political interest seems to be influenced mostly by their mothers, with the mother--daughter political interest link stronger than all other combinations [@beauregard2008; @owen1988; @prior2019], although @rebenstorf2004political strikes a discordant note, finding that fathers' potential to transmit political interest to their children is stronger than mothers'. Finally, it is not clear whether sons' political interest is influenced mostly by their father [@beauregard2008; @owen1988] or mother [@prior2019].

The strength of parent--child political interest correlations by gender is inquired by @prior2019, whose findings are more recent and rely on panel data. In the British dataset, they find mother--daughter political interest correlations are the strongest (0.43), while mother--son, father--daughter and father--son correlations all sit between 0.31 and 0.34. In their German dataset, all four pairs are relatively close together, but exact numbers are not provided by the author. Finally, in the Swiss dataset, mother--daughter correlations are the strongest (0.36), followed by mother--son (0.31), father--son (0.21) and father--daughter (0.2). There is overall a stronger causal effect of mothers' political interest on their children's political interest, compared with fathers.

Studies have investigated the gender patterns in parent--child political discussions, but the amount of political discussions does not seem to vary based on parents' and children's gender. While earlier studies found that children discuss politics more often with their fathers than mothers [@noller1985; @oswald1998], more recent research has found no significant difference between fathers and mothers [@hooghe2015; @mayer2004; @shulman2014]. @noller1985 also find that parents discuss politics more often with their sons. However, most studies again find a different pattern in which parents discuss politics with their daughters as much as with their sons [@dowse1971; @lawless2015; @mayer2004]. Overall, it seems reasonable to assume, as @hooghe2015 do, that political discussions involving the father tend to revolve mostly around partisan politics, while political discussions with mothers might center on other topics --- presumably health care, education, gender issues, and so on.

Parents' education can also affect their children's interest in politics. Status transmission theory suggests that "well-educated parents are more likely to provide a politically stimulating home environment" and therefore have politically engaged children [@gidengil2016, p.&nbsp;373]. It is not clear if status transmission theory works in gendered ways. @beauregard2008 and @janmaat2022 find that parents' education has a positive relationship with their adolescents' political interest regardless of parents' gender, while other studies find a stronger effect of one parent. @neundorf2013 and @koskimaa2015 find a positive relationship between fathers' education and children's interest in politics, @sanjuan2022 instead find a positive relationship for mothers, @jennings2009 find no relationship for both fathers and mothers, and @koskimaa2015 even find a *negative* relationship between mothers' education and children's political interest.

Other parental characteristics can also influence the development of children's political interest in gendered or non-gendered ways. @gidengil2010 and @cicognani2012 find that a mother's level of political participation has a positive link with her daughter's political interest. @jennings2009 do not find a relationship between parents' income and children's political interest, but @neundorf2013 find a positive relationship between fathers' income and children's political interest, which they suggest may be due to an indirect process in which fathers' income increases children's level of education and development of higher class civic attitudes, which then increases their level of interest in politics. @borkowska2022 find that political interest transmission patterns vary between immigrant and non-immigrant families, with intergenerational transmission somewhat weaker in immigrant families than other families, notably because other factors --- such as socialization in a good-performing democracy or not --- have a more significant influence. The authors find no moderating effect for naturalization.

## Social Learning Theory

As highlighted in @sec-chap1, social learning theory suggests that children learn by observing their parents' behaviour, attitudes, habits and values, and model their own behaviour, attitudes, habits and values after them [@gidengil2016; @shehata2019]. This process applies to the transmission of political interest [@jennings2009; @prior2019; @shehata2019]. @prior2019 suggests the transmission process starts in early adolescence, when children still live in the family home but start acquiring an understanding of what politics is. Its effectiveness is higher when parents give consistent and strong cues about their own level of political interest [@jennings2009; @prior2019]. Parents' political opinions and leanings are also more likely to be known by their children when the family environment is politicized, since this environment can foster social learning [@jennings2009; @neundorf2013]. In a political home environment, children can either feel social pressure to become interested in politics to create or maintain a sense of social belonging in the family unit, be exposed to more news media content, listen to or participate in more political discussions at home, or all of these [@shehata2019]. @bandura1969 further explains that observer--model similarity leads children to model their behaviour, values, attitudes and habits on models that resemble them, notably their parent(s) of the same gender.

Social learning theory and observer--model similarity both assume that political interest should be more strongly correlated between mothers and daughters and between fathers and sons than any other combination. While this assumption has somewhat effectively been tested in past research, mothers seem to have an overall stronger influence on the development of political interest in their children. Moreover, it remains unclear if this finding applies similarly across various political topics, or if the topics in which women typically report being more interested --- health care and education --- can better be transmitted by mothers while those topics in which men typically report being more interested --- law and crime, international affairs, and partisan politics --- can better be transmitted by fathers. Four hypotheses are therefore tested alongside each other:

- ***Hypothesis 1a: Children's interest in specific political topics is more affected by political discussions with their same-gender parent(s) than other-gender parent(s).*

- ***Hypothesis 1b: Children's interest in the politics of health care and education is more affected by political discussions with their mother(s) than father(s).*

- ***Hypothesis 1c: Children's interest in law and crime, international affairs, and partisan politics is more affected by political discussions with their father(s) than mother(s).*

- ***Hypothesis 1d: Children's interest in specific political topics is more affected by political discussions with their mother(s) than father(s).*

According to Hypothesis 1a, which is the one most consistent with social learning theory, a parent's interest in a specific political topic should influence interest in that topic more strongly for their same-gender children than other-gender children. On average, a mother is expected to have more transmission potential of her interest in health care issues to her daughters than sons through political discussion, for example. Given the importance of parents' role in socialization in childhood and adolescence, this transmission process would explain part of the gender gap in interest into topics such as health care and partisan politics that exists among adults.

The other three hypotheses provide different answers to the question of how political interest gender gaps could be transmitted by parents. Hypothesis 1a supposes similar effects of parental discussions for different topics but supposes the key factor explaining interest transmission is parent--child gender congruence. To the contrary, hypotheses 1b and 1c suppose the key factor explaining interest transmission is the nature of the topics themselves --- with mothers having more transmission potential of interest in topics for which the average woman reports more interest, and the same for fathers and men's political interests. Finally, hypothesis 1d supposes that mothers have more transmission potential than fathers for all topics. Hypotheses 1a and 1d provide further tests of the findings of @beauregard2008, @owen1988, and @prior2019, while hypotheses 1b and 1c are alternate hypotheses. Since @campbellrosie2008 suggest that women tend to see political issues through the lens of cooperation while men tend to see them through a lens of self-assertion, hypotheses 1c and 1d test the suggestion that parental socialization of boys favours agency-related political topics while parental socialization of girls favours communality-related political topics.

While past studies have found no differences in the frequency of political discussions by child gender or parent gender, these studies have not studied whether the nature of political discussions in the family could vary by parent gender or child gender. Are fathers more likely than mothers to discuss agency-related political topics with their children? Are parents more likely to discuss communality-related political topics with their daughters than sons? Hypotheses 1e--1j provide hypotheses about the frequency of political discussions by gender in the family unit.

- ***Hypothesis 1e: Mothers are as likely as fathers to discuss any specific political topic with their children.*

- ***Hypothesis 1f: Mothers are more likely than fathers to discuss the politics of health care and education with their children.*

- ***Hypothesis 1g: Fathers are more likely than mothers to discuss law and crime, international affairs, and partisan politics with their children.*

- ***Hypothesis 1h: Parents are as likely to discuss any specific political topic with their sons as with their daughters.*

- ***Hypothesis 1i: Parents are more likely to discuss the politics of health care and education with their daughters than sons.*

- ***Hypothesis 1j: Parents are more likely to discuss law and crime, international affairs, and partisan politics with their sons than daughters.*

Hypotheses 1e, 1f and 1g focus their attention on discrepancies between mothers' and fathers' likelihood of discussing certain topics with their children, while hypotheses 1h, 1i and 1j focus their attention on discrepancies between daughters and sons in their likelihood of discussing politics with --- or being told about politics by --- their parents. Hypotheses 1e and 1h seek to corroborate previous findings by @mayer2004 and others that political discussion frequency in the family does not vary by gender. To the contrary, hypotheses 1f and 1g suggest that parent gender is a significant factor that explains which topics will be discussed with children, and hypotheses 1i and 1j suggest that child gender is a key variable that explains which topics they will discuss with their parents.

## Data Analysis

Several questions are then used to assess the role of parents in transmitting interest to their children. Given the important role of political discussion in the transmission of political interest, this variable is used to measure parents' transmission potential. First, the following question is asked: "For each of the following topics, which parent do you discuss most often with?" Answers are either "My mother", "My father" or "Don't know/Prefer not to answer". Second, students are asked "Among these five topics, which one do you discuss most often with your mother(s)?" Each of the five topics is listed, and the same question is then asked for the father(s). While topics such as health care and education can be spoken about without referring to their political aspects, children were given examples of political issues related to each of this topic shortly before in the same survey when they were asked which topic they were most interested in (e.g., "Health care (i.e., pandemic restrictions, working conditions of nurses)"). Nevertheless, it is possible that children will answer while keeping the non-political aspects of these topics in mind (e.g., "How was school today? Are you feeling sick?"), and results are therefore interpreted with caution for these two topics.

Relying on children's assessments of what their parents think is commonplace among both studies of parent--child political discussions and of parent--child transmission of political interest [@beauregard2008; @dostiegoulet2009en; @dowse1971; @lawless2015; @mayer2004; @oswald1998; @shehata2019; @shulman2014]. This practice can be an imperfect metric, since @beauregard2008 finds that girls tend to rate their parents' political interest more highly than boys do and @stattin2022 find that youth's political interest can cause changes in their perceptions of others' political interest.

### Parent--Child Political Discussions by Gender

@fig-parents shows, for each topic, which parent students report discussing the most often with. Out of 698 students, after removing non-answers and missing data, 82% of students report discussing health care more often with their mother than father, and 74% say the same for education. To the contrary, 64% of students say they discuss law and crime more often with their father, as well as 68% for partisan politics and 71% for international affairs. There are more non-answers and missing data for partisan politics, which presumably means no parent discusses the topic at home with their children. When these results are broken down by students' gender (see Appendix IV), very similar results are found for boys and girls.

```{r}
#| label: fig-parents
#| echo: false
#| fig-cap: "Topic most often discussed with parents, 2023 CCPIS data"
knitr::include_graphics("_graphs/ParentTopics.pdf")
```

These results match with what previous literature has found about women reporting more interest in education and health care, and men reporting more interest in partisan politics, law and crime, and international affairs. Also in line with the literature, parents seem to talk about politics just as much with their sons as with their daughters.

@fig-parentsmomdad shows the extent to which mothers discuss each of the five topics, and the same for fathers. Mothers overwhelmingly discuss education and health care according to their children, while fathers discuss international affairs more than other topics but seem like a more heterogeneous group. Despite partisan politics being the least discussed topic by both parents, fathers are much more likely to discuss it than mothers according to this metric. For law and crime too, fathers are more likely to discuss it. Again, when these results are broken down by students' gender (see Appendix IV), very similar results are found for boys and girls.

```{r}
#| label: fig-parentsmomdad
#| echo: false
#| fig-cap: "Topic most often discussed by mothers and fathers, 2023 CCPIS data"
knitr::include_graphics("_graphs/ParentTopicsMomDad.pdf")
```

The gender gap in political topics discussed is pretty stark between mothers and fathers, according to their children's assessments. It is important to specify that given the question's phrasing, it is not clear if these interactions are initiated by parents or by their children, and if it is reflective of top-down or trickle-up political socialization [@mcdevitt2002]. Moreover, the question does not ask children to specify a percentage of interactions about each topic started by the mother or father; it simply asks them to pick the parent most likely to have these discussions with them. There is no assumption that mothers are the only ones talking about health care and fathers the only ones to talk about international affairs, but on average, they are more likely to raise this topic than the other-gender parent.

Regardless, the fact that results among boys and girls both strongly point in the same direction is revealing. It makes little doubt that mothers and fathers speak differently about politics when they raise the topic with their children. Both questions were formulated in concrete ways, asking children the extent to which they discuss these topics with their parents. There is no Hawthorne effect among kids; they would not know what they are expected to answer. Parents speak in starkly different ways about politics to their children, and this confirms that political socialization is a deeply gendered process.

### Topics Parents Discuss the Most

When it comes to parents' role in political interest transmission, @tbl-lmeParentAlt shows the relationship between students' interest in each of the five topics and the gender of the parent who discusses the topic with them the most.^[Students who do not have one parent of either gender are removed from the analysis.] Children's interest in any of the five topics does not seem to be related to the gender of the parent who discusses the topic the most; all relationships are statistically insignificant. Only when aggregated, and only for boys, it seems that the gender of the parent who discusses one topic the most is related to the child's level of interest in that topic.^[Appendix V shows very similar findings without controls.]

Table: Interest in topic by gender of parent who discusses that topic the most {#tbl-lmeParentAlt}

|                                         |    All   | Health care | International affairs | Law and crime | Education | Partisan politics | 
|:----------------------------------------|---------:|------------:|----------------------:|--------------:|----------:|------------------:|
| **Results among boys**                  |          |             |                       |               |           |                   | 
| (Intercept)                             |   1.437  |    7.825    |         -3.757        |     2.568     |  36.781+  |       17.051      | 
|                                         | (11.191) |   (19.394)  |        (16.316)       |    (23.200)   |  (20.133) |      (21.989)     | 
| Mother discusses topic more than father | -0.489** |    -0.244   |         -0.380        |     -0.065    |   0.180   |       0.221       | 
|                                         |  (0.179) |   (0.468)   |        (0.397)        |    (0.432)    |  (0.442)  |      (0.524)      | 
| Age                                     |  -0.165  |    -1.156   |         1.447         |     -0.417    |  -5.153+  |       -2.456      | 
|                                         |  (1.515) |   (2.647)   |        (2.240)        |    (3.169)    |  (2.748)  |      (3.018)      | 
| Age squared                             |   0.013  |    0.048    |         -0.050        |     0.023     |   0.185+  |       0.086       | 
|                                         |  (0.051) |   (0.090)   |        (0.076)        |    (0.109)    |  (0.094)  |      (0.104)      | 
| Race (1 = white)                        |  0.463*  |    0.305    |        1.082**        |     0.171     |   -0.337  |       0.437       | 
|                                         |  (0.221) |   (0.442)   |        (0.385)        |    (0.474)    |  (0.466)  |      (0.564)      | 
| Immigrant                               |  -0.761* |    -0.443   |        -1.025*        |     -0.310    |  -1.148+  |       -1.081      | 
|                                         |  (0.319) |   (0.617)   |        (0.519)        |    (0.695)    |  (0.672)  |      (0.839)      | 
| English spoken at home                  |  -0.220  |    -0.392   |         -1.252        |     0.156     |   -0.535  |       0.630       | 
|                                         |  (0.469) |   (0.901)   |        (0.803)        |    (0.948)    |  (0.933)  |      (1.258)      | 
| French spoken at home                   |  -0.077  |    0.238    |        -0.882+        |     0.364     |   -0.076  |       0.035       | 
|                                         |  (0.276) |   (0.539)   |        (0.453)        |    (0.594)    |  (0.578)  |      (0.692)      | 
| Agency                                  | 3.009*** |   3.580**   |                       |     2.976*    |   1.450   |      4.552**      | 
|                                         |  (0.602) |   (1.163)   |                       |    (1.309)    |  (1.253)  |      (1.588)      | 
| Communality                             |  1.793*  |    0.476    |                       |     2.174     |   3.515*  |       2.407       | 
|                                         |  (0.697) |   (1.355)   |                       |    (1.527)    |  (1.466)  |      (1.749)      | 
| SD (Intercept Class)                    |   0.761  |    0.482    |         0.829         |     0.174     |   0.425   |       0.000       | 
| SD (Observations)                       |   2.549  |    2.314    |         2.413         |     2.582     |   2.634   |       2.821       | 
| Num.Obs.                                |    850   |     178     |          220          |      166      |    187    |        147        | 
| R2 Marg.                                |   0.071  |    0.087    |         0.067         |     0.072     |   0.100   |       0.101       | 
| _______________
| **Results among girls**                 |          |             |                       |               |           |                   | 
| (Intercept)                             |   6.187  |   -11.514   |        25.110+        |     14.368    |   1.183   |       28.707      | 
|                                         |  (8.435) |   (14.607)  |        (14.569)       |    (15.973)   |  (13.913) |      (24.214)     | 
| Mother discusses topic more than father |  -0.022  |    0.505    |         0.453         |     0.297     |   0.004   |       0.314       | 
|                                         |  (0.192) |   (0.518)   |        (0.494)        |    (0.475)    |  (0.486)  |      (0.561)      | 
| Age                                     |  -0.512  |    1.587    |         -3.060        |     -1.506    |   0.049   |       -3.633      | 
|                                         |  (1.158) |   (1.989)   |        (2.026)        |    (2.199)    |  (1.914)  |      (3.303)      | 
| Age squared                             |   0.025  |    -0.041   |         0.108         |     0.059     |   0.010   |       0.126       | 
|                                         |  (0.040) |   (0.068)   |        (0.070)        |    (0.076)    |  (0.066)  |      (0.113)      | 
| Race (1 = white)                        |  -0.362  |    -0.055   |         -0.424        |     -0.333    |   -0.611  |       0.034       | 
|                                         |  (0.246) |   (0.435)   |        (0.535)        |    (0.518)    |  (0.469)  |      (0.660)      | 
| Immigrant                               |  -0.390  |    -0.325   |         -0.528        |    -1.533+    |   0.517   |       -0.220      | 
|                                         |  (0.354) |   (0.628)   |        (0.777)        |    (0.835)    |  (0.667)  |      (0.876)      | 
| English spoken at home                  |  -0.354  |    -0.567   |         0.306         |     -0.849    |   0.305   |       0.304       | 
|                                         |  (0.430) |   (0.711)   |        (0.889)        |    (0.845)    |  (0.808)  |      (1.039)      | 
| French spoken at home                   |   0.120  |    0.213    |         0.185         |     -0.262    |   0.248   |       0.168       | 
|                                         |  (0.269) |   (0.485)   |        (0.594)        |    (0.595)    |  (0.517)  |      (0.717)      | 
| Agency                                  |  -0.386  |    -0.781   |         -0.256        |     0.139     |   -0.070  |       0.748       | 
|                                         |  (0.695) |   (1.279)   |        (1.588)        |    (1.558)    |  (1.358)  |      (1.812)      | 
| Communality                             |   1.164  |    1.856    |         1.938         |     1.077     |   0.115   |       -0.024      | 
|                                         |  (0.754) |   (1.295)   |        (1.607)        |    (1.621)    |  (1.499)  |      (2.027)      | 
| SD (Intercept Class)                    |   0.769  |    0.857    |         0.000         |     0.674     |   0.758   |       0.796       | 
| SD (Observations)                       |   2.521  |    2.172    |         2.631         |     2.668     |   2.416   |       2.597       | 
| Num.Obs.                                |    732   |     164     |          144          |      151      |    160    |        113        | 
| R2 Marg.                                |   0.026  |    0.082    |         0.039         |     0.046     |   0.077   |       0.022       | 

__Note:__
+ p < 0.1, * p < 0.05, ** p < 0.01, *** p < 0.001

__Note:__
Multilevel regression with random effects at the classroom level

### Topics Most Often Discussed with Mothers

@tbl-lmeMotherAlt shows students' interest in a specific topic depending on it being the topic they most often discuss with their mother(s). For both boys and girls, interest in law and crime is related to their mothers discussing law and crime (*p*<0.1 and *p*<0.001 respectively). If a girl's mother discusses mostly law and crime among the five topics, her interest in law and crime is expected to increase by a large 3.1 points on the 11-point scale. For boys, the expected increase is 1.6 points. For girls, interest in education is also related to their mothers discussing the topic at home (*p*<0.01), but not for boys. If a girl's mother discusses mostly education among the five topics, her interest in this topic is expected to increase by 0.9 points. All other relationships are non-significant at the *p*<0.05 level.

Table: Interest in topic most often discussed with one's mother {#tbl-lmeMotherAlt}

|                                         |    All   | Health care | International affairs | Law and crime | Education | Partisan politics | 
|:----------------------------------------|---------:|------------:|----------------------:|--------------:|----------:|------------------:|
| _______________
**Results among boys**  |          |             |                       |               |           |                   | 
| (Intercept)                             |   7.517  |    3.188    |         7.773         |     18.119    |   24.785  |       24.672      | 
|                                         |  (9.946) |   (17.546)  |        (17.436)       |    (17.617)   |  (18.994) |      (18.866)     | 
| Main topic discussed with mother?       |   0.059  |    0.299    |         1.481*        |     1.310+    |   0.301   |       1.520       | 
|                                         |  (0.211) |   (0.363)   |        (0.722)        |    (0.722)    |  (0.399)  |      (1.297)      | 
| Age                                     |  -1.156  |    -0.567   |         -1.093        |     -2.583    |   -3.688  |       -3.291      | 
|                                         |  (1.345) |   (2.395)   |        (2.382)        |    (2.419)    |  (2.592)  |      (2.583)      | 
| Age squared                             |   0.050  |    0.031    |         0.045         |     0.100     |   0.139   |       0.111       | 
|                                         |  (0.046) |   (0.082)   |        (0.081)        |    (0.083)    |  (0.089)  |      (0.089)      | 
| Race (1 = white)                        |  0.388+  |    0.472    |        1.246**        |     -0.027    |   -0.261  |       0.207       | 
|                                         |  (0.218) |   (0.420)   |        (0.439)        |    (0.451)    |  (0.468)  |      (0.506)      | 
| Immigrant                               |  -0.570+ |    -0.169   |        -1.419*        |     -0.661    |   -0.840  |       -0.787      | 
|                                         |  (0.313) |   (0.599)   |        (0.629)        |    (0.647)    |  (0.667)  |      (0.712)      | 
| English spoken at home                  |   0.227  |    -0.454   |         -1.329        |     0.113     |   0.970   |       0.881       | 
|                                         |  (0.524) |   (0.986)   |        (1.035)        |    (1.060)    |  (1.093)  |      (1.170)      | 
| French spoken at home                   |  -0.122  |    0.047    |        -1.631**       |     -0.057    |   0.265   |       0.245       | 
|                                         |  (0.270) |   (0.515)   |        (0.542)        |    (0.552)    |  (0.573)  |      (0.612)      | 
| Agency                                  | 2.750*** |    2.860*   |        3.540**        |     2.872*    |   1.477   |       3.325*      | 
|                                         |  (0.569) |   (1.108)   |        (1.173)        |    (1.200)    |  (1.230)  |      (1.326)      | 
| Communality                             | 2.091*** |    0.249    |        4.046**        |     1.732     |  3.617**  |       2.062       | 
|                                         |  (0.631) |   (1.219)   |        (1.268)        |    (1.320)    |  (1.344)  |      (1.443)      | 
| SD (Intercept Class)                    |   0.764  |    0.876    |         0.599         |     0.001     |   0.692   |       0.000       | 
| SD (Observations)                       |   2.589  |    2.259    |         2.424         |     2.544     |   2.550   |       2.811       | 
| Num.Obs.                                |    941   |     188     |          189          |      188      |    188    |        188        | 
| R2 Marg.                                |   0.079  |    0.081    |         0.199         |     0.111     |   0.123   |       0.081       | 
| _______________
**Results among girls** |          |             |                       |               |           |                   | 
| (Intercept)                             |  11.521  |   -11.420   |        28.362+        |     4.925     |   7.745   |      42.821**     | 
|                                         |  (7.923) |   (13.409)  |        (15.184)       |    (14.939)   |  (15.046) |      (14.364)     | 
| Main topic discussed with mother?       | 0.904*** |    0.531    |         1.442*        |    2.922***   |   0.717+  |       3.600       | 
|                                         |  (0.216) |   (0.354)   |        (0.688)        |    (0.716)    |  (0.385)  |      (2.641)      | 
| Age                                     |  -1.151  |    1.820    |        -3.506+        |     -0.210    |   -0.726  |      -5.449**     | 
|                                         |  (1.082) |   (1.834)   |        (2.076)        |    (2.045)    |  (2.054)  |      (1.962)      | 
| Age squared                             |   0.044  |    -0.053   |         0.125+        |     0.016     |   0.034   |      0.182**      | 
|                                         |  (0.037) |   (0.063)   |        (0.071)        |    (0.070)    |  (0.070)  |      (0.067)      | 
| Race (1 = white)                        |  -0.355  |    0.004    |         -0.174        |     -0.561    |   -0.612  |       -0.456      | 
|                                         |  (0.234) |   (0.427)   |        (0.479)        |    (0.490)    |  (0.467)  |      (0.480)      | 
| Immigrant                               |  -0.452  |    -0.130   |         -0.677        |    -1.981**   |   0.715   |       -0.373      | 
|                                         |  (0.313) |   (0.572)   |        (0.650)        |    (0.674)    |  (0.621)  |      (0.649)      | 
| English spoken at home                  |  -0.627  |    -0.383   |        -1.404+        |     -0.806    |   -0.090  |       -0.198      | 
|                                         |  (0.401) |   (0.701)   |        (0.770)        |    (0.758)    |  (0.789)  |      (0.736)      | 
| French spoken at home                   |  -0.113  |    -0.245   |         -0.301        |     -0.627    |   0.222   |       0.187       | 
|                                         |  (0.255) |   (0.474)   |        (0.536)        |    (0.555)    |  (0.512)  |      (0.541)      | 
| Agency                                  |   0.342  |    -0.930   |         1.228         |     0.519     |   0.299   |       1.245       | 
|                                         |  (0.620) |   (1.177)   |        (1.325)        |    (1.363)    |  (1.261)  |      (1.337)      | 
| Communality                             |   0.492  |    1.310    |         0.564         |     0.826     |   -0.910  |       0.287       | 
|                                         |  (0.657) |   (1.206)   |        (1.358)        |    (1.385)    |  (1.313)  |      (1.359)      | 
| SD (Intercept Class)                    |   0.627  |    0.624    |         0.555         |     0.210     |   0.930   |       0.000       | 
| SD (Observations)                       |   2.595  |    2.224    |         2.529         |     2.635     |   2.370   |       2.601       | 
| Num.Obs.                                |    900   |     181     |          179          |      178      |    181    |        181        | 
| R2 Marg.                                |   0.034  |    0.060    |         0.076         |     0.146     |   0.089   |       0.064       |

__Note:__
+ p < 0.1, * p < 0.05, ** p < 0.01, *** p < 0.001

__Note:__
Multilevel regression with random effects at the classroom level

### Topics Most Often Discussed with Fathers

@tbl-lmeFather shows students' interest in a specific topic depending on it being the topic they most often discuss with their father(s). Again, for both boys and girls, interest in law and crime is related to their fathers discussing law and crime (*p*<0.05 and *p*<0.001 respectively). If a girl's father discusses mostly law and crime among the five topics, her interest in law and crime is expected to increase by 1.6 points on the 11-point scale. For boys, the expected increase is 1 point. Boys' interest in partisan politics is also related to their fathers discussing the topic at home (*p*<0.01), but the same result does not apply to girls. If a boy's father discusses mostly partisan politics among the five topics, his interest in this topic is expected to increase by 1.9 points. All other relationships are non-significant at the *p*<0.05 level.

Table: Interest in topic most often discussed with one's father {#tbl-lmeFatherAlt}

|                                         |    All   | Health care | International affairs | Law and crime | Education | Partisan politics | 
|:----------------------------------------|---------:|------------:|----------------------:|--------------:|----------:|------------------:|
| _______________
**Results among boys**  |          |             |                       |               |           |                   | 
| (Intercept)                             |   5.353  |    3.085    |         0.812         |     18.145    |   28.165  |      33.127+      | 
|                                         | (10.025) |   (17.260)  |        (17.961)       |    (17.816)   |  (19.195) |      (18.514)     | 
| Main topic discussed with father?       | 0.958*** |    0.199    |         -0.187        |     1.204*    |   0.141   |      2.652***     | 
|                                         |  (0.210) |   (0.617)   |        (0.398)        |    (0.468)    |  (0.474)  |      (0.720)      | 
| Age                                     |  -0.769  |    -0.505   |         0.047         |     -2.622    |   -4.044  |      -4.448+      | 
|                                         |  (1.357) |   (2.359)   |        (2.458)        |    (2.450)    |  (2.629)  |      (2.538)      | 
| Age squared                             |   0.035  |    0.026    |         0.006         |     0.098     |   0.149   |       0.151+      | 
|                                         |  (0.046) |   (0.081)   |        (0.084)        |    (0.084)    |  (0.090)  |      (0.087)      | 
| Race (1 = white)                        |  0.371+  |    0.206    |        1.258**        |     0.191     |   -0.285  |       0.127       | 
|                                         |  (0.211) |   (0.423)   |        (0.445)        |    (0.437)    |  (0.478)  |      (0.475)      | 
| Immigrant                               |  -0.603+ |    -0.208   |        -1.400*        |     -0.050    |   -1.107  |       -0.750      | 
|                                         |  (0.307) |   (0.597)   |        (0.639)        |    (0.634)    |  (0.689)  |      (0.683)      | 
| English spoken at home                  |   0.323  |    0.307    |         -1.595        |     0.551     |   0.580   |       0.521       | 
|                                         |  (0.498) |   (0.946)   |        (1.002)        |    (0.990)    |  (1.075)  |      (1.085)      | 
| French spoken at home                   |  -0.005  |    0.462    |        -1.495**       |     0.479     |   0.075   |       0.434       | 
|                                         |  (0.279) |   (0.535)   |        (0.566)        |    (0.558)    |  (0.609)  |      (0.603)      | 
| Agency                                  | 2.883*** |   3.122**   |         2.665*        |    3.316**    |   1.562   |       3.384*      | 
|                                         |  (0.587) |   (1.151)   |        (1.221)        |    (1.208)    |  (1.306)  |      (1.302)      | 
| Communality                             |  1.371*  |    0.247    |         3.254*        |     1.635     |   3.515*  |       1.461       | 
|                                         |  (0.644) |   (1.245)   |        (1.324)        |    (1.306)    |  (1.430)  |      (1.411)      | 
| SD (Intercept Class)                    |   0.766  |    0.678    |         0.506         |     0.000     |   0.546   |       0.124       | 
| SD (Observations)                       |   2.536  |    2.275    |         2.451         |     2.476     |   2.632   |       2.664       | 
| Num.Obs.                                |    911   |     182     |          183          |      182      |    182    |        182        | 
| R2 Marg.                                |   0.080  |    0.068    |         0.145         |     0.115     |   0.094   |       0.146       | 
| _______________
**Results among girls** |          |             |                       |               |           |                   | 
| (Intercept)                             |   5.300  |    1.411    |         21.839        |     6.238     |   -6.377  |      33.506*      | 
|                                         |  (7.852) |   (12.270)  |        (14.358)       |    (13.915)   |  (14.190) |      (14.250)     | 
| Main topic discussed with father?       |  0.684** |    0.030    |         0.332         |    1.829**    |   0.381   |       -0.616      | 
|                                         |  (0.224) |   (0.533)   |        (0.442)        |    (0.575)    |  (0.438)  |      (0.841)      | 
| Age                                     |  -0.460  |    -0.076   |         -2.678        |     -0.327    |   0.855   |      -4.200*      | 
|                                         |  (1.075) |   (1.681)   |        (1.972)        |    (1.909)    |  (1.943)  |      (1.954)      | 
| Age squared                             |   0.021  |    0.014    |         0.096         |     0.018     |   -0.018  |       0.140*      | 
|                                         |  (0.037) |   (0.058)   |        (0.068)        |    (0.066)    |  (0.067)  |      (0.067)      | 
| Race (1 = white)                        |  -0.256  |    0.112    |         0.240         |     -0.678    |   -0.588  |       -0.502      | 
|                                         |  (0.237) |   (0.421)   |        (0.491)        |    (0.496)    |  (0.463)  |      (0.491)      | 
| Immigrant                               |  -0.206  |    -0.334   |         -0.647        |    -1.679*    |   1.133+  |       -0.160      | 
|                                         |  (0.339) |   (0.616)   |        (0.724)        |    (0.718)    |  (0.661)  |      (0.711)      | 
| English spoken at home                  |  -0.496  |    -0.358   |         -1.313        |     -1.256    |   0.639   |       0.254       | 
|                                         |  (0.415) |   (0.697)   |        (0.813)        |    (0.796)    |  (0.790)  |      (0.816)      | 
| French spoken at home                   |   0.158  |    0.274    |         -0.424        |     -0.299    |   0.765   |       0.470       | 
|                                         |  (0.263) |   (0.482)   |        (0.570)        |    (0.569)    |  (0.519)  |      (0.568)      | 
| Agency                                  |   0.444  |    -1.449   |         0.947         |     0.502     |   1.252   |       0.483       | 
|                                         |  (0.682) |   (1.244)   |        (1.455)        |    (1.453)    |  (1.339)  |      (1.463)      | 
| Communality                             |  1.225+  |    2.027    |         1.505         |     0.275     |   0.671   |       1.023       | 
|                                         |  (0.713) |   (1.286)   |        (1.520)        |    (1.501)    |  (1.398)  |      (1.515)      | 
| SD (Intercept Class)                    |   0.748  |    0.537    |         0.464         |     0.000     |   1.020   |       0.627       | 
| SD (Observations)                       |   2.546  |    2.186    |         2.590         |     2.625     |   2.293   |       2.547       | 
| Num.Obs.                                |    811   |     163     |          161          |      161      |    163    |        163        | 
| R2 Marg.                                |   0.030  |    0.064    |         0.050         |     0.108     |   0.101   |       0.052       | 

__Note:__
+ p < 0.1, * p < 0.05, ** p < 0.01, *** p < 0.001

__Note:__
Multilevel regression with random effects at the classroom level

Results from @tbl-lmeParent, @tbl-lmeMother and @tbl-lmeFather are somewhat intriguing. On the one hand, whichever parent discusses a topic with them has little impact on students' interest in that topic. Moreover, parents' interest in law and crime seems to be transmitted to their children in a consistent way, while interest in health care and international relations seems to have little to do with parents' discussion of these topics at home. It is relevant to note that substantively, daughters seem to be more influenced by their parents' discussion of law and crime than boys: their rating of law and crime increases by 1.6--3.1 points on the 11-point scale, compared with 1--1.7 points for boys.^[Appendix V reproduces @tbl-lmeParent, @tbl-lmeMother and @tbl-lmeFather while adding controls for age, age squared, language, immigrant status and race. The same relationships are found. Mother--daughter international affairs interest transmission also becomes marginally significant at the *p*<0.05 level.]

On the other hand, social learning theory partly seems to apply for the transmission of interest in education and partisan politics. Mothers' frequent discussion of --- and perhaps higher interest in --- education seems to trickle down to their daughters but not their sons, while fathers' discussion of education is unrelated to their children's interest in the topic. The opposite pattern can be observed for partisan politics: fathers' frequent discussion of it --- and perhaps higher interest in partisan politics --- is associated with their sons' higher interest in the topic, but not their daughters', while mothers' discussion of partisan politics is unrelated to their children's interest in the topic. Since adult women are generally more interested in education and men in partisan politics, these patterns match with the theory --- but nothing similar can be said for international affairs and health care. It may be the case that for these topics other socialization agents have more influence.

## Discussion

Descriptive results show a clear trend: mothers are much more likely to speak about education and health care, while fathers are more likely to speak about the other three topics. These trends mirror previous literature on interest in these topics. Moreover, fathers' discussion of partisan politics is related to their sons' interest in the topic, while mothers' discussion of education is related to their daughters' interest in the topic. However, other findings show the limitations of this theory, since transmission of interest in health care and international relations does not seem to happen in the family. Moreover, when students are interested in one specific topic, it does not seem that this interest is necessarily related to a parent's gender.

While results are mixed, they highlight the role parents can have on transmitting some forms of interests to their children and they reinforce the idea that, in some respects, there are different socialization routes for girls and boys.

## References
