# Gender Homophily and Social Learning: How Can Friends Shape Political Interests? {#sec-chap5}

Peers play a crucial role in children's socialization, especially during adolescence. This chapter aims to decipher peer-to-peer political interest transmission and the role gender can play in that transmission. To what extent do female peers and friends influence political interest development? Does that influence grow as children age? Do female friends have a greater influence in transmitting political interest to boys or girls? What about male peers? Do these results hold for different political topics, when political interest is measured by sector? The chapter relies on gender homophily theory and social learning. It seeks to test this dissertation's second hypothesis highlighted in @sec-chap1: *Children's interest in specific political topics is more related to political discussions with their same-gender peer(s) than other-gender peer(s).*

As @sec-chap1 lays it out, theories of gender homophily and social learning suggest that children are influenced by role models, especially those who share their gender, and this finding applies to attitudes such as political interest. Gender homophily can lead to social learning with observer--model similarity: as children and teenagers stick with other children who resemble them --- mostly of their gender --- they can start imitating their habits, attitudes and behaviours, including interests in various political topics.

## Peer Political Interest Transmission

Friends and acquaintances sometimes discuss political issues. This is especially true starting in adolescence since this period of life is associated with increased social interactions between peers [@berndt1982; @hunter1985]. While children generally report having more political discussions with parents and teachers than with their friends [@dostiegoulet2009en; @hess1967], @oswald1998 find that teenagers become more likely to discuss politics with best friends, other friends, and classmates between ages 16 and 18, while increases in political discussions with other socialization agents remain more modest. These results are confirmed by @shehata2019, who show that political discussions with parents and friends increase with age, but discussions with friends increase the most between ages 13 and 18. Both of these studies still find more political discussions happen within the family than within friend groups.

Political discussions between friends and acquaintances matter since they have been associated with higher political interest among child and adult respondents. Those who find the greatest influence for friends are probably @koskimaa2015, whose results suggest that friends' interest in politics and political discussions with these friends influence 16-to-18--year-old Finnish teenagers' political interest to a greater extent than family and school. Using the same two indicators to determine friends' influence, @furman2022 find Polish young adults are significantly influenced by their friends, controlling for the influence of other sources of socialization. They also find similar results in dozens of in-depth qualitative interviews. Among Danish teenagers and young adults, political discussion with their best friends is also associated with political interest [@levinsen2015]. @janmaat2022 also find that having politically interested friends reduces the influence of parental education on young teenagers' political interest evolution. @dostiegoulet2009en also finds 14-to-17--year-old children's political interest is explained both by their parents' political interest and by their friends' political interest, but contrary to @koskimaa2015, she finds parents are more significant in that respect. Interestingly, her results suggest that the influence of friends' political interest on personal political interest does not increase and may even decrease as children age. Another study by @klofstad2007 shows that, among first-year college students sharing their dormitory with a randomly assigned roommate, there is a strong correlation (0.46, *p*<0.001) between discussing politics and current events often with that roommate and saying discussions with that roommate have increased their political interest.

Beyond associations between peers' discussion of political issues and political interest, studies inquiring about longitudinal or causal effects have gotten more mixed results. @dostiegoulet2009en finds evidence that *changes* in political interest from one year to the next are linked to changes in the number of political discussions with friends just as much as changes in the number of such discussions with parents. @shehata2019 also show that, using a two-way fixed effects panel model, changes in children's frequency of political discussions with peers are related to their changes in political interest just as much as the frequency of political discussions with parents. However, another test they make finds that political discussions with peers at ages 14--15 do not lead to more political interest two years later, after controlling for initial levels of political interest and frequency of political discussions with parents, suggesting parents play a more long-lasting role in the socialization process over the long run. @stattin2022 further temper the results of this literature by showing that adolescents' initial level of political interest can predict changes in their perceptions of their peers' political interest, while their peers' initial political interest (as perceived by the child) does not predict changes in their own future political interest.

Other than political discussions, friends can also contribute to children's political interest development in other ways. @shehata2019 find that variations in the number of friends who are interested in what happens in the world and consume news are related to variations in political interest: the more news-watching peers, the higher the political interest. They find that this consumption of news by friends encourages children to watch more political news, but a direct long-term impact of friends' news consumption on adolescent political interest several years later is not found. Peer-group diversity can also contribute to the development of political interest. The size of one's social network has been found to have a positive relationship with political interest, while political disagreement is not related to it [@pattie2009].

## Gender Differences in Peer Transmission

Political discussions between friends typically exhibit gender effects. Among adults, leaving aside relatives, 84% of men report discussing politics only with men, while 64% of women report discussing politics only with women [@huckfeldt1995]. This could be a by-product of gender homophily since people generally spend more time with other people of the same gender. @lawless2013 find that, among United States college students aged 18--25, 20% of women and 27% of men report discussing politics frequently with friends, and a similar significant 6-point gap is found for discussion of current events with friends. @kittilson2012 concur, showing that, among 29 countries located on different continents, there is a gender gap in how often women and men discuss politics with their friends, with men being on average 7 percentage points more likely to do so than women. Among Canadian respondents, they estimate a 9-point gap. In Denmark, young men have also been found to have more political discussions with their friends than young women [@levinsen2015].

Interestingly, despite gender homophily studies finding children are more likely to discuss with other children who share their gender, older literature has not always found this to be the case for political discussions. @hess1967 find children are as likely to discuss politics with their peers regardless of their gender, while @dowse1971 find boys and girls are just as likely to have heated political arguments with their friends.

There are four main limitations to the scientific literature about peer-to-peer political discussions and political interest transmission. Each of these limitations is associated with distinct hypotheses that are tested in this chapter.

First, studies about political interest and socialization by peers have not tested the effects of peer discussions or peer political interest on chidlren's interest in political sub-topics, despite recent political interest literature often measuring interest by sub-topic given the important gender differences in interest found between issues [@campbellrosie2008; @coffe2013; @ferrin2020; @hayes1993; @kuhn2004; @sabella2004; @tormos2022; @verba1997]. This chapter's hypotheses integrate this concern by inquiring about discussions of politics with friends by topic instead of suggesting a single-item measure of political discussion and political interest.

Second, no recent study has reported the relative frequency of children's political discussions with female and male friends in any country. This chapter seeks to provide that information in the Canadian context. While research has historically not found large gender gaps in children's political discussions with friends, recent study results among adult populations have found men tend to discuss politics more often with their friends than women do. Moreover, studies about gender homophily suggest children mostly discuss with their same-gender peers in general [@shrum1988; @stehle2013]. Hypotheses 2a and 2b suggest children are more likely to discuss agency-focused topics with their male friends and communality-focused topics with their female friends.

- *__Hypothesis 2a__: Children are more likely to discuss the politics of health care and education with their female friends than male friends.*

- *__Hypothesis 2b__: Children are more likely to discuss law and crime, international affairs, and partisan politics with their male friends than female friends.*

Third, social learning theory and observer--model similarity suggest that attitudes such as political interest should be transmitted by same-gender peers more easily than by other-gender peers, an assumption yet untested in any study. Studies about social learning theory have instead mostly focused their attention on gender differences in parental transmission of political interest. As @sec-chap4 reports, gender congruence between a parent and their children plays an important role in explaining parental transmission of political interest through discussions. Most of the literature has found friends play an important role in transmitting political interest, although some recent studies such as @shehata2019 and @stattin2022 suggest the role of friends may be limited or nonexistent. Delving into the gendered aspects of the transmission process may help clarify the role friends can play in children's political interest development. Hypothesis 2c suggests that discussion of a political topic with same-gender friends should be associated with higher interest in that topic compared with discussion with other-gender friends.

- *__Hypothesis 2c__: Children's interest in specific political topics is more affected by political discussions with their same-gender friends than their other-gender friends.*

Fourth, past studies have found mixed results about whether political interest transmission by friends increases as children get older, and they have not inquired about the role of gender in that process. Since gender homophily starts at a very young age, it is not clear whether interest transmission by same-gender vs. other-gender friends becomes more prevalent as children get older. Hypothesis 2d will test the theory that political discussions with friends become more important predictors of political interest in that topic as children grow.

- *__Hypothesis 2d__: Children's interest in specific political topics becomes more and more affected by political discussions with their friends as they age.*

## Data and Methods

The 2022--23 Canadian Children Political Interest Survey (CCPIS) is used to study relationships between students' interest in certain topics and their friends' discussions of these same topics. This web-collected bilingual dataset includes survey responses from 698 Canadian children and adolescents aged 9 to 18. The CCPIS includes information about students' interest in five political topics: health care, international affairs, law and crime, education, and partisan politics. Further information about the dataset and question wording for interest questions are found in @sec-chap2.

In order to determine the importance of peers depending on their gender, the following question was asked to children: "What is the gender of most of your friends? (a) Girls; (b) Boys; (c) About the same for both genders".  Several questions are then used to assess the role of peers in transmitting interest through discussions: "Among these five topics, which one do you discuss most often with your male friends?" Each of the five topics is listed, and the same question is then asked about female friends. While topics such as health care and education can be spoken about without referring to their political aspects, children were given examples of political issues related to each of these topics shortly before in the same survey when they were asked which topic they were most interested in.

All multilevel regression results presented in this chapter include controls for socio-economic status, agency, communality, and classroom fixed effects, except if otherwise specified. A simple multilevel regression model with one explanatory variable --- often a given role model's interest in a specific topic --- is followed by multiple regression models with control variables for three blocs of variables: (1) socio-economic status variables, (2) personality traits, and (3) all other role models (parents, teachers, and influencers). Socio-economic status variables include confounding variables that have been linked with political interest and could therefore mediate any given relationship found between two persons' political interest. These include gender, age, language, ethnicity, and immigrant status. Personality traits include the agency and communality scales developed in @sec-chap2.

## Results

### Peer-to-Peer Political Discussions by Gender

@fig-peers-genderage shows the percentage of boys and girls who report having more friends of their gender, more of the other gender or about the same for both genders. Boys and girls aged 9--15 and 16--18 all say they have more friends of their gender, with percentages of 67% and 65% for younger boys and girls, respectively, against 68% and 46% of older boys and girls. On the contrary, only 3 to 8% of boys and girls of either age group report having more friends of the other gender than friends of their own gender. Finally, 28--30% of boys and girls aged 9--15 and 31% of boys aged 16--18 say they have as many friends of both genders. The percentage is higher for girls aged 16--18, at 46%.

```{r}
#| label: fig-peers-genderage
#| echo: false
#| fig-cap: "Children's Friends by Gender and Age Group, 2022 CCPIS"
knitr::include_graphics("_graphs/PeersGenderAge.pdf")
```

Gender homophily therefore holds among younger and older students, who mostly bond with students of their gender. When excluding students who say they have as many friends of both genders, no less than 95% of boys and 91% of girls report that most of their friends are of the same gender as theirs. The high proportion of girls aged 16--18 reporting having as many friends of either gender may be due to the diminishing effects gender homophily has over time, which has been reported by @shrum1988, @stehle2013 and others, although the fact that boys do not see this trend could mean gender homophily remains present for a longer time frame among them.

@fig-peers-topics shows which of the five topics is most discussed with friends of each gender among boys and girls. In a similar vein to @fig-parentsmomdad in @sec-chap4, boys and girls alike are more likely to report discussing the cooperation-focused topics of health care and education with female role models --- mothers there, female friends here. They are also more likely to report discussing the other three more assertion-focused topics of law and crime, international affairs, and partisan politics with male role models --- fathers in @sec-chap4, male friends here. These results provide evidence to support Hypotheses 2a and 2b. For each of the five topics, expectations derived from studies among adults are corroborated: children discuss education politics with female friends, international affairs with male friends, etc. A simple t-test shows that these differences are all statistically significant at the *p*<0.05 level, except partisan politics since few male and female friends discuss this topic more than all others.

```{r}
#| label: fig-peers-topics
#| echo: false
#| fig-cap: "Topic most often discussed with friends by child gender and friends' gender, 2023 CCPIS data"
knitr::include_graphics("_graphs/PeersTopicsGrey.pdf")
```

Education is the most commonly discussed topic with female friends for both boys and girls, while the most commonly discussed political topics with male friends are international affairs --- especially among boys, highest percentage at 30% --- and law and crime --- especially among girls. Partisan politics is infrequently mentioned as the most discussed topic by either boys or girls.

### Topics Most Often Discussed with Female Friends

When it comes to female friends' role in political interest transmission, @fig-femalefriends shows students' interest in a specific topic depending on it being the topic they most often discuss with their female friends. The left part of the figure shows coefficients among boys and the right part shows coefficients among girls. "All topics" is an index that aggregates all peer-to-peer linkages matched by topic. Analysis centers on the first bloc at the top, which includes no controls, as well as the fourth bloc at the bottom of the figure, which includes all controls. In the first bloc, for boys, interest in law and crime is related to their female friends discussing law and crime (*p*<0.01). At the aggregate level, interest in any of the five topics is related to discussing this topic with female friends (*p*<0.05). If a boy's female friends discuss mostly one of the five topics, his interest in that topic is expected to increase by 0.5 points on the 11-point interest scale, roughly half the size of the gender gaps in interest in international affairs (1) or partisan politics (0.9) among children. For girls, interest in law and crime (*p*<0.05), but also international affairs (*p*<0.05) and partisan politics (*p*<0.01), is associated with their interest in those matters. At the aggregate level, interest in any of the five topics is related to discussing this topic with female friends (*p*<0.001). If a girl's female friends discuss mostly one of the five topics, her interest in that topic is expected to increase by 0.9 points on the 11-point scale, similar to the gender gaps in interest in international affairs or partisan politics and higher than the 0.5-point coefficient found among boys. So far, Hypothesis 2c tentatively seems to be corroborated, with a larger effect size for girls' political discussions than girls with boys, although both effect sizes are within each other's confidence intervals.

```{r}
#| label: fig-femalefriends
#| echo: false
#| fig-cap: "Interest in topic most often discussed with one's female friends"
knitr::include_graphics("_graphs/FemaleFriendsDiscuss.pdf")
```

In the fourth bloc, after adding controls, for boys, interest in law and crime is related to their female friends discussing law and crime, and the same now goes for education (both *p*<0.01). At the aggregate level, interest in any of the five topics is related to discussing this topic with female friends (*p*<0.05). If a boy's female friends discuss mostly one of the five topics, his interest in that topic is expected to increase by 0.7 points on the 11-point scale. For girls, however, none of the relationships are statistically significant at *p*<0.05 in this model, and the coefficients for health care and law and crime are even negative. These results are opposite to what Hypothesis 2c predicts: other-gender influence on political interests is marginally significant, but not same-gender influence.

### Topics Most Often Discussed with Male Friends

@fig-malefriends shows students' interest in a specific topic depending on it being the topic they most often discuss with their male friends. The left part of the figure again shows coefficients among boys and the right part shows coefficients among girls. "All topics" is an index that aggregates all peer-to-peer linkages matched by topic. Analysis again centers on the first and fourth blocs. In the first bloc, without controls, for boys, interest in international affairs and law and crime is associated with discussions of these topics with male friends (both *p*<0.001). For girls, interest in partisan politics is associated with discussions of this topic with male friends (*p*<0.05). At the aggregate level, more discussions of any of the five topics with male friends are associated with a higher interest in that topic among both boys and girls (*p*<0.001). If a boy's male friends discuss one topic more, his interest in that topic is expected to increase by 1.5 points on the 11-point scale, larger than any gender gaps in interest reported in @sec-chap3. If a girl's male friends discuss one topic more, her interest in that topic is also expected to increase by 0.9 points on the 11-point scale. In this model, the effect of male friends seems to be twice as large for boys than girls, which is in line with Hypothesis 2c.

```{r}
#| label: fig-malefriends
#| echo: false
#| fig-cap: "Interest in Topic Most Often Discussed with One's Male Friends, 2022 CCPIS"
knitr::include_graphics("_graphs/MaleFriendsDiscuss.pdf")
```

In the fourth bloc, when adding controls for both boys and girls, the coefficient for interest in partisan politics is the highest for both genders. It is significant at the 99% confidence level among girls, similar to the model without controls, but does not quite reach statistical significance among boys. Other than partisan politics, discussions of political topics with male friends are not associated with higher sectoral political interest for the child. Discussion of health care with male friends is even associated with lower interest in health care among boys (*p*<0.05). Yet, at the aggregate level, more discussions of any of the five topics with male friends remain marginally associated with a higher interest in that topic among both boys and girls (*p*<0.05). Both coefficients are only marginally statistically significant. If a boy's male friends discuss one topic more, his interest in that topic is expected to increase by 0.6 points on the 11-point scale. If a girl's male friends discuss one topic more, her interest in that topic is also expected to increase by 0.6 points on the 11-point scale (*p*<0.05). Contrary to the model with no controls, where male friends' influence on boys was higher, in this case, both coefficients have the same size, which goes against Hypothesis 2c.^[The complete regression table for models with all controls in @fig-femalefriends and @fig-malefriends is found in @sec-chap4, Table \ref{tab:lmeAgentsCtrl}, while complete regression tables without controls can be found in Tables \ref{tab:lmeFemaleFriends} and {tab:lmeMaleFriends} in @sec-appendix7.]

Patterns from @fig-femalefriends and @fig-malefriends are somewhat complex to understand and do not seem consistent with gender homophily theory. Discussion of a political topic with peers has a significant and positive relationship with interest in that topic in most models, confirming the influence of peers in the socialization process overall, but gender patterns are not those expected. If anything, other-gender peers seem to have a larger influence than same-gender peers. Hypothesis 2c is therefore not corroborated.

### Age Trends

@fig-discusspeersyoctrl shows age patterns in peer-to-peer political interest transmission. Topic-by-topic analyses are excluded, with all analyses using the aggregate index. Interest in a political topic is related to female friends' discussion of that topic for boys aged 9--15, but not for boys aged 16--18. Similarly for the other different-gender pair, interest in a political topic is related to male friends' discussion of that same topic for girls aged 9--15, but not for boys aged 16--18. The only significant relationship among children aged 16--18 is for male friends with boys, and it barely reaches statistical significance at *p*<0.05.

```{r}
#| label: fig-discusspeersyoctrl
#| echo: false
#| fig-cap: "Interest in Topics by Gender, Age and Discussion with Peers, 2022 CCPIS"
knitr::include_graphics("_graphs/DiscussPeersYOCtrl.pdf")
```

Again, these results seem to be highly dependent on the inclusion or exclusion of certain control variables. @fig-discusspeersyo in @sec-appendix6 finds that, when controls are removed, discussions of political topics with male friends have a positive and significant influence on political interest in that topic among children of both genders and age groups. Coefficients for female friends are also all positive, but are weaker and only reach statistical significance in the case of girls aged 16 to 18. Since more coefficients are significant among older adolescents only, Hypothesis 2d is not corroborated.

## Discussion

This chapter provides mixed results regarding gendered peer influences on political interest development. On the one hand, some results confirm the presumption that friends matter and that gendered patterns appear in children's discussions of political topics. Among children who say their friends are mostly of one gender, more than 90%, both girls and boys, report they have more friends of their own gender. Moreover, female friends tend to discuss topics that are more cooperation-focused, while male friends tend to discuss topics that are more assertion-focused. In all models but one, in the aggregate, discussions with male friends or female friends of a certain topic are associated with a higher interest in that topic among children. These include models who control for background characteristics, personality traits and even the influence of other role models. Clearly, peers seem to matter in the socialization process, like past research has found. Multilevel regression models with no controls also seem to indicate children's discussion of political topics is mostly related to discussions of these topics with same-gender friends as opposed to other-gender friends, although effect sizes' confidence intervals do overlap.

On the other hand, in multilevel regression models with all controls, discussions with other-gender friends seem to have a similarly-sized influence on children's political interests than those with same-gender friends. Moreover, the influence of discussions with friends on political interescts does not seem to grow as children age.

Overall, it seems that social learning theory may not apply in the same gendered way for friends than it does for parents, despite results confirming the presence of gender homophily and gendered patterns in political discussions. It could be the case that much of the gendered transmission of interests is done by parents, while friends' influence is more limited --- and less gendered, after taking into account influences from other role models.